---
title: "RNA-Seq_A-to-I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RNA-Seq_A-to-I}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(VeloRM)
```
## RNA-seq A-to-I velocity pipeline
### data download
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE146773
### alignment
```{bash eval=FALSE, include=FALSE}
### trim
# Extract SRR numbers into an array (skip the header line)
srr_list=($(awk -F',' 'NR>1 {print $1}' /gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SraRunTable.csv))

# Check the extraction results
echo "Found ${#srr_list[@]} SRR samples:"
printf "%s\n" "${srr_list[@]}"



# Base directories
input_base="/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR"
output_base="/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR"


# Run trim_galore for each SRR sample
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    # Input FASTQ file (assuming it's in ${input_base}/${srr}/${srr}.fastq)
    # Run trim_galore (single-end mode)
    trim_galore \
        --output_dir "${output_base}/${srr}" \
        --basename "${srr}" \
        "${input_base}/${srr}/${srr}.fastq"
done

### STAR build index
(nohup STAR --runThreadN 10 --runMode genomeGenerate --genomeDir /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/star_indx/ --genomeFastaFiles /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/Homo_sapiens.GRCh38.dna.primary_assembly.fa --sjdbGTFfile /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/Homo_sapiens.GRCh38.109.gtf --sjdbOverhang 150 > /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/star_indx/build_index.out)&   




### STAR align

# Run trim_galore for each SRR sample
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    STAR --runThreadN 10 --outSAMtype BAM Unsorted\
    --readFilesIn "${input_base}/${srr}/${srr}_trimmed.fq"\
    --genomeDir /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/star_indx\
    --sjdbGTFfile /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/Homo_sapiens.GRCh38.109.gtf\
    --outFileNamePrefix "${output_base}/${srr}/${srr}"\
    --outSAMattributes All --outSAMunmapped Within\
    --sjdbOverhang 150 > "${output_base}/${srr}/${srr}.log"
done



###  sort

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools sort -@ 1 -o "${output_base}/${srr}/${srr}_sorted.bam" "${output_base}/${srr}/${srr}Aligned.out.bam"
done


### filter above mapQ30



for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools view -h -b -q 30 \
    "${input_base}/${srr}/${srr}_sorted.bam" \
        -o "${input_base}/${srr}/${srr}_above.mapQ30.bam"
done


```
### variance calling
```{bash eval=FALSE, include=FALSE}
#### call variance
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bcftools mpileup -f /gpfs/work/bio/haozhewang17/data/VeloRM_1/hg38/Homo_sapiens.GRCh38.dna.primary_assembly.fa\
    "${input_base}/${srr}/${srr}_sorted.bam" | \
    bcftools call -mv -Oz -o "${input_base}/${srr}/${srr}_calls.vcf"
done

```
### find relevant mutation
```{r eval=FALSE, include=FALSE}

sra_table <- read.csv("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SraRunTable.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

# Extract the first column which contains SRR information
srr_info <- sra_table[,1]

library(GenomicRanges)
library(vcfR)
library(Biostrings)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)

for (srr in srr_info){
  # find relevant mutation in R script
  data <- read.vcfR(paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"_calls.vcf"))
  
  data2 <- as.data.frame(data@fix)
  
  indx <- which(data2$REF == 'A' | data2$REF == 'G' | data2$REF == 'C' | data2$REF == 'T')
  
  data2 <- data2[indx,]
  
  indx <- which(data2$ALT == 'A' | data2$ALT == 'G' | data2$ALT == 'C' | data2$ALT == 'T')
  
  data2 <- data2[indx,]
  if(dim(data2)[1]==0){
    saveRDS(gr,paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"SNP.rds"))
    saveRDS(gr,paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"SNP_filtered.rds"))
  } else{
    gr <- GRanges(seqnames = paste0('chr',data2$CHROM),
                  IRanges(start = as.numeric(data2$POS),
                          width = 1),
                  strand = '+',
                  ref = data2$REF,
                  alt = data2$ALT,
                  quality = as.numeric(data2$QUAL),
                  info = data2$INFO)
    
    gr <- keepStandardChromosomes(gr,pruning.mode = "coarse")
    
    ref <- as.character(DNAStringSet(Views(Hsapiens,gr)))
    
    indx <- which(gr$ref != ref)
    
    if(length(indx) != 0){
      gr <- gr[-indx]
    }
    read.vcfR(paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"_calls.vcf"))
    saveRDS(gr,paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"SNP.rds"))  
    
    # filter mutation
    transcript <- transcripts(TxDb.Hsapiens.UCSC.hg38.knownGene)
    
    transcript_plus <- transcript[which(strand(transcript)== "+")]
    
    transcript_minus <- transcript[which(strand(transcript)== "-")]
    
    gr <- readRDS(paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"SNP.rds"))
    
    gr_all <- GRanges()
    
    # 1.	Coordinate  (+) strand: A
    gr_plus <- gr
    olp <- findOverlaps(gr_plus,transcript_plus)
    q <- queryHits(olp)
    s <- subjectHits(olp)
    gr_plus <- gr_plus[q]
    gr_plus$transcript_overlapped <- transcript_plus[s]$tx_name
    if( length(gr_plus$transcript_strand)==0){
      gr_plus=NULL
    }else{
      gr_plus$transcript_strand <- '+'
      gr_plus <- unique(gr_plus)
    }
    
    
    
    indx <- which(gr_plus$ref == "A")
    gr_plus <- gr_plus[indx]
    indx <- which(gr_plus$alt == 'G')
    gr_plus <- gr_plus[indx]
    
    # 2.	Coordinate  (-) strand: 
    gr_minus <- gr
    strand(gr_minus) <- '-'
    
    olp <- findOverlaps(gr_minus,transcript_minus)
    q <- queryHits(olp)
    s <- subjectHits(olp)
    gr_minus <- gr_minus[q]
    gr_minus$transcript_overlapped <- transcript_minus[s]$tx_name
    if( length(gr_minus$transcript_strand)==0){
      gr_minus=NULL
    }else{
      gr_minus$transcript_strand <- '-'
      gr_minus <- unique(gr_minus)
    }
    
    if(!is.null(gr_minus)){
      strand(gr_minus) <- '+'
      indx <- which(gr_minus$ref == 'T')
      gr_minus <- gr_minus[indx]
      indx <- which(gr_minus$alt == 'C')
      gr_minus <- gr_minus[indx]
    }
    
    
    gr <- c(gr_plus,gr_minus)
    saveRDS(gr,paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/",srr,"SNP_filtered.rds"))
    print(paste0("finish",srr))
  }
  
  
}

```
### divide bam
```{bash eval=FALSE, include=FALSE}

############ bedtools to divide the bam file
###### File prepare
## get exon-no-ambiguity bed by substract (exon.bed - intron.bed = exon-na.bed)
bedtools subtract -a UCSC_Exons.bed -b UCSC_Introns.bed > exon-na.bed

## get intron-no-ambituity bed by subtract (inton.bed - exon.bed = inron-na.bed)
bedtools subtract -a UCSC_Introns.bed -b UCSC_Exons.bed > intron-na.bed

###### divide the file 


#### mRNA.bam
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/UCSC_Introns.bed\
    -v -wa -split > "${input_base}/${srr}/${srr}_mRNA_temp.bam"
done

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_mRNA_temp.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/UCSC_Exons.bed\
    -wa -split > "${input_base}/${srr}/${srr}_mRNA.bam"
    samtools index "${input_base}/${srr}/${srr}_mRNA.bam"
done
  
 #### mRNA_spliced.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools view -h "${input_base}/${srr}/${srr}_mRNA.bam"\
    | awk '$0 ~ /^@/ || $6 ~ /N/' | samtools view -b >\
    "${input_base}/${srr}/${srr}_mRNA_spliced.bam"
    samtools index "${input_base}/${srr}/${srr}_mRNA_spliced.bam"
done

  
  #### pre-mRNA.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/intron-na.bed\
    -wa -split > "${input_base}/${srr}/${srr}_pre-mRNA.bam"
    samtools index "${input_base}/${srr}/${srr}_pre-mRNA.bam"
done

  
  #### isoform-a.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/UCSC_Introns.bed\
    -wa -split > "${input_base}/${srr}/${srr}_maybe_pre_mRNA.bam"
done

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_maybe_pre_mRNA.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/intron-na.bed\
    -v -wa -split > "${input_base}/${srr}/${srr}_isoform-a.bam"
    samtools index "${input_base}/${srr}/${srr}_isoform-a.bam"
done


```
### base counting
```{r eval=FALSE, include=FALSE}


### base counting
sra_table <- read.csv("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SraRunTable.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

# Extract the first column which contains SRR information
srr_info <- sra_table[,1]

library(GenomicRanges)
library(Biostrings)
library(GenomicAlignments)
library(Rsamtools)


gr_all <- readRDS('/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SNP_all.rds')[,-c(1:5)]
seqlevels(gr_all) <- gsub('chr','',seqlevels(gr_all))

## base counting
j <- 0
for (srr in srr_info) {
  
  time_0 <- Sys.time()
  j <- j+1
  name <- srr
  path <- paste0("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/",srr,"/")
  
  ################################## mRNA_spliced.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_mRNA_spliced.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  time_1 <- Sys.time()
  ################################## pre-mRNA.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_pre-mRNA.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  time_2 <- Sys.time()
  ################################## isoform.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_isoform-a.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
    time_3 <- Sys.time()
  
  print(j)
  print(paste0("spliced mRNA runing time", time_1-time_0))
  print(paste0("pre-mRNA runing time", time_2-time_1))
    print(paste0("isoform-mRNA runing time", time_3-time_2))
}


### combine A-to-I


sra_table <- read.csv("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SraRunTable.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

# Extract the first column which contains SRR information
srr_info <- sra_table[,1]


# Initialize the spliced_list
spliced_list <- list()
j = 1
for (srr in srr_info) {
  name <- srr
  path <- paste0('/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/', srr,"/")
  spliced_list[[j]] <- readRDS(paste0(path, srr, "_mRNA_spliced.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of spliced_list
names(spliced_list) <- srr_info
saveRDS(spliced_list, "/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/spliced_A_to_I.rds")

# Initialize the pre_list
pre_list <- list()
j = 1
for (srr in srr_info) {
  name <- srr
  path <- paste0('/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/', srr,"/")
  pre_list[[j]] <- readRDS(paste0(path, srr, "_pre-mRNA.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of isoform_list
names(pre_list) <- srr_info
saveRDS(pre_list, "/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/pre_A_to_I.rds")


# Initialize the isoform_list
isoform_list <- list()
j = 1
for (srr in srr_info) {
  name <- srr
  path <- paste0('/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/', srr,"/")
  isoform_list[[j]] <- readRDS(paste0(path, srr, "_isoform-a.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of pre_list
names(isoform_list) <- srr_info
saveRDS(isoform_list, "/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/SRR/isoform_A_to_I.rds")

```

We will get the 
Processed data:

SNP_all

Frequency data: isoform_A_to_I, pre_A_to_I, spliced_A_to_I
## pre-process
### extract the modification read counts and un-modification read counts
```{r eval=FALSE, include=FALSE}

##  pre



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/pre_A_to_I.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_pre.rds")
saveRDS(frequency_all,"reads_counts_pre.rds")
saveRDS(error_reads_list,"error_reads_list_pre.rds")
saveRDS(methylation_rate,"methylation_rate_pre.rds")



##  spliced



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/spliced_A_to_I.rds")



testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_spliced.rds")
saveRDS(frequency_all,"reads_counts_spliced.rds")
saveRDS(error_reads_list,"error_reads_list_spliced.rds")
saveRDS(methylation_rate,"methylation_rate_spliced.rds")


##  isoform



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/isoform_A_to_I.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_isoform.rds")
saveRDS(frequency_all,"reads_counts_isoform.rds")
saveRDS(error_reads_list,"error_reads_list_isoform.rds")
saveRDS(methylation_rate,"methylation_rate_isoform.rds")
```
### calculate the error rate for site
```{r eval=FALSE, include=FALSE}
# calculate teh meta_error_rate for each site
## spliced

rm(list = ls())
gc()
error_reads_list_spliced <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_reads_list_spliced.rds")

error_reads_list_spliced_sum <- Reduce("+", error_reads_list_spliced)

reads_counts_spliced <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_spliced.rds")

summed_vectors_list <- lapply(reads_counts_spliced, function(df) rowSums(df))

reads_counts_spliced_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_spliced <- error_reads_list_spliced_sum/
  (error_reads_list_spliced_sum+reads_counts_spliced_sum) 


saveRDS(error_rate_meta_spliced,"/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_rate_meta_spliced.rds")
## pre


rm(list = ls())
gc()
error_reads_list_pre <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_reads_list_pre.rds")

error_reads_list_pre_sum <- Reduce("+", error_reads_list_pre)

reads_counts_pre <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_pre.rds")

summed_vectors_list <- lapply(reads_counts_pre, function(df) rowSums(df))

reads_counts_pre_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_pre <- error_reads_list_pre_sum/
  (error_reads_list_pre_sum+reads_counts_pre_sum) 

saveRDS(error_rate_meta_pre,"/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_rate_meta_pre.rds")

```

### filter the site 
From the initial analysis of 1,152 cells, we identified 48,556 potential editing sites. After applying stringent quality filters that excluded sites with error rates exceeding 1%, we retained 1,159 high-confidence A-to-I editing sites. To ensure data reliability, we further filtered the dataset by removing cells with fewer than 200 reads in either spliced or unspliced mRNA fractions, resulting in a final set of 1,119 cells for downstream analysis.

```{r eval=FALSE, include=FALSE}
# Load necessary library
library(VeloRM)
library(ggplot2)

# Read metadata
error_rate_meta_spliced <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_rate_meta_spliced.rds")
error_rate_meta_pre <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/error_rate_meta_pre.rds")
index_error <- intersect(which(error_rate_meta_spliced <= 0.01), which(error_rate_meta_pre <= 0.01))



# Read counts
reads_counts_spliced<- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_spliced.rds")

# Process spliced data
spliced_meth_test <- as.data.frame(lapply(reads_counts_spliced, function(df) df[["methylated_reads"]]))
spliced_unmeth_test <- as.data.frame(lapply(reads_counts_spliced, function(df) df[["unmethylated_reads"]]))


reads_counts_pre<- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_pre.rds")


# Process spliced data
pre_meth_test <- as.data.frame(lapply(reads_counts_pre, function(df) df[["methylated_reads"]]))
pre_unmeth_test <- as.data.frame(lapply(reads_counts_pre, function(df) df[["unmethylated_reads"]]))


# Set row names
rownames(spliced_meth_test) <- rownames(spliced_unmeth_test) <- paste0("site_", 1:dim(spliced_unmeth_test)[1])

rownames(pre_meth_test) <- rownames(pre_unmeth_test)  <- paste0("site_", 1:dim(pre_unmeth_test)[1])

# Calculate spliced and pre-mRNA test matrices
spliced_test <- spliced_meth_test+spliced_unmeth_test


pre_test <- pre_meth_test + pre_unmeth_test

#discard the cell have lower reads
index_cell_discard_spliced <- which(colSums(spliced_test)<=200)
index_cell_discard_pre <- which(colSums(pre_test)<=200)

spliced_test_discarded <- spliced_test[,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_test_discarded <- pre_test[,-union(index_cell_discard_spliced,index_cell_discard_pre)]




index <- index_error

spliced_meth_test_processed <- spliced_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
spliced_unmeth_test_processed <- spliced_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_meth_test_processed <- pre_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_unmeth_test_processed <- pre_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]


saveRDS(union(index_cell_discard_spliced,index_cell_discard_pre),"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/index_cell_discard.rds")
saveRDS(index_error,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/index_error.rds")
saveRDS(index,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/index.rds")


reads_counts_isoform <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_isoform.rds")
reads_counts_isoform <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/reads_counts_isoform.rds")

# Process pre-mRNA data
isoform_meth_test <- as.data.frame(lapply(reads_counts_isoform, function(df) df[["methylated_reads"]]))
isoform_unmeth_test <- as.data.frame(lapply(reads_counts_isoform, function(df) df[["unmethylated_reads"]]))



rownames(isoform_meth_test) <- rownames(isoform_unmeth_test) <-  paste0("site_", 1:dim(isoform_unmeth_test)[1])

isoform_meth_test_processed <- isoform_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
isoform_unmeth_test_processed <- isoform_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]

test.list <- list(spliced_meth_test_processed,spliced_unmeth_test_processed,pre_meth_test_processed,pre_unmeth_test_processed,
                  isoform_meth_test_processed,isoform_unmeth_test_processed)
control.list <- NULL

res_preprocess_0.5 <- methylation.sites.preprocess(test.list,control.list,0.5)
res_preprocess <- methylation.sites.preprocess(test.list,control.list)

saveRDS(test.list,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/test.list.rds")
saveRDS(control.list,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/control.list.rds")
saveRDS(res_preprocess_0.5,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_preprocess_0.5.rds")
saveRDS(res_preprocess,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
# 
SNP <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/origin/SNP_all.rds")
SNP_index <- SNP[index]
saveRDS(SNP_index,"D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/SNP_1159.rds")
```
### DESeq2 to test the differential methylation between the spliced mRNA and pre-mRNA
```{r eval=FALSE, include=FALSE}
# Load input data
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data

# Set random seed for reproducibility
set.seed(42)

# Record start time
start_time <- Sys.time()

# Load DESeq2 package
library(DESeq2)

# Define DESeq2 differential expression analysis function
DESeq2_test <- function(meth_control, meth_test, unmeth_control, unmeth_test) {
  # Convert input data to data frames
  meth_control <- data.frame(meth_control)
  meth_test <- data.frame(meth_test)
  unmeth_control <- data.frame(unmeth_control)
  unmeth_test <- data.frame(unmeth_test)
  
  # Function to calculate size factor
  sizeFactor <- function(data) {
    # Ensure data is not less than 0
    data <- as.matrix(data)
    
    # Log-transform the data
    log_data <- log(data)
    log_data[is.infinite(log_data)] <- NA
    log_mean <- rowMeans(log_data, na.rm = TRUE)
    log_s <- log_data - log_mean
    
    # Calculate size factor
    s_size <- exp(apply(log_s, 2, function(x) median(x, na.rm = TRUE)))
    return(s_size)
  }
  
  # Calculate size factors for spliced and unspliced data
  spliced_total <- meth_control + unmeth_control
  unspliced_total <- meth_test + unmeth_test
  
  spliced_size_factor <- sizeFactor(spliced_total)
  unspliced_size_factor <- sizeFactor(unspliced_total)
  
  # Get the number of control and test samples
  num_control <- dim(unmeth_control)[2]
  num_test <- dim(unmeth_test)[2]
  
  # Combine all data
  counts <- cbind(meth_control, meth_test, unmeth_control, unmeth_test)
  
  # Design the experiment matrix
  design = data.frame(
    Trt = rep(c(rep("control", num_control), rep("test", num_test)), 2), 
    Meth = c(rep("meth", (num_control + num_test)), rep("unmeth", (num_control + num_test)))
  )
  
  # Define the model
  model = ~ Meth + Trt + Meth:Trt
  
  # Create DESeqDataSet object
  dds <- DESeqDataSetFromMatrix(countData = counts,
                                colData = design,
                                design = model)
  
  # Set size factors
  sizeFactors(dds) = rep(c(spliced_size_factor, unspliced_size_factor), 2)
  
  # Perform differential expression analysis
  dds <- DESeq(dds, test = "Wald")
  
  # Extract results
  res = DESeq2::results(dds, name = "Methunmeth.Trttest", cooksCutoff = FALSE)
  
  return(res)
}

# Call the function to perform differential expression analysis
result <- DESeq2_test(spliced.meth.test, unspliced.meth.test, spliced.unmeth.test, unspliced.unmeth.test)

# Record end time
end_time <- Sys.time()

# Print the runtime
print(end_time - start_time)


# Save the results
saveRDS(result, "/gpfs/work/bio/haozhewang17/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_DESeq2.rds")
```

We will get the 
Processed data:
SNP_1159: the SNP information for the A-to-I site for further analysis.
test_list: the test cell reads counts (spliced_meth_test,spliced_unmeth_test,pre_meth_test,pre_unmeth_test,
                   isoform_meth_test,isoform_unmeth_test)
control_list: NULL
res_preprocess: the list processed by VeloRM
res_DESeq2: the DESeq2 results
## analysis
### splicing junction distance of the SNP sites
```{r eval=FALSE, include=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(dplyr)
library(tidyr)

snp <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/SNP_1159.rds")
snp_pos <- snp@ranges@start
snp_chr_length <- snp@seqnames@lengths
snp_chr_seq <- snp@seqnames@values
snp_chr <- as.character(rep(snp_chr_seq,snp_chr_length)) 
snp_strand <- snp@elementMetadata@listData[["transcript_strand"]]
snp_data <- as.data.frame(cbind(snp_chr,snp_pos,snp_strand))
snp_data$snp_pos <- as.numeric(snp_data$snp_pos)




# Read the GTF file from Ensembl
gtf_file <- "D:/data/VeloRM_1/RNA-Seq_A-to-I/origin/gencode.v47.chr_patch_hapl_scaff.annotation.gtf"
gtf_data <- import(gtf_file)

# Filter for exons
exons <- gtf_data[gtf_data$type == "exon"]

# Convert to data frame
exons_df <- data.frame(
chr = as.character(seqnames(exons)),
start = start(exons),
end = end(exons),
strand = as.character(strand(exons)),
gene_id = mcols(exons)$gene_id,
transcript_id = mcols(exons)$transcript_id
)

# Sort exons by transcript and position
exons_df <- exons_df %>%
arrange(transcript_id, chr, start)

# Extract splicing junctions using intron boundaries
junctions <- exons_df %>%
group_by(transcript_id) %>%
arrange(if_else(strand == "+", start, -start)) %>%  # Sort based on strand
mutate(
  donor_site = end,                # End of the upstream exon
  acceptor_site = lead(start)      # Start of the downstream exon
) %>%
filter(!is.na(acceptor_site)) %>%   # Remove the last exon (no downstream exon)
ungroup() %>%
select(chr, donor_site, acceptor_site, gene_id, transcript_id, strand)


junctions_unique <- junctions %>%
 distinct(chr, donor_site, acceptor_site, .keep_all = TRUE)



junctions_long <- junctions_unique %>%
 pivot_longer(cols = c(donor_site, acceptor_site), 
              names_to = "site_type", 
              values_to = "site_position") %>%
 select(chr, site_position, site_type, gene_id, transcript_id, strand)



# Calculate distance from each SNP to the nearest junction (matching chr and strand)
snp_with_distance <- snp_data %>%
 rowwise() %>%
 mutate(
   nearest_junction_distance = min(
     abs(junctions_long$site_position[junctions_long$chr == snp_chr & junctions_long$strand == snp_strand] - snp_pos),
     na.rm = TRUE
   )
 )


# get normal chr
standard_chromosomes <- paste0("chr", c(1:22, "X", "Y"))

# get the exon with the normal chr
exons_filtered <- exons[seqnames(exons) %in% standard_chromosomes]



# calculate the distance 
exon_lengths <- width(exons_filtered)

# calculate the prob 
prob_weights <- exon_lengths / sum(exon_lengths)

# random sample
set.seed(123) 


#random_distances_list <- list()

#for (i in 1:10){
 selected_exons <- sample(seq_along(exons_filtered), size = 1159, replace = TRUE, prob = prob_weights)
 # choose the site in the select the exons
 random_positions <- start(exons[selected_exons]) + 
   floor(runif(1159, min = 0, max = exon_lengths[selected_exons]))
 
 # create the data frame
 random_sites <- data.frame(
   chr = as.character(seqnames(exons[selected_exons])),
   position = random_positions,
   strand = as.character(strand(exons[selected_exons]))
 )
 
 
 # Calculate distance from each SNP to the nearest junction (matching chr and strand)
 random_sites_with_distance <- random_sites %>%
   rowwise() %>%
   mutate(
     nearest_junction_distance = min(
       abs(junctions_long$site_position[junctions_long$chr == chr & junctions_long$strand == strand] - position),
       na.rm = TRUE
     )
   )
 random_distances_list  <- random_sites_with_distance$nearest_junction_distance
#}



random_distances_df <- data.frame(
 distance = random_distances_list,
 group = "Random"
)

# merge the data frame
real_distances_df <- data.frame(
 distance = snp_with_distance$nearest_junction_distance,
 group = "Real"
)


all_distances_df <- bind_rows(random_distances_df, real_distances_df)
group_order <- c("Real","Random")
all_distances_df$group <- factor(all_distances_df$group, levels = group_order)
library(ggplot2)
#  Boxplot
plot <- ggplot(all_distances_df, aes(x = group, y = log10(distance+1), fill = group)) +
 geom_violin(trim = FALSE, alpha = 0.5) +
 geom_boxplot(width = 0.1, outlier.shape = NA)+
 scale_fill_manual(values = c("red",rep("lightblue", 10))) +
 labs(title = NULL,
      x = NULL,
      y = "log10(Distance (bp)+1)") +
 theme_minimal() +
 theme(
       plot.title = element_text(hjust = 0.5),
       legend.position = "none")+coord_cartesian(ylim = c(0,5))


saveRDS(all_distances_df,"D:/data/VeloRM_1/RNA-Seq_A-to-I/results/splicing_junction_analysis/all_distances_df.rds")

```
#### visulization
We can see that the SNP is close to the splicing junction comparing to random sample
```{r}
library(ggplot2)
all_distances_df <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/results/splicing_junction_analysis/all_distances_df.rds")
#  Boxplot
plot <- ggplot(all_distances_df, aes(x = group, y = log10(distance+1), fill = group)) +
 geom_violin(trim = FALSE, alpha = 0.5) +
 geom_boxplot(width = 0.1, outlier.shape = NA)+
 scale_fill_manual(values = c("red",rep("lightblue", 10))) +
 labs(title = NULL,
      x = NULL,
      y = "log10(Distance (bp)+1)") +
 theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5),
       legend.position = "none")+coord_cartesian(ylim = c(0,5))
plot

```
### overall analysis
#### overall reads counts
```{r}
# Load input data
test.list <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data



spliced.meth.sum <- sum(spliced.meth.test)
spliced.unmeth.sum <- sum(spliced.unmeth.test)
unspliced.meth.sum <- sum(unspliced.meth.test)
unspliced.unmeth.sum <- sum(unspliced.unmeth.test)


# spliced.meth.sum, spliced.unmeth.sum, unspliced.meth.sum, unspliced.unmeth.sum

# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Create a data frame
data <- data.frame(
  category = c("mature mRNA", "pre mRNA"),  # spliced = mature mRNA, unspliced = pre mRNA
  meth = c(spliced.meth.sum, unspliced.meth.sum),      # methylated counts
  unmeth = c(spliced.unmeth.sum, unspliced.unmeth.sum) # unmethylated counts
)

# Calculate proportions
data <- data %>%
  mutate(
    total = meth + unmeth,                # total counts for each category
    meth_ratio = meth / total             # methylation proportion
  )

# Calculate mature mRNA (spliced) proportion of total
spliced_ratio <- (spliced.meth.sum + spliced.unmeth.sum) / 
                 (spliced.meth.sum + spliced.unmeth.sum + 
                  unspliced.meth.sum + unspliced.unmeth.sum)

# Visualization 1: Pie chart of mature mRNA vs pre mRNA proportion
pie_data <- data.frame(
  group = c("mature mRNA", "pre mRNA"),
  value = c(spliced_ratio, 1 - spliced_ratio)
)

ggplot(pie_data, aes(x = "", y = value, fill = group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(value*100, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(title = "mature mRNA/(mature mRNA + pre mRNA) Ratio") +
  theme_void()

# Visualization 2: Bar plot of methylation ratio in each category
ggplot(data, aes(x = category, y = meth_ratio, fill = category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(meth_ratio*100, 1), "%")), vjust = -0.5) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.52)) +
  labs(title = "Methylation Ratio in Spliced and Unspliced",
       y = "Methylation Ratio",
       x = "") +
  theme_minimal()

```


#### read the DESeq2 results
```{r}
res_DESeq2 <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_DESeq2.rds")

log2FC <- res_DESeq2@listData[["log2FoldChange"]]
pvals <- res_DESeq2@listData[["pvalue"]]

plot_data <- data.frame(log2FC = log2FC, 
                       pvalue = pvals,
                       neg_log10_p = -log10(pvals))


ggplot(plot_data, aes(x = log2FC, y = neg_log10_p)) +
  # Add light gray background for -1 < log2FC < 1 (non-significant FC region)
  annotate("rect", xmin = -1, xmax = 1, ymin = 0, ymax = 30, 
           fill = "gray90", alpha = 0.3) +
  
  # Color points based on significance and log2FC:
  # - Red: log2FC > 1 & p < 0.05 (significantly up-regulated)
  # - Blue: log2FC < -1 & p < 0.05 (significantly down-regulated)
  # - Gray: Not significant or |log2FC| ≤ 1
  geom_point(aes(color = case_when(
    log2FC > 1 & pvalue < 0.05 ~ "Spliced",
    log2FC < -1 & pvalue < 0.05 ~ "Unspliced",
    TRUE ~ "Not significant"
  )), alpha = 0.6) +
  
  # Manual color scale (red, blue, gray)
  scale_color_manual(
    values = c(
      "Spliced" = "red",
      "Unspliced" = "blue",
      "Not significant" = "gray50"
    ),
    name = "Regulation"  # Legend title
  ) +
  
  # Add dashed lines for thresholds:
  # - Vertical: log2FC = ±1
  # - Horizontal: p-value = 0.05 (-log10(0.05) ≈ 1.3)
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  
  # Axis labels and theme
  labs(
    x = "log2 Fold Change", 
    y = "-log10(p-value)",
    title = NULL
  ) +
  theme_minimal() +
  
  # Set axis limits (zoomed-in view)
  coord_cartesian(ylim = c(0, 30), xlim = c(-4, 4))

index_DESeq2_significant <- which(res_DESeq2@listData$pvalue<=0.05)


# the meta OR infomation
res_preprocess <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")




library(ggplot2)
library(tidyr)




spliced_num_sites <- length(which(res_preprocess$spliced_methylation_M_value_meta-res_preprocess$unspliced_methylation_M_value_meta>=1&res_DESeq2@listData$pvalue<=0.05))
unspliced_num_sites <- length(which(res_preprocess$unspliced_methylation_M_value_meta-res_preprocess$spliced_methylation_M_value_meta>=1&res_DESeq2@listData$pvalue<=0.05))
Other_num_sites <- length(res_preprocess$spliced_methylation_M_value_meta) - spliced_num_sites -unspliced_num_sites

# Load necessary package
library(ggplot2)

# Create a data frame for pie chart
df_pie <- data.frame(
  Type = c("Spliced", "Unspliced", "Other"),
  Count = c(spliced_num_sites, unspliced_num_sites, Other_num_sites)
)

# Calculate percentages
df_pie$Percentage <- round(df_pie$Count / sum(df_pie$Count) * 100, 1)

# Create separate data frames for each piece of text you want to display
df_pie$text_count <- as.character(df_pie$Count)
df_pie$text_percentage <- paste0(df_pie$Percentage, "%")

# Plot the pie chart with 4 separate texts
pie_chart <- ggplot(df_pie, aes(x = "", y = Count, fill = Type)) +
  geom_bar(stat = "identity", width = 1) +  
  coord_polar("y", start = 0) +  # Convert to pie chart
  geom_text(aes(label = text_count), 
            position = position_stack(vjust = 0.15), size = 5) +  # Position for the number
   geom_text(aes(label = text_percentage), 
            position = position_stack(vjust = 0.85), size = 5) +  # Position for the percentage
  labs(title = "Significant Sites Proportion") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank(),   # Remove x-axis labels
    axis.ticks = element_blank(),    # Remove x-axis ticks
    panel.grid = element_blank(),    # Remove grid
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, vjust = 0, size = 14),  # Move title outside center
    legend.title = element_text(size = 10),  # Larger legend title text
    legend.text = element_text(size = 8),  # Larger legend labels
    legend.key.size = unit(0.8, "cm")  # Increase the size of the legend keys (colors)
  )
pie_chart
```

### regulatory mechanisms

#### Find the site that is differentially methylated in spliced and pre-mRNA.
```{r}
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(MetaTX)
library(ggplot2)
library(dplyr)

res_DESeq2 <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_DESeq2.rds")


index_differentially_meth <- which(res_DESeq2@listData$pvalue<=0.05)


```

#### Finding sites where methylation levels are high in pre-mRNA but low in spliced mRNA, as well as sites where methylation levels are high in spliced mRNA but low in pre-mRNA.
```{r}
res_preprocess <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
spliced_methylation_M_value_meta <- res_preprocess$spliced_methylation_M_value_meta
unspliced_methylation_M_value_meta <- res_preprocess$unspliced_methylation_M_value_meta


index_pre_high_spliced_low <- which(unspliced_methylation_M_value_meta-spliced_methylation_M_value_meta>=1)
index_spliced_high_pre_low <- which(spliced_methylation_M_value_meta-unspliced_methylation_M_value_meta>=1)
```


#### metaTX plot 
```{r}
intersect_pre_dominate <- intersect(index_pre_high_spliced_low,index_differentially_meth)

intersect_spliced_dominate <- intersect(index_spliced_high_pre_low,index_differentially_meth)

index_not_dominate <- setdiff(1:1159,union(intersect_pre_dominate,intersect_spliced_dominate)) 

# metaTX


SNP_1159 <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/SNP_1159.rds")
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene


remap_results_pre_dominate <- remapCoord(features = SNP_1159[intersect_pre_dominate], txdb = txdb)
remap_results_spliced_dominate <- remapCoord(features = SNP_1159[intersect_spliced_dominate], txdb = txdb)
remap_results_not_dominate <- remapCoord(features = SNP_1159[index_not_dominate], txdb = txdb)
  
# Step 2: Create plots for each subset
plot_pre_dominate <- directPlot(remap_results_pre_dominate)
plot_spliced_dominate <- directPlot(remap_results_spliced_dominate)
plot_not_dominate <- directPlot(remap_results_not_dominate)

data_a <- ggplot_build(plot_pre_dominate)$data[[1]] %>% mutate(Group = "Precursor")
data_b <- ggplot_build(plot_spliced_dominate)$data[[1]] %>% mutate(Group = "Mature")
data_c <- ggplot_build(plot_not_dominate)$data[[1]] %>% mutate(Group = "Other")

# Combine all datasets
combined_data <- bind_rows(data_a, data_b, data_c)#, data_intersect)

region_rects <- data.frame(
    xmin = c(0.2, 0.4, 0.6),  # Start positions of each region
    xmax = c(0.4, 0.6, 0.8),  # End positions of each region
    ymin = min(combined_data$y, na.rm = TRUE) - 0.1,  # Adjust below data range
    ymax = min(combined_data$y, na.rm = TRUE) - 0.05, # Slightly higher to make room
    Region = c("5' UTR", "CDS", "3' UTR")
)

# Calculate the height of 5' UTR and 3' UTR regions
utr_height <- region_rects$ymax[region_rects$Region == "5' UTR"] - region_rects$ymin[region_rects$Region == "5' UTR"]

# Increase the height of CDS region by 20% (can be adjusted as needed)
cds_height <- utr_height * 2 # Increase height by 30%

# Set ymin and ymax for CDS so that it is centered with respect to 5' UTR and 3' UTR
region_rects$ymin[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] - 0.5*utr_height  # Decrease ymin for CDS
region_rects$ymax[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] + cds_height  # Increase ymax for CDS


compare_metaTX_plot_2 <-ggplot(combined_data, aes(x = x, y = y, color = Group))+
  geom_line(size=1)+
    # Add region highlights using geom_rect, with consistent positioning across all facets
    geom_rect(data = region_rects, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = Region),
              inherit.aes = FALSE, color = "black", alpha = 0.4, 
              linewidth = 0.5) +  # Default linewidth for all regions
    # Set fill for CDS region to grey (lighter than black)
    geom_rect(data = subset(region_rects, Region == "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE, fill = "grey", color = "black", alpha = 0.4, linewidth = 0.5) +  # Grey fill for CDS
    # Adjust 5' UTR and 3' UTR fills to a very dark grey, closer to black
    geom_rect(data = subset(region_rects, Region != "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                                                                fill = factor(Region, levels = c("5' UTR", "3' UTR"))), 
              inherit.aes = FALSE, color = "black", alpha = 0.3, linewidth = 0.5) + 
    scale_fill_manual(values = c("5' UTR" = grey(0.15),  # Very dark grey for 5' UTR, closer to black
                                "CDS" = "grey",           # Grey for CDS
                                "3' UTR" = grey(0.15))) +  # Very dark grey for 3' UTR
    # Adjust y position for the text annotations to place them above the rectangles
    geom_text(data = region_rects, aes(x = (xmin + xmax) / 2, y = ymax - 0.5, label = Region),
              inherit.aes = FALSE, color = "black", size = 3, hjust = 0.5) +
    # Remove Region fill legend
    guides(fill = "none")+scale_color_manual(values = c(
  "Other" = "#00ba38",      
  "Mature" = "#619cff",     
  "Precursor" = "#F8766d"         
),name = "Higher on")+theme_minimal()+
  theme(
    legend.position = c(0.8, 0.7)  )+xlab(NULL)+ylab(NULL)


saveRDS(intersect_pre_dominate,"D:/data/VeloRM_1/RNA-Seq_A-to-I/results/regulatory_mechanisms/index_pre_dominate.rds")
saveRDS(intersect_spliced_dominate,"D:/data/VeloRM_1/RNA-Seq_A-to-I/results/regulatory_mechanisms/index_spliced_dominate.rds")
saveRDS(index_not_dominate,"D:/data/VeloRM_1/RNA-Seq_A-to-I/results/regulatory_mechanisms/index_not_dominate.rds")
```

#### motif plot
```{r}
library(GenomicRanges)
library(Biostrings)
library(BSgenome)
library(BSgenome.Hsapiens.UCSC.hg38)
genome <- BSgenome.Hsapiens.UCSC.hg38

## pre 
SNP_intersect_pre_dominate <- SNP_1159[intersect_pre_dominate]
# transcript_strand  +
SNP_intersect_pre_dominate_plus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="+")] 


extended_range_intersect_pre_dominate_plus <- resize(SNP_intersect_pre_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_pre_dominate_plus <- getSeq(genome, extended_range_intersect_pre_dominate_plus)



# transcript_strand  -
SNP_intersect_pre_dominate_minus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="-")] 

extended_range_intersect_pre_dominate_minus <- resize(SNP_intersect_pre_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_pre_dominate_minus <- getSeq(genome, extended_range_intersect_pre_dominate_minus)



## spliced
SNP_intersect_spliced_dominate <- SNP_1159[intersect_spliced_dominate]
# transcript_strand  +
SNP_intersect_spliced_dominate_plus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="+")] 

extended_range_intersect_spliced_dominate_plus <- resize(SNP_intersect_spliced_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_spliced_dominate_plus <- getSeq(genome, extended_range_intersect_spliced_dominate_plus)



# transcript_strand  -
SNP_intersect_spliced_dominate_minus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="-")] 

extended_range_intersect_spliced_dominate_minus <- resize(SNP_intersect_spliced_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_spliced_dominate_minus <- getSeq(genome, extended_range_intersect_spliced_dominate_minus)








## not 
SNP_intersect_not_dominate <- SNP_1159[index_not_dominate]
# transcript_strand  +
SNP_intersect_not_dominate_plus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="+")] 


extended_range_intersect_not_dominate_plus <- resize(SNP_intersect_not_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_not_dominate_plus <- getSeq(genome, extended_range_intersect_not_dominate_plus)



# transcript_strand  -
SNP_intersect_not_dominate_minus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="-")] 

extended_range_intersect_not_dominate_minus <- resize(SNP_intersect_not_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_not_dominate_minus <- getSeq(genome, extended_range_intersect_not_dominate_minus)





sequence_intersect_pre_dominate_minus_rc <- reverseComplement(sequence_intersect_pre_dominate_minus)
sequence_intersect_pre_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus_rc))
sequence_intersect_pre_dominate_plus_rc <- reverseComplement(sequence_intersect_pre_dominate_plus)
sequence_intersect_pre_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus_rc))


sequence_intersect_pre_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus))
sequence_intersect_pre_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus))



sequence_intersect_spliced_dominate_minus_rc <- reverseComplement(sequence_intersect_spliced_dominate_minus)
sequence_intersect_spliced_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus_rc))
sequence_intersect_spliced_dominate_plus_rc <- reverseComplement(sequence_intersect_spliced_dominate_plus)
sequence_intersect_spliced_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus_rc))


sequence_intersect_spliced_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus))
sequence_intersect_spliced_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus))

sequence_intersect_not_dominate_minus_rc <- reverseComplement(sequence_intersect_not_dominate_minus)
sequence_intersect_not_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus_rc))
sequence_intersect_not_dominate_plus_rc <- reverseComplement(sequence_intersect_not_dominate_plus)
sequence_intersect_not_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus_rc))



sequence_intersect_not_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus))
sequence_intersect_not_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus))


sequence_intersect_pre_dominate <- c(sequence_intersect_pre_dominate_plus,sequence_intersect_pre_dominate_minus_rc)

sequence_intersect_spliced_dominate <- c(sequence_intersect_spliced_dominate_plus,sequence_intersect_spliced_dominate_minus_rc)

sequence_intersect_not_dominate <- c(sequence_intersect_not_dominate_plus,sequence_intersect_not_dominate_minus_rc)



library(ggseqlogo)
seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate),
    "pre_dominate_plus" = as.character(sequence_intersect_pre_dominate_plus),
    "spliced_dominate_plus" = as.character(sequence_intersect_spliced_dominate_plus),
    "not_dominate_plus" = as.character(sequence_intersect_not_dominate_plus),
    "pre_dominate_minus" = as.character(sequence_intersect_pre_dominate_minus),
    "spliced_dominate_minus" = as.character(sequence_intersect_spliced_dominate_minus),
    "not_dominate_minus" = as.character(sequence_intersect_not_dominate_minus)
)
ggseqlogo(seq_list, method = 'bits') + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")

seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate))
ggseqlogo(seq_list, method = 'bits', seq_type = "RNA") + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")  +
  theme_grey() 
ggseqlogo(seq_list, method = 'prob', seq_type = "RNA") + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")  +
  theme_grey() 

library(universalmotif)

motif_list <- lapply(seq_list,function(x) create_motif(x, alphabet = "RNA"))
```




#### find the index of site only detect methylated reads in spliced or pre mRNA
```{r}
index_meth_only_detect_pre <-  which(res_preprocess[["vg.test"]] %in% res_preprocess[["index"]][["methylation only occurs at unspliced RNA"]])

index_meth_only_detect_spliced <-  which(res_preprocess[["vg.test"]] %in%res_preprocess[["index"]][["methylation only occurs at spliced RNA"]])

# only detect at spliced
index_meth_only_detect_spliced_int <-  intersect(index_meth_only_detect_spliced,intersect_spliced_dominate)
SNP_1159[index_meth_only_detect_spliced_int]

library(GenomicRanges)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)  # For converting Entrez IDs to Gene Symbols

# Assume your data is in a data.frame containing chr and pos
# Example:
# snp_data <- data.frame(chr = c("chr17", "chr21"), pos = c(4797798, 25597391))

# Use TxDb to find overlapping genes
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene  # Load hg38 gene annotations from UCSC
genes <- genes(txdb)  # Get genomic coordinates of all genes

# Find genes overlapping with SNPs
overlaps <- findOverlaps(SNP_1159[index_meth_only_detect_spliced_int], genes)
gene_ids <- genes[subjectHits(overlaps)]$gene_id  # Extract Entrez IDs

# Convert Entrez IDs to Gene Symbols
gene_symbols_meth_only_detect_spliced_int <- mapIds(org.Hs.eg.db,
                       keys = as.character(gene_ids),
                       keytype = "ENTREZID",
                       column = "SYMBOL")

gene_symbols_meth_only_detect_spliced_int


# only detect at pre
index_meth_only_detect_pre_int <- intersect(index_meth_only_detect_pre,intersect_pre_dominate)
SNP_1159[index_meth_only_detect_pre_int]


# Find genes overlapping with SNPs
overlaps <- findOverlaps(SNP_1159[index_meth_only_detect_pre_int], genes)
gene_ids <- genes[subjectHits(overlaps)]$gene_id  # Extract Entrez IDs

# Convert Entrez IDs to Gene Symbols
gene_symbols_meth_only_detect_pre_int <- mapIds(org.Hs.eg.db,
                       keys = as.character(gene_ids),
                       keytype = "ENTREZID",
                       column = "SYMBOL")

gene_symbols_meth_only_detect_pre_int
```

### cell cycle trajectory using A-to-I velocity
#### run velocity
```{r,warning=FALSE}
library(ggplot2)
res_preprocess <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
# the input data
test.list <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/test.list.rds")
control.list <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/control.list.rds")


spliced.meth.test <- test.list[[1]]
spliced.unmeth.test <- test.list[[2]]
unspliced.meth.test <- test.list[[3]]
unspliced.unmeth.test <- test.list[[4]]

kCells = 1; deltaT =1; kSites = 1
library(VeloRM)
# calculate the emat_size and nmat_size

emat_size <- colSums(spliced.meth.test + spliced.unmeth.test)
nmat_size <- colSums(unspliced.meth.test + unspliced.unmeth.test)

res <- methylation.site.relative.velocity.estimates(res_preprocess$test, res_preprocess$control,
                                                    kCells = kCells, deltaT = deltaT, kSites = kSites, 
                                                    fit.quantile = 0.05, n.cores = 1, emat.size = emat_size, 
                                                    nmat.size = nmat_size)

# saveRDS(res, "D:/data/VeloRM_1/RNA-Seq_A-to-I/results/cell_cycle/res_A-to-I_velocity.rds")

```


#### load the embedding
```{r,warning=FALSE}
sra_table <- read.csv("D:/data/VeloRM_1/RNA-Seq_A-to-I/SRR/SraRunTable.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

# Extract the first column which contains SRR information
srr_info <- sra_table[,1]
phase_info <- sra_table[,8]
GSM_info <- sra_table[,16]

index_cell_discard <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/index_cell_discard.rds")

label_phase <- phase_info[-index_cell_discard]
srr_info_filter <- srr_info[-index_cell_discard]
GSM_info_filter <- GSM_info[-index_cell_discard]
label_phase <- label_phase[order(GSM_info_filter)]
srr_info_filter <- srr_info_filter[order(GSM_info_filter)]
GSM_info_filter <- GSM_info_filter[order(GSM_info_filter)]



sample_info <- read.csv("D:/data/VeloRM_1/RNA-Seq_A-to-I/SRR/sample.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

plate_info_filter <- sample_info[sample_info[,1]%in%GSM_info_filter,2]
plate_info_filter <- sub("^Single U2OS cell ", "", plate_info_filter)
GSM_info_filter2 <- sample_info[sample_info[,1]%in%GSM_info_filter,1]




embedding_info <- read.csv("D:/data/VeloRM_1/RNA-Seq_A-to-I/SRR/Embedding.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

index_with_embedding <- which(plate_info_filter%in%embedding_info[,1])

embedding_x <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),3]
names(embedding_x) <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),1]

embedding_y <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),4]
names(embedding_y) <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),1]

embedding_h <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),2]
names(embedding_h) <- embedding_info[which(embedding_info[,1]%in%plate_info_filter[index_with_embedding]),1]


srr_info_filter <- srr_info_filter[index_with_embedding]
label_phase_filter <- label_phase[index_with_embedding]


umap_data <- data.frame(cbind(embedding_x,embedding_y,label_phase_filter,srr_info_filter,embedding_h))
colnames(umap_data) <- c("x","y","color","srr","time")
umap_data$x <- as.numeric(umap_data$x)
umap_data$y <- as.numeric(umap_data$y)
umap_data$time <- as.numeric(umap_data$time)

umap_data <- umap_data[which(umap_data$color%in%c("G1","S-ph","G2M")),]




ggplot()+geom_point(data=umap_data,aes(x=x,y=y,color=color))

ggplot()+geom_point(data=umap_data,aes(x=x,y=y,color=time))+
  scale_color_gradientn(colors = c("blue", "green", "yellow", "red"))

# 
# saveRDS(umap_data, "D:/data/VeloRM_1/RNA-Seq_A-to-I/results/cell_cycle/umap_data_origin.rds")

```
#### filter
```{r,warning=FALSE}

SNP_1159 <- readRDS("D:/data/VeloRM_1/RNA-Seq_A-to-I/pre_process/SNP_1159.rds")

names(SNP_1159) <- rownames(test.list[[1]])

spliced_methylation_Beta_value_meta_index <- res_preprocess[["spliced_methylation_Beta_value_meta"]]
unspliced_methylation_Beta_value_meta_index <-res_preprocess[["unspliced_methylation_Beta_value_meta"]]


Beta_not_1or0 <- apply(res[[3]][["current"]]$Beta,1,function(x) length(which(x!=0)))
index_site_name_int_intersect <-   names(which(Beta_not_1or0>=1))


Beta_not_1or0_cell <- apply(res[[3]][["current"]]$Beta[index_site_name_int_intersect,],2,function(x) length(which(x!=0)))
index_cell_phase <-(which(Beta_not_1or0_cell>=1))

```


#### trajectory Beta
```{r,warning=FALSE}



umap_results <- umap_data[,1:2]
rownames(umap_results) <- umap_data[,4]
umap_color <-umap_data$color
names(umap_color) <- umap_data[,4]

ggplot() +
  geom_point(
    data = umap_data[which(rownames(umap_results) %in% names(index_cell_phase)), ],
    aes(x = x, y = y, color = time),
    size = 2,
    alpha = 0.7
  ) +scale_color_gradientn(
  colors = c("#00008B", "#1E90FF", "#00BFFF", "#00FF00", "#FFFF00", "#FF8C00", "#FF0000"),
  name = "Time"
)+theme_minimal()


arrow.scale=0.01; cell.alpha=0.4; cell.cex=1; fig.height=4;

vis_umap_Beta_emb_Beta_linear_index  <- show.velocity.on.embedding.cor(umap_results[which(rownames(umap_results)%in%names(index_cell_phase)),],
             res[[3]][["current"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             res[[3]][["projected"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             res[[3]][["delta"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             cell.colors=(umap_color[which(rownames(umap_results)%in%names(index_cell_phase))]),n=500,scale='linear',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,rm_uniform="mode1")
vis_umap_Beta_emb_Beta_linear_index[[1]]

arrow.scale=0.02; cell.alpha=0.4; cell.cex=1; fig.height=4;
vis_umap_Beta_emb_Beta_linear_index_grid <- show.velocity.on.embedding.cor(umap_results[which(rownames(umap_results)%in%names(index_cell_phase)),],
             res[[3]][["current"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             res[[3]][["projected"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             res[[3]][["delta"]]$Beta[index_site_name_int_intersect,intersect(rownames(umap_results),names(index_cell_phase))],
             cell.colors=(umap_color[which(rownames(umap_results)%in%names(index_cell_phase))]),n=500,scale='linear',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,show.grid.flow = TRUE,grid.n = 20,rm_uniform="mode1")
vis_umap_Beta_emb_Beta_linear_index_grid[[1]]+theme_minimal()

```

