---
title: "RNA-Seq_A-to-I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RNA-Seq_A-to-I}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(VeloRM)
```
## RNA-seq A-to-I velocity pipeline
### data download
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE99933
### alignment
```{bash eval=FALSE}
### trim
# Extract SRR numbers into an array (skip the header line)
srr_list=($(find /gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR -maxdepth 1 -type d -name "*SRR*" -exec basename {} \;))

# Check the extraction results
echo "Found ${#srr_list[@]} SRR samples:"
printf "%s\n" "${srr_list[@]}"

# alignment

# Base directories
input_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"
output_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"


# Run trim_galore for each SRR sample
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    # Input FASTQ file (assuming it's in ${input_base}/${srr}/${srr}.fastq)
    # Run trim_galore (single-end mode)
    trim_galore \
        --output_dir "${output_base}/${srr}" \
        --basename "${srr}" \
        "${input_base}/${srr}/${srr}.fastq"
done

### STAR build index

### STAR align

for srr in "${srr_list[@]}"; do 
    echo "Processing sample: $srr"
    STAR --runThreadN 1 --outSAMtype BAM Unsorted\
    --readFilesIn "${input_base}/${srr}/${srr}_trimmed.fq"\
    --genomeDir /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39/star_indx\
    --sjdbGTFfile /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39/mm39.knownGene.gtf\
    --outFileNamePrefix "${output_base}/${srr}/${srr}"\
    --outSAMattributes All --outSAMunmapped Within\
    --sjdbOverhang 150 > "${output_base}/${srr}/${srr}.log"
done


###  sort




# Base directories
input_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"
output_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"


for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools sort -@ 1 -o "${output_base}/${srr}/${srr}_sorted.bam" "${output_base}/${srr}/${srr}Aligned.out.bam"
done



### filter above mapQ30



for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools view -h -b -q 30 \
    "${input_base}/${srr}/${srr}_sorted.bam" \
        -o "${input_base}/${srr}/${srr}_above.mapQ30.bam"
done



```

### divide bam
```{bash eval=FALSE}
############ bedtools to divide the bam file
###### File prepare
## get exon-no-ambiguity bed by substract (exon.bed - intron.bed = exon-na.bed)
bedtools subtract -a /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Exons_mm39.bed -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Introns_mm39.bed > /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/exon-na.bed

## get intron-no-ambituity bed by subtract (inton.bed - exon.bed = inron-na.bed)
bedtools subtract -a /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Introns_mm39.bed -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Exons_mm39.bed > /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/intron-na.bed



# Extract SRR numbers into an array (skip the header line)
srr_list=($(find /gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR -maxdepth 1 -type d -name "SRR*" ! -name "SRR" -exec basename {} \;))

# Check the extraction results
echo "Found ${#srr_list[@]} SRR samples:"
printf "%s\n" "${srr_list[@]}"



# Base directories
input_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"
output_base="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"




#### mRNA.bam
for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Introns.bed\
    -v -wa -split > "${input_base}/${srr}/${srr}_mRNA_temp.bam"
done

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_mRNA_temp.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Exons.bed\
    -wa -split > "${input_base}/${srr}/${srr}_mRNA.bam"
    samtools index "${input_base}/${srr}/${srr}_mRNA.bam"
done
  
 #### mRNA_spliced.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    samtools view -h "${input_base}/${srr}/${srr}_mRNA.bam"\
    | awk '$0 ~ /^@/ || $6 ~ /N/' | samtools view -b >\
    "${input_base}/${srr}/${srr}_mRNA_spliced.bam"
    samtools index "${input_base}/${srr}/${srr}_mRNA_spliced.bam"
done

  
  #### pre-mRNA.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/intron-na.bed\
    -wa -split > "${input_base}/${srr}/${srr}_pre-mRNA.bam"
    samtools index "${input_base}/${srr}/${srr}_pre-mRNA.bam"
done

  
  #### isoform-a.bam

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_above.mapQ30.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/UCSC_Introns.bed\
    -wa -split > "${input_base}/${srr}/${srr}_maybe_pre_mRNA.bam"
done

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bedtools intersect -a "${input_base}/${srr}/${srr}_maybe_pre_mRNA.bam"\
    -b /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39_bed/intron-na.bed\
    -v -wa -split > "${input_base}/${srr}/${srr}_isoform-a.bam"
    samtools index "${input_base}/${srr}/${srr}_isoform-a.bam"
done


#### we can also do further to clean the splicing reads missing classify to ambiguous and pre-mRNA, and merge the splicing reads to mRNA.
  
  
#  for srr in "${srr_list[@]}"; do
#    echo "Processing sample: $srr"
#    samtools view -h "${input_base}/${srr}/${srr}_pre-mRNA.bam"\
#    | awk '$0 ~ /^@/ || $6 ~ /N/' | samtools view -b >\
#    "${input_base}/${srr}/${srr}_pre-mRNA_spliced.bam"
#    samtools index "${input_base}/${srr}/${srr}_pre-mRNA_spliced.bam"
#  done
  
#  for srr in "${srr_list[@]}"; do
#    echo "Processing sample: $srr"
#    samtools view -h "${input_base}/${srr}/${srr}_isoform-a.bam"\
#    | awk '$0 ~ /^@/ || $6 ~ /N/' | samtools view -b >\
#    "${input_base}/${srr}/${srr}_isoform-a_spliced.bam"
#    samtools index "${input_base}/${srr}/${srr}_isoform-a_spliced.bam"
#  done
  
  
#  for srr in "${srr_list[@]}"; do
#    echo "Processing sample: $srr"
#    bedtools subtract -a "${input_base}/${srr}/${srr}_pre-mRNA.bam"\
#     -b "${input_base}/${srr}/${srr}_pre-mRNA_spliced.bam"\
#      > "${input_base}/${srr}/${srr}_pre-mRNA_clean.bam"
#  done
  
#  for srr in "${srr_list[@]}"; do
#    echo "Processing sample: $srr"
#    bedtools subtract -a "${input_base}/${srr}/${srr}_isoform-a.bam"\
#     -b "${input_base}/${srr}/${srr}_isoform-a_spliced.bam"\
#      > "${input_base}/${srr}/${srr}_isoform-a_clean.bam"
#  done
  
#  for srr in "${srr_list[@]}"; do
#    echo "Processing sample: $srr"
#    samtools merge "${input_base}/${srr}/${srr}_mRNA_spliced_combine.bam"\
#     "${input_base}/${srr}/${srr}_pre-mRNA_spliced.bam"\
#     "${input_base}/${srr}/${srr}_isoform-a_spliced.bam"\
#     "${input_base}/${srr}/${srr}_mRNA_spliced.bam"
#  done


```
### variance calling
```{bash eval=FALSE}
#### call variance

for srr in "${srr_list[@]}"; do
    echo "Processing sample: $srr"
    bcftools mpileup -f /gpfs/work/bio/haozhewang17/data/VeloRM_1/mm39/mm39.fa\
    "${input_base}/${srr}/${srr}_sorted.bam" | \
    bcftools call -mv -Oz -o "${input_base}/${srr}/${srr}_calls.vcf"
done

```
### find relevant mutation
```{r eval=FALSE}
# Extract SRR sample folder names into a character vector (skip the header line)
srr_info <- list.dirs("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", recursive = FALSE, full.names = FALSE)
srr_info <- srr_info[grepl("SRR", srr_info)]  # Keep only directories containing "SRR"

# Load required packages
library(GenomicRanges)
library(vcfR)
library(Biostrings)
library(TxDb.Mmusculus.UCSC.mm39.knownGene)  # Changed to mouse mm39
library(BSgenome.Mmusculus.UCSC.mm39)        # Changed to mouse mm39

# Loop over each SRR sample
for (srr in srr_info) {
  # Load VCF file
  data <- read.vcfR(paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", srr, "/", srr, "_calls.vcf"))
  data2 <- as.data.frame(data@fix)

  # Keep entries with valid nucleotide bases
  data2 <- data2[data2$REF %in% c('A', 'C', 'G', 'T') & data2$ALT %in% c('A', 'C', 'G', 'T'), ]

  if (nrow(data2) == 0) {
    # If no valid SNPs, save empty object
    saveRDS(NULL, paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", srr, "/", srr, "SNP.rds"))
  } else {
    # Create GRanges object for SNPs
    gr <- GRanges(seqnames = paste0(data2$CHROM),
                  ranges = IRanges(start = as.numeric(data2$POS), width = 1),
                  strand = '+',
                  ref = data2$REF,
                  alt = data2$ALT,
                  quality = as.numeric(data2$QUAL),
                  info = data2$INFO)

    # Keep only standard chromosomes
    gr <- keepStandardChromosomes(gr, pruning.mode = "coarse")

    # Validate reference bases with genome
    ref <- as.character(DNAStringSet(Views(Mmusculus, gr)))
    gr <- gr[gr$ref == ref]

    saveRDS(gr, paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", srr, "/", srr, "SNP.rds"))

    # Load transcript annotation
    transcript <- transcripts(TxDb.Mmusculus.UCSC.mm39.knownGene)
    transcript_plus <- transcript[strand(transcript) == "+"]
    transcript_minus <- transcript[strand(transcript) == "-"]

    # Filter A>G mutations on the + strand
    gr_plus <- gr
    olp <- findOverlaps(gr_plus, transcript_plus)
    gr_plus <- gr_plus[queryHits(olp)]
    gr_plus$transcript_overlapped <- transcript_plus[subjectHits(olp)]$tx_name
    if (length(gr_plus) > 0) {
      gr_plus$transcript_strand <- '+'
      gr_plus <- unique(gr_plus)
      gr_plus <- gr_plus[gr_plus$ref == "A" & gr_plus$alt == "G"]
    } else {
      gr_plus <- NULL
    }

    # Filter T>C mutations on the - strand (reverse-complement)
    gr_minus <- gr
    strand(gr_minus) <- '-'
    olp <- findOverlaps(gr_minus, transcript_minus)
    gr_minus <- gr_minus[queryHits(olp)]
    gr_minus$transcript_overlapped <- transcript_minus[subjectHits(olp)]$tx_name
    if (length(gr_minus) > 0) {
      gr_minus$transcript_strand <- '-'
      gr_minus <- unique(gr_minus)
      strand(gr_minus) <- '+'
      gr_minus <- gr_minus[gr_minus$ref == "T" & gr_minus$alt == "C"]
    } else {
      gr_minus <- NULL
    }

    # Combine both strand results and save
    gr_filtered <- c(gr_plus, gr_minus)
    saveRDS(gr_filtered, paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", srr, "/", srr, "SNP_filtered.rds"))
  }

  print(paste0("Finished processing ", srr))
}

# Combine all filtered SNPs across samples
gr_all <- GRanges()
for (srr in srr_info) {
  gr <- readRDS(paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/", srr, "/", srr, "SNP_filtered.rds"))
  if (!is.null(gr)) {
    gr_all <- c(gr_all, gr)
  }
}
gr_all <- unique(gr_all)

# Save combined SNP results
saveRDS(gr_all, "/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/SNP_all.rds")


```
### base counting
```{r eval=FALSE}
# Extract SRR sample folder names into a character vector (skip the header line)
srr_info <- list.dirs("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR", recursive = FALSE, full.names = FALSE)
srr_info <- srr_info[grepl("SRR", srr_info)]  # Keep only directories containing "SRR"

library(GenomicRanges)
library(Biostrings)
library(GenomicAlignments)
library(Rsamtools)


gr_all<- readRDS('/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/SNP_all.rds')


gr_all <-gr_all[,-c(1:5)]


## base counting
j <- 0
for (srr in srr_info) {
  
  time_0 <- Sys.time()
  j <- j+1
  name <- srr
  path <- paste0("/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR/",srr,"/")
  
  ################################## mRNA_spliced.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_mRNA_spliced.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  time_1 <- Sys.time()
  ################################## pre-mRNA.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_pre-mRNA.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  time_2 <- Sys.time()
  ################################## isoform.bam
  ##################################
  bamfilePath <- paste0(path,srr,'_isoform-a.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
    time_3 <- Sys.time()
  
  print(j)
  print(paste0("spliced mRNA runing time", time_1-time_0))
  print(paste0("pre-mRNA runing time", time_2-time_1))
    print(paste0("isoform-mRNA runing time", time_3-time_2))
}

```
### summary the count
```{bash eval=FALSE}


module load samtools/1.16.1-gcc-8.5.0-teyetiz 



root_dir="/gpfs/work/bio/haozhewang17/data/RNA-Seq_A-to-I_SRR_GSE99933/SRR"


output_file="bam_read_counts.csv"



echo "Folder,isoform-a.bam,mRNA_spliced.bam,pre-mRNA.bam" > "$output_file"


for folder in "$root_dir"/SRR*; do
    if [ -d "$folder" ]; then
        folder_name=$(basename "$folder")
        echo "Processing folder: $folder_name"
        
        
        isoform_a_count="Not found"
        mrna_spliced_count="Not found"
        pre_mrna_count="Not found"
        
        #isoform-a.bam
        isoform_a_file=$(find "$folder" -name "*_isoform-a.bam" | head -n 1)
        if [ -n "$isoform_a_file" ] && [ -f "$isoform_a_file" ]; then
            isoform_a_count=$(samtools view -c "$isoform_a_file")
        fi
        
        #mRNA_spliced.bam
        mrna_spliced_file=$(find "$folder" -name "*_mRNA_spliced.bam" | head -n 1)
        if [ -n "$mrna_spliced_file" ] && [ -f "$mrna_spliced_file" ]; then
            mrna_spliced_count=$(samtools view -c "$mrna_spliced_file")
        fi
        
        #pre-mRNA.bam
        pre_mrna_file=$(find "$folder" -name "*_pre-mRNA.bam" | head -n 1)
        if [ -n "$pre_mrna_file" ] && [ -f "$pre_mrna_file" ]; then
            pre_mrna_count=$(samtools view -c "$pre_mrna_file")
        fi
        
        #
        echo "$folder_name,$isoform_a_count,$mrna_spliced_count,$pre_mrna_count" >> "$output_file"
    fi
done

echo "Processing complete. Results saved to $output_file"
```

We will get the 
Processed data:

SNP_all

bam_read_counts

Frequency data: isoform_A_to_I, pre_A_to_I, spliced_A_to_I

## Overall reads counts
```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

counts_data <- read.csv("D:/data/VeloRM_data/RNA-Seq_A-to-I/origin/bam_read_counts.csv", 
                       header = TRUE,  # Assumes first row contains column names
                       stringsAsFactors = FALSE)  # Prevents automatic conversion to factors

# View the first few rows

counts_data_sum <- apply(counts_data,1,function(x) sum(as.numeric(x[2:4])))
isoform_prop <- as.numeric(counts_data[,2])/counts_data_sum
spliced_prop <- as.numeric(counts_data[,3])/counts_data_sum
pre_prop <- as.numeric(counts_data[,4])/counts_data_sum

df <- data.frame(
  `mature RNA` = spliced_prop,
  `pre RNA` = pre_prop,
  `ambiguous` = isoform_prop,
  check.names = FALSE
)
df_long <- pivot_longer(df, cols = everything(), 
                       names_to = "Type", 
                       values_to = "Probability")

box_plot <- ggplot(df_long, aes(x = "", y = Probability, fill = Type)) +  
    geom_violin(alpha = 0.6, width = 0.7) +
    geom_boxplot(width = 0.2, alpha = 0.8, 
                 outlier.shape = NA,
                 coef = 0) + 
    facet_wrap(~Type, scales = "free_y", ncol = 3) +  
    labs(
        title = NULL,
        x = NULL,  
        y = "Probability"
    ) +
    # Use theme_gray() which is the default ggplot2 theme
    theme_gray(base_size = 14)  +
    scale_fill_manual(
        values = c("mature RNA" = "#FF9999",  
                   "pre RNA" = "#99CCFF",     
                   "ambiguous" = "#99CC99"),   
        breaks = c("mature RNA", "pre RNA", "ambiguous") 
    )+
  theme(
    legend.position = "none"
  )

box_plot

# Create a data frame with the summed proportions
plot_data <- data.frame(
  category = c("Spliced", "Unspliced"),
  proportion = c(sum(counts_data[,3]), sum(counts_data[,4]))
)

# Create the barplot
barplot <- ggplot(plot_data, aes(x = category, y = proportion, fill = category)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_text(aes(label = round(proportion, 2)), vjust = -0.5, size = 5) + # Add values on top
  scale_fill_manual(values = c("Spliced" = "salmon", "Unspliced" = "steelblue")) +
  labs(title = NULL,
       x = NULL,
       y = "Total read counts") +
  theme_grey(base_size = 14) + # Using theme_grey with larger base font size
  theme(legend.position = "none", # Remove legend
        plot.title = element_text(hjust = 0.5)) +# Center title
  coord_cartesian(ylim = c(0,3.5e7))
barplot

res_preprocess <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
spliced_index <- which(res_preprocess[["spliced_methylation_M_value_meta"]]>=1)
unspliced_index <- which(res_preprocess[["unspliced_methylation_M_value_meta"]]>=1)


spliced_index_l <- length(spliced_index)
unspliced_index_l <- length(unspliced_index)
intersect_l <- length(intersect(spliced_index,unspliced_index))

library(VennDiagram)

venn <-draw.pairwise.venn(
  area1 = spliced_index_l,
  area2 = unspliced_index_l,
  cross.area = intersect_l,
  category = c("Spliced", "Unspliced"),
  fill = c("#E74C3C","skyblue"),
  alpha = 0.5,
  lty = "blank",
  cex = 2,
  cat.cex = 1.5
)
grid.newpage()
grid.draw(venn)

```
## pre-process
### extract the modification read counts and un-modification read counts
```{r eval=FALSE}

##  pre



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/pre_A_to_I.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_pre.rds")
saveRDS(frequency_all,"reads_counts_pre.rds")
saveRDS(error_reads_list,"error_reads_list_pre.rds")
saveRDS(methylation_rate,"methylation_rate_pre.rds")



##  spliced



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/spliced_A_to_I.rds")



testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_spliced.rds")
saveRDS(frequency_all,"reads_counts_spliced.rds")
saveRDS(error_reads_list,"error_reads_list_spliced.rds")
saveRDS(methylation_rate,"methylation_rate_spliced.rds")


##  isoform



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/isoform_A_to_I.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/origin/SNP_all.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,3]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,4]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]+frequency_all[[i]][strand_plus,4]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,1]+frequency_all[[i]][strand_minus,3]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process")
saveRDS(error_rate_list,"error_rate_list_isoform.rds")
saveRDS(frequency_all,"reads_counts_isoform.rds")
saveRDS(error_reads_list,"error_reads_list_isoform.rds")
saveRDS(methylation_rate,"methylation_rate_isoform.rds")
```
### calculate the error rate for site
```{r eval=FALSE}
# calculate teh meta_error_rate for each site
## spliced

rm(list = ls())
gc()
error_reads_list_spliced <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_reads_list_spliced.rds")

error_reads_list_spliced_sum <- Reduce("+", error_reads_list_spliced)

reads_counts_spliced <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_spliced.rds")

summed_vectors_list <- lapply(reads_counts_spliced, function(df) rowSums(df))

reads_counts_spliced_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_spliced <- error_reads_list_spliced_sum/
  (error_reads_list_spliced_sum+reads_counts_spliced_sum) 


saveRDS(error_rate_meta_spliced,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_rate_meta_spliced.rds")
## pre


rm(list = ls())
gc()
error_reads_list_pre <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_reads_list_pre.rds")

error_reads_list_pre_sum <- Reduce("+", error_reads_list_pre)

reads_counts_pre <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_pre.rds")

summed_vectors_list <- lapply(reads_counts_pre, function(df) rowSums(df))

reads_counts_pre_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_pre <- error_reads_list_pre_sum/
  (error_reads_list_pre_sum+reads_counts_pre_sum) 

saveRDS(error_rate_meta_pre,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_rate_meta_pre.rds")

```

### filter the site 
From the initial analysis of 1,152 cells, we identified 48,556 potential editing sites. After applying stringent quality filters that excluded sites with error rates exceeding 1%, we retained 1,159 high-confidence A-to-I editing sites. To ensure data reliability, we further filtered the dataset by removing cells with fewer than 200 reads in either spliced or unspliced mRNA fractions, resulting in a final set of 1,119 cells for downstream analysis.

```{r eval=FALSE}
# Load necessary library
library(VeloRM)
library(ggplot2)

# Read metadata
error_rate_meta_spliced <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_rate_meta_spliced.rds")
error_rate_meta_pre <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/error_rate_meta_pre.rds")
index_error <- intersect(which(error_rate_meta_spliced <= 0.01), which(error_rate_meta_pre <= 0.01))



# Read counts
reads_counts_spliced<- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_spliced.rds")

# Process spliced data
spliced_meth_test <- as.data.frame(lapply(reads_counts_spliced, function(df) df[["methylated_reads"]]))
spliced_unmeth_test <- as.data.frame(lapply(reads_counts_spliced, function(df) df[["unmethylated_reads"]]))


reads_counts_pre<- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_pre.rds")


# Process spliced data
pre_meth_test <- as.data.frame(lapply(reads_counts_pre, function(df) df[["methylated_reads"]]))
pre_unmeth_test <- as.data.frame(lapply(reads_counts_pre, function(df) df[["unmethylated_reads"]]))


# Set row names
rownames(spliced_meth_test) <- rownames(spliced_unmeth_test) <- paste0("site_", 1:dim(spliced_unmeth_test)[1])

rownames(pre_meth_test) <- rownames(pre_unmeth_test)  <- paste0("site_", 1:dim(pre_unmeth_test)[1])

# Calculate spliced and pre-mRNA test matrices
spliced_test <- spliced_meth_test+spliced_unmeth_test


pre_test <- pre_meth_test + pre_unmeth_test

#discard the cell have lower reads
index_cell_discard_spliced <- which(colSums(spliced_test)<=200)
index_cell_discard_pre <- which(colSums(pre_test)<=200)

spliced_test_discarded <- spliced_test[,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_test_discarded <- pre_test[,-union(index_cell_discard_spliced,index_cell_discard_pre)]




index <- index_error

spliced_meth_test_processed <- spliced_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
spliced_unmeth_test_processed <- spliced_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_meth_test_processed <- pre_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
pre_unmeth_test_processed <- pre_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]


saveRDS(union(index_cell_discard_spliced,index_cell_discard_pre),"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/index_cell_discard.rds")
saveRDS(index_error,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/index_error.rds")
saveRDS(index,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/index.rds")


reads_counts_isoform <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_isoform.rds")
reads_counts_isoform <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_isoform.rds")

# Process pre-mRNA data
isoform_meth_test <- as.data.frame(lapply(reads_counts_isoform, function(df) df[["methylated_reads"]]))
isoform_unmeth_test <- as.data.frame(lapply(reads_counts_isoform, function(df) df[["unmethylated_reads"]]))



rownames(isoform_meth_test) <- rownames(isoform_unmeth_test) <-  paste0("site_", 1:dim(isoform_unmeth_test)[1])

isoform_meth_test_processed <- isoform_meth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]
isoform_unmeth_test_processed <- isoform_unmeth_test[index,-union(index_cell_discard_spliced,index_cell_discard_pre)]

test.list <- list(spliced_meth_test_processed,spliced_unmeth_test_processed,pre_meth_test_processed,pre_unmeth_test_processed,
                  isoform_meth_test_processed,isoform_unmeth_test_processed)
control.list <- NULL

res_preprocess_0.5 <- methylation.sites.preprocess(test.list,control.list,0.5)
res_preprocess <- methylation.sites.preprocess(test.list,control.list)

saveRDS(test.list,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/test.list.rds")
saveRDS(control.list,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/control.list.rds")
saveRDS(res_preprocess_0.5,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_preprocess_0.5.rds")
saveRDS(res_preprocess,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
# 
SNP <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/origin/SNP_all.rds")
SNP_index <- SNP[index]
saveRDS(SNP_index,"D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/SNP_1159.rds")
```
### DESeq2 to test the differential methylation between the spliced mRNA and pre-mRNA
```{r eval=FALSE}
# Load input data
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data

# Set random seed for reproducibility
set.seed(42)

# Record start time
start_time <- Sys.time()

# Load DESeq2 package
library(DESeq2)

# Define DESeq2 differential expression analysis function
DESeq2_test <- function(meth_control, meth_test, unmeth_control, unmeth_test) {
  # Convert input data to data frames
  meth_control <- data.frame(meth_control)
  meth_test <- data.frame(meth_test)
  unmeth_control <- data.frame(unmeth_control)
  unmeth_test <- data.frame(unmeth_test)
  
  # Function to calculate size factor
  sizeFactor <- function(data) {
    # Ensure data is not less than 0
    data <- as.matrix(data)
    
    # Log-transform the data
    log_data <- log(data)
    log_data[is.infinite(log_data)] <- NA
    log_mean <- rowMeans(log_data, na.rm = TRUE)
    log_s <- log_data - log_mean
    
    # Calculate size factor
    s_size <- exp(apply(log_s, 2, function(x) median(x, na.rm = TRUE)))
    return(s_size)
  }
  
  # Calculate size factors for spliced and unspliced data
  spliced_total <- meth_control + unmeth_control
  unspliced_total <- meth_test + unmeth_test
  
  spliced_size_factor <- sizeFactor(spliced_total)
  unspliced_size_factor <- sizeFactor(unspliced_total)
  
  # Get the number of control and test samples
  num_control <- dim(unmeth_control)[2]
  num_test <- dim(unmeth_test)[2]
  
  # Combine all data
  counts <- cbind(meth_control, meth_test, unmeth_control, unmeth_test)
  
  # Design the experiment matrix
  design = data.frame(
    Trt = rep(c(rep("control", num_control), rep("test", num_test)), 2), 
    Meth = c(rep("meth", (num_control + num_test)), rep("unmeth", (num_control + num_test)))
  )
  
  # Define the model
  model = ~ Meth + Trt + Meth:Trt
  
  # Create DESeqDataSet object
  dds <- DESeqDataSetFromMatrix(countData = counts,
                                colData = design,
                                design = model)
  
  # Set size factors
  sizeFactors(dds) = rep(c(spliced_size_factor, unspliced_size_factor), 2)
  
  # Perform differential expression analysis
  dds <- DESeq(dds, test = "Wald")
  
  # Extract results
  res = DESeq2::results(dds, name = "Methunmeth.Trttest", cooksCutoff = FALSE)
  
  return(res)
}

# Call the function to perform differential expression analysis
result <- DESeq2_test(spliced.meth.test, unspliced.meth.test, spliced.unmeth.test, unspliced.unmeth.test)

# Record end time
end_time <- Sys.time()

# Print the runtime
print(end_time - start_time)


# Save the results
saveRDS(result, "/gpfs/work/bio/haozhewang17/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_DESeq2.rds")
```

We will get the 
Processed data:
SNP_1159: the SNP information for the A-to-I site for further analysis.
test_list: the test cell reads counts (spliced_meth_test,spliced_unmeth_test,pre_meth_test,pre_unmeth_test,
                   isoform_meth_test,isoform_unmeth_test)
control_list: NULL
res_preprocess: the list processed by VeloRM
res_DESeq2: the DESeq2 results

## analysis
### splicing jucntion distance
```{r eval=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(dplyr)
library(tidyr)

snp <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/SNP_1019.rds")
snp_pos <- snp@ranges@start
snp_chr_length <- snp@seqnames@lengths
snp_chr_seq <- snp@seqnames@values
snp_chr <- as.character(rep(snp_chr_seq,snp_chr_length)) 
snp_strand <- snp@elementMetadata@listData[["transcript_strand"]]
snp_data <- as.data.frame(cbind(snp_chr,snp_pos,snp_strand))
snp_data$snp_pos <- as.numeric(snp_data$snp_pos)




# Read the GTF file from Ensembl
gtf_file <- "D:/data/VeloRM_data/RNA-Seq_A-to-I/origin/gencode.vM33.annotation.gtf"
gtf_data <- import(gtf_file)

# Filter for exons
exons <- gtf_data[gtf_data$type == "exon"]

# Convert to data frame
exons_df <- data.frame(
chr = as.character(seqnames(exons)),
start = start(exons),
end = end(exons),
strand = as.character(strand(exons)),
gene_id = mcols(exons)$gene_id,
transcript_id = mcols(exons)$transcript_id
)

# Sort exons by transcript and position
exons_df <- exons_df %>%
arrange(transcript_id, chr, start)

# Extract splicing junctions using intron boundaries
junctions <- exons_df %>%
group_by(transcript_id) %>%
arrange(if_else(strand == "+", start, -start)) %>%  # Sort based on strand
mutate(
  donor_site = end,                # End of the upstream exon
  acceptor_site = lead(start)      # Start of the downstream exon
) %>%
filter(!is.na(acceptor_site)) %>%   # Remove the last exon (no downstream exon)
ungroup() %>%
select(chr, donor_site, acceptor_site, gene_id, transcript_id, strand)


junctions_unique <- junctions %>%
 distinct(chr, donor_site, acceptor_site, .keep_all = TRUE)



junctions_long <- junctions_unique %>%
 pivot_longer(cols = c(donor_site, acceptor_site), 
              names_to = "site_type", 
              values_to = "site_position") %>%
 select(chr, site_position, site_type, gene_id, transcript_id, strand)



# Calculate distance from each SNP to the nearest junction (matching chr and strand)
snp_with_distance <- snp_data %>%
 rowwise() %>%
 mutate(
   nearest_junction_distance = min(
     abs(junctions_long$site_position[junctions_long$chr == snp_chr & junctions_long$strand == snp_strand] - snp_pos),
     na.rm = TRUE
   )
 )


# get normal chr
standard_chromosomes <- paste0("chr", c(1:22, "X", "Y"))

# get the exon with the normal chr
exons_filtered <- exons[seqnames(exons) %in% standard_chromosomes]



# calculate the distance 
exon_lengths <- width(exons_filtered)

# calculate the prob 
prob_weights <- exon_lengths / sum(exon_lengths)

# random sample
set.seed(123) 


#random_distances_list <- list()

#for (i in 1:10){
 selected_exons <- sample(seq_along(exons_filtered), size = 1019, replace = TRUE, prob = prob_weights)
 # choose the site in the select the exons
 random_positions <- start(exons[selected_exons]) + 
   floor(runif(1019, min = 0, max = exon_lengths[selected_exons]))
 
 # create the data frame
 random_sites <- data.frame(
   chr = as.character(seqnames(exons[selected_exons])),
   position = random_positions,
   strand = as.character(strand(exons[selected_exons]))
 )
 
 
 # Calculate distance from each SNP to the nearest junction (matching chr and strand)
 random_sites_with_distance <- random_sites %>%
   rowwise() %>%
   mutate(
     nearest_junction_distance = min(
       abs(junctions_long$site_position[junctions_long$chr == chr & junctions_long$strand == strand] - position),
       na.rm = TRUE
     )
   )
 random_distances_list  <- random_sites_with_distance$nearest_junction_distance
#}



random_distances_df <- data.frame(
 distance = random_distances_list,
 group = "Random"
)

# merge the data frame
real_distances_df <- data.frame(
 distance = snp_with_distance$nearest_junction_distance,
 group = "Real"
)


all_distances_df <- bind_rows(random_distances_df, real_distances_df)
group_order <- c("Real","Random")
all_distances_df$group <- factor(all_distances_df$group, levels = group_order)
library(ggplot2)
#  Boxplot
plot <- ggplot(all_distances_df, aes(x = group, y = log10(distance+1), fill = group)) +
 geom_violin(trim = FALSE, alpha = 0.5) +
 geom_boxplot(width = 0.1, outlier.shape = NA)+
 scale_fill_manual(values = c("red",rep("lightblue", 10))) +
 labs(title = NULL,
      x = NULL,
      y = "log10(Distance (bp)+1)") +
 theme_minimal() +
 theme(
       plot.title = element_text(hjust = 0.5),
       legend.position = "none")+coord_cartesian(ylim = c(0,7))


saveRDS(all_distances_df,"D:/data/VeloRM_data/RNA-Seq_A-to-I/results/splicing_junction_analysis/all_distances_df.rds")


```
#### vis
```{r warning=FALSE}
 all_distances_df <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/results/splicing_junction_analysis/all_distances_df.rds")

library(ggplot2)
#  Boxplot
plot <- ggplot(all_distances_df, aes(x = group, y = log10(distance+1), fill = group)) +
 geom_violin(trim = FALSE, alpha = 0.5) +
 geom_boxplot(width = 0.1, outlier.shape = NA)+
 scale_fill_manual(values = c("red",rep("lightblue", 10))) +
 labs(title = NULL,
      x = NULL,
      y = "log10(Distance (bp)+1)") +
 theme_grey() +
 theme(plot.title = element_text(hjust = 0.5),
       legend.position = "none")+coord_cartesian(ylim = c(0,7))
plot
```


### read the DESeq2 results
```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

res_DESeq2 <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_DESeq2.rds")


log2FC <-res_DESeq2@listData[["log2FoldChange"]]
pvals <- res_DESeq2@listData[["pvalue"]]

plot_data <- data.frame(log2FC = log2FC, 
                       pvalue = pvals,
                       neg_log10_p = -log10(pvals))


volcano_plot_DESeq2 <- ggplot(plot_data, aes(x = log2FC, y = neg_log10_p)) +
  # Add light gray background for -1 < log2FC < 1 (non-significant FC region)
  annotate("rect", xmin = -1, xmax = 1, ymin = 0, ymax = 30, 
           fill = "gray90", alpha = 0.3) +
  
  # Color points based on significance and log2FC:
  # - Red: log2FC > 1 & p < 0.05 (significantly up-regulated)
  # - Blue: log2FC < -1 & p < 0.05 (significantly down-regulated)
  # - Gray: Not significant or |log2FC| ≤ 1
  geom_point(aes(color = case_when(
    log2FC > 1 & pvalue < 0.05 ~ "Unspliced",
    log2FC < -1 & pvalue < 0.05 ~ "Spliced",
    TRUE ~ "Not significant"
  )), alpha = 0.3) +
  
  # Manual color scale (red, blue, gray)
  scale_color_manual(
    values = c(
      "Spliced" = "red",
      "Unspliced" = "blue",
      "Not significant" = "gray50"
    ),
    name = "Regulation"  # Legend title
  ) +
  
  # Add dashed lines for thresholds:
  # - Vertical: log2FC = ±1
  # - Horizontal: p-value = 0.05 (-log10(0.05) ≈ 1.3)
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  
  # Axis labels and theme
  labs(
    x = "log2 Fold Change", 
    y = "-log10(p-value)",
    title = NULL
  ) +
   theme_grey() +
   # Set axis limits (zoomed-in view)
   coord_cartesian(ylim = c(0, 30),xlim = c(-4, 4))
volcano_plot_DESeq2

n_unspliced <- length(which(log2FC >= 1 & pvals <= 0.05))  
n_spliced <- length(which(log2FC <= -1 & pvals <= 0.05))  
n_nonsig <- sum(pvals > 0.05 | (log2FC > -1 & log2FC < 1))

total <- n_unspliced + n_spliced + n_nonsig
Percentage <- round(c(n_unspliced, n_spliced, n_nonsig) / total * 100, 1)

df <- data.frame(
  Category = c("Unspliced", "Spliced", "Insignificant"),
  Count = c(n_unspliced, n_spliced, n_nonsig)
)



pie_chart_DESeq2 <- ggplot(df, aes(x = "", y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 1, color = "white",alpha=0.6) +
  coord_polar("y", start = 0) + 
  geom_text(aes(label = paste0(Count, "\n(", Percentage, "%)")), 
            position = position_stack(vjust = 0.5), size = 5) +
  scale_fill_manual(values = c("Unspliced" = "blue", "Spliced" = "red","Insignificant"="gray50")) +
  labs(
    title = NULL,
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    legend.position = "none"  
  )

pie_chart_DESeq2
```

### meth density
```{r warning=FALSE}

# Load input data
test.list <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data
isoform.meth.test <- test.list[[5]]  # Methylated isoform data
isoform.unmeth.test <- test.list[[6]]  # Unmethylated isoform data





spliced.meth.sum.cell <- apply(spliced.meth.test,2,sum)#,apply(spliced.meth.control,2,sum))
spliced.unmeth.sum.cell <- apply(spliced.unmeth.test,2,sum)#,apply(spliced.unmeth.control,2,sum))
unspliced.meth.sum.cell <- apply(unspliced.meth.test,2,sum)#,apply(unspliced.meth.control,2,sum))
unspliced.unmeth.sum.cell <- apply(unspliced.unmeth.test,2,sum)#,apply(unspliced.unmeth.control,2,sum))
isoform.meth.sum.cell <- apply(isoform.meth.test,2,sum)#,apply(unspliced.meth.control,2,sum))
isoform.unmeth.sum.cell <- apply(isoform.unmeth.test,2,sum)#,apply(unspliced.unmeth.control,2,sum))

spliced.sum.cell <- spliced.meth.sum.cell + spliced.unmeth.sum.cell
unspliced.sum.cell <- unspliced.meth.sum.cell + unspliced.unmeth.sum.cell
isoform.sum.cell <- isoform.meth.sum.cell + isoform.unmeth.sum.cell

spliced.prob.cell <- spliced.sum.cell/(spliced.sum.cell+unspliced.sum.cell+isoform.sum.cell)
unspliced.prob.cell <- unspliced.sum.cell/(spliced.sum.cell+unspliced.sum.cell+isoform.sum.cell)
isoform.prob.cell <- isoform.sum.cell/(spliced.sum.cell+unspliced.sum.cell+isoform.sum.cell)


spliced.cell.prob <- spliced.meth.sum.cell/(spliced.unmeth.sum.cell+spliced.meth.sum.cell)
unspliced.cell.prob <- unspliced.meth.sum.cell/(unspliced.unmeth.sum.cell+unspliced.meth.sum.cell)

df <- data.frame(
  Probability = c(spliced.cell.prob, unspliced.cell.prob),
  Type = rep(c("mature RNA", "pre RNA"), 
             c(length(spliced.cell.prob), length(unspliced.cell.prob)))
)

# Create the density plot
ggplot(df, aes(x = Probability, fill = Type)) +
  geom_density(alpha = 0.3) +
  labs(title = "Density Plot of Methylation Probabilities",
       x = "Methylation Probability",
       y = "Density") +
  theme_grey()+
  theme(
    plot.title = element_text(hjust = 0.5, size = 14),
    legend.position = c(0.8, 0.7),
    legend.background = element_blank())+coord_cartesian(xlim=c(0.2,0.8))
```

### Regulatory mechanisms
#### Finding sites where methylation levels are high in pre-mRNA but low in spliced mRNA, as well as sites where methylation levels are high in spliced mRNA but low in pre-mRNA.
```{r warning=FALSE}
library(TxDb.Mmusculus.UCSC.mm39.knownGene)
library(MetaTX)
library(ggplot2)
library(dplyr)
res_preprocess <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
spliced_methylation_M_value_meta <- res_preprocess$spliced_methylation_M_value_meta
unspliced_methylation_M_value_meta <- res_preprocess$unspliced_methylation_M_value_meta



index_pre_high_spliced_low <- which(unspliced_methylation_M_value_meta-spliced_methylation_M_value_meta>=1)
index_spliced_high_pre_low <- which(spliced_methylation_M_value_meta-unspliced_methylation_M_value_meta>=1)
```



#### metaTX plot 
```{r warning=FALSE}
intersect_pre_dominate <-index_pre_high_spliced_low

intersect_spliced_dominate <-index_spliced_high_pre_low

reads_counts_isoform <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_isoform.rds")
reads_counts_pre <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_pre.rds")
reads_counts_spliced <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/reads_counts_spliced.rds")
reads_counts_meth <- as.data.frame(lapply(reads_counts_isoform,function(x)x[["methylated_reads"]]))+
  as.data.frame(lapply(reads_counts_pre,function(x)x[["methylated_reads"]]))+
  as.data.frame(lapply(reads_counts_spliced,function(x)x[["methylated_reads"]]))

reads_counts_unmeth <- as.data.frame(lapply(reads_counts_isoform,function(x)x[["unmethylated_reads"]]))+
  as.data.frame(lapply(reads_counts_pre,function(x)x[["unmethylated_reads"]]))+
  as.data.frame(lapply(reads_counts_spliced,function(x)x[["unmethylated_reads"]]))

reads_counts_M_overall <- ((rowSums(reads_counts_meth)+1e-7)/(rowSums(reads_counts_unmeth)+rowSums(reads_counts_meth)+1e-7))


index_not_dominate <- order(reads_counts_M_overall,decreasing = TRUE)[1:500]

# metaTX

SNP_all <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/origin/SNP_all.rds")
SNP_1019 <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/SNP_1019.rds")
txdb <- TxDb.Mmusculus.UCSC.mm39.knownGene


SNP_intersect_pre_dominate <- SNP_1019[intersect_pre_dominate]
SNP_intersect_spliced_dominate <- SNP_1019[intersect_spliced_dominate]
SNP_intersect_not_dominate <- SNP_all[index_not_dominate]
SNP_background <- SNP_1019

SNP_to_exclude <- c(SNP_intersect_pre_dominate, SNP_intersect_spliced_dominate)

library(GenomicRanges)
hits <- findOverlaps(SNP_intersect_not_dominate, SNP_to_exclude)

if(length(hits)!=0){
  SNP_intersect_not_dominate<- SNP_intersect_not_dominate[-queryHits(hits)]
}

remap_results_pre_dominate <- remapCoord(features = SNP_1019[intersect_pre_dominate], txdb = txdb)
remap_results_spliced_dominate <- remapCoord(features = SNP_1019[intersect_spliced_dominate], txdb = txdb)
remap_results_not_dominate <- remapCoord(features = SNP_intersect_not_dominate, txdb = txdb)
remap_results_background <- remapCoord(features = SNP_background, txdb = txdb)


# Step 2: Create plots for each subset
plot_pre_dominate <- directPlot(remap_results_pre_dominate)
plot_spliced_dominate <- directPlot(remap_results_spliced_dominate)
plot_not_dominate <- directPlot(remap_results_not_dominate)
plot_background <- directPlot(remap_results_background)

data_a <- ggplot_build(plot_pre_dominate)$data[[1]] %>% mutate(Group = "Precursor")
data_b <- ggplot_build(plot_spliced_dominate)$data[[1]] %>% mutate(Group = "Mature")
data_c <- ggplot_build(plot_not_dominate)$data[[1]] %>% mutate(Group = "Other")
data_d <- ggplot_build(plot_background)$data[[1]] %>% mutate(Group = "Real")

# Combine all datasets
combined_data <- bind_rows(data_a, data_b, data_c, data_d)#, data_intersect)

region_rects <- data.frame(
    xmin = c(0.2, 0.4, 0.6),  # Start positions of each region
    xmax = c(0.4, 0.6, 0.8),  # End positions of each region
    ymin = min(combined_data$y, na.rm = TRUE) - 0.1,  # Adjust below data range
    ymax = min(combined_data$y, na.rm = TRUE) - 0.05, # Slightly higher to make room
    Region = c("5' UTR", "CDS", "3' UTR")
)

# Calculate the height of 5' UTR and 3' UTR regions
utr_height <- region_rects$ymax[region_rects$Region == "5' UTR"] - region_rects$ymin[region_rects$Region == "5' UTR"]

# Increase the height of CDS region by 20% (can be adjusted as needed)
cds_height <- utr_height * 2 # Increase height by 30%

# Set ymin and ymax for CDS so that it is centered with respect to 5' UTR and 3' UTR
region_rects$ymin[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] - 0.5*utr_height  # Decrease ymin for CDS
region_rects$ymax[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] + cds_height  # Increase ymax for CDS


compare_metaTX_plot <-ggplot(combined_data, aes(x = x, y = y, color = Group))+
  geom_line(size=1)+
    # Add region highlights using geom_rect, with consistent positioning across all facets
    geom_rect(data = region_rects, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = Region),
              inherit.aes = FALSE, color = "black", alpha = 0.4, 
              linewidth = 0.5) +  # Default linewidth for all regions
    # Set fill for CDS region to grey (lighter than black)
    geom_rect(data = subset(region_rects, Region == "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE, fill = "grey", color = "black", alpha = 0.4, linewidth = 0.5) +  # Grey fill for CDS
    # Adjust 5' UTR and 3' UTR fills to a very dark grey, closer to black
    geom_rect(data = subset(region_rects, Region != "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                                                                fill = factor(Region, levels = c("5' UTR", "3' UTR"))), 
              inherit.aes = FALSE, color = "black", alpha = 0.3, linewidth = 0.5) + 
    scale_fill_manual(values = c("5' UTR" = grey(0.15),  # Very dark grey for 5' UTR, closer to black
                                "CDS" = "grey",           # Grey for CDS
                                "3' UTR" = grey(0.15))) +  # Very dark grey for 3' UTR
    # Adjust y position for the text annotations to place them above the rectangles
    geom_text(data = region_rects, aes(x = (xmin + xmax) / 2, y = ymax - 0.5, label = Region),
              inherit.aes = FALSE, color = "black", size = 3, hjust = 0.5) +
    # Remove Region fill legend
    guides(fill = "none")+scale_color_manual(values = c(
  "Other" = "#00ba38",      
  "Mature" = "#619cff",     
  "Precursor" = "#F8766d" ,
  "Real" = "#4D4D4D"  
),name = "Higher on")+theme_grey()+
  theme(
    legend.position = c(0.2, 0.7) ,legend.background = element_blank() )+xlab(NULL)+ylab(NULL)

compare_metaTX_plot
```

#### motif plot
```{r warning=FALSE}
library(GenomicRanges)
library(Biostrings)
library(BSgenome)
library(BSgenome.Mmusculus.UCSC.mm39)
genome <- BSgenome.Mmusculus.UCSC.mm39

## pre 

# transcript_strand  +
SNP_intersect_pre_dominate_plus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="+")] 


extended_range_intersect_pre_dominate_plus <- resize(SNP_intersect_pre_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_pre_dominate_plus <- getSeq(genome, extended_range_intersect_pre_dominate_plus)



# transcript_strand  -
SNP_intersect_pre_dominate_minus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="-")] 

extended_range_intersect_pre_dominate_minus <- resize(SNP_intersect_pre_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_pre_dominate_minus <- getSeq(genome, extended_range_intersect_pre_dominate_minus)



## spliced
# transcript_strand  +
SNP_intersect_spliced_dominate_plus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="+")] 

extended_range_intersect_spliced_dominate_plus <- resize(SNP_intersect_spliced_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_spliced_dominate_plus <- getSeq(genome, extended_range_intersect_spliced_dominate_plus)



# transcript_strand  -
SNP_intersect_spliced_dominate_minus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="-")] 

extended_range_intersect_spliced_dominate_minus <- resize(SNP_intersect_spliced_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_spliced_dominate_minus <- getSeq(genome, extended_range_intersect_spliced_dominate_minus)








## not 

# transcript_strand  +
SNP_intersect_not_dominate_plus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="+")] 


extended_range_intersect_not_dominate_plus <- resize(SNP_intersect_not_dominate_plus, width = 5, fix = "center")

# getseq
sequence_intersect_not_dominate_plus <- getSeq(genome, extended_range_intersect_not_dominate_plus)



# transcript_strand  -
SNP_intersect_not_dominate_minus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="-")] 

extended_range_intersect_not_dominate_minus <- resize(SNP_intersect_not_dominate_minus, width = 5, fix = "center")

# getseq
sequence_intersect_not_dominate_minus <- getSeq(genome, extended_range_intersect_not_dominate_minus)





sequence_intersect_pre_dominate_minus_rc <- reverseComplement(sequence_intersect_pre_dominate_minus)
sequence_intersect_pre_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus_rc))
sequence_intersect_pre_dominate_plus_rc <- reverseComplement(sequence_intersect_pre_dominate_plus)
sequence_intersect_pre_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus_rc))


sequence_intersect_pre_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus))
sequence_intersect_pre_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus))



sequence_intersect_spliced_dominate_minus_rc <- reverseComplement(sequence_intersect_spliced_dominate_minus)
sequence_intersect_spliced_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus_rc))
sequence_intersect_spliced_dominate_plus_rc <- reverseComplement(sequence_intersect_spliced_dominate_plus)
sequence_intersect_spliced_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus_rc))


sequence_intersect_spliced_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus))
sequence_intersect_spliced_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus))

sequence_intersect_not_dominate_minus_rc <- reverseComplement(sequence_intersect_not_dominate_minus)
sequence_intersect_not_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus_rc))
sequence_intersect_not_dominate_plus_rc <- reverseComplement(sequence_intersect_not_dominate_plus)
sequence_intersect_not_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus_rc))



sequence_intersect_not_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus))
sequence_intersect_not_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus))


sequence_intersect_pre_dominate <- c(sequence_intersect_pre_dominate_plus,sequence_intersect_pre_dominate_minus_rc)

sequence_intersect_spliced_dominate <- c(sequence_intersect_spliced_dominate_plus,sequence_intersect_spliced_dominate_minus_rc)

sequence_intersect_not_dominate <- c(sequence_intersect_not_dominate_plus,sequence_intersect_not_dominate_minus_rc)



library(ggseqlogo)
seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate),
    "pre_dominate_plus" = as.character(sequence_intersect_pre_dominate_plus),
    "spliced_dominate_plus" = as.character(sequence_intersect_spliced_dominate_plus),
    "not_dominate_plus" = as.character(sequence_intersect_not_dominate_plus),
    "pre_dominate_minus" = as.character(sequence_intersect_pre_dominate_minus),
    "spliced_dominate_minus" = as.character(sequence_intersect_spliced_dominate_minus),
    "not_dominate_minus" = as.character(sequence_intersect_not_dominate_minus)
)
ggseqlogo(seq_list, method = 'bits') + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")

seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate))
ggseqlogo(seq_list, method = 'bits', seq_type = "RNA") + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")  +
  theme_grey() 
ggseqlogo(seq_list, method = 'prob', seq_type = "RNA") + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")  +
  theme_grey() 

library(universalmotif)

motif_list <- lapply(seq_list,function(x) create_motif(x, alphabet = "RNA"))
```




#### find the index of site only detect methylated reads in spliced or pre mRNA
```{r warning=FALSE}
index_meth_only_detect_pre <-  which(res_preprocess[["vg.test"]] %in% res_preprocess[["index"]][["methylation only occurs at unspliced RNA"]])

index_meth_only_detect_spliced <-  which(res_preprocess[["vg.test"]] %in%res_preprocess[["index"]][["methylation only occurs at spliced RNA"]])

# only detect at spliced
index_meth_only_detect_spliced_int <-  intersect(index_meth_only_detect_spliced,intersect_spliced_dominate)
SNP_1019[index_meth_only_detect_spliced_int]



library(biomaRt)

# Ensembl (Mouse)
ensembl <- useEnsembl(biomart = "genes", dataset = "mmusculus_gene_ensembl")

# ENST 
transcript_ids <- SNP_1019[index_meth_only_detect_spliced_int]$transcript_overlapped

# clean
transcript_ids_clean <- gsub("\\..*", "", transcript_ids)

# find
transcript_info <- getBM(
  attributes = c("ensembl_transcript_id", "ensembl_gene_id", "external_gene_name", 
                 "chromosome_name", "start_position", "end_position", "strand"),
  filters = "ensembl_transcript_id",
  values = transcript_ids_clean,
  mart = ensembl
)

idx <- match(transcript_ids_clean, transcript_info$ensembl_transcript_id)
transcript_info_ordered <- transcript_info[idx, ]
transcript_info_ordered <-transcript_info_ordered [,c(4,3,7)]
transcript_info_ordered$position <- SNP_1019[index_meth_only_detect_spliced_int]@ranges@start
transcript_info_ordered$position <- transcript_info_ordered$position-transcript_info_ordered$strand


transcript_info_ordered$meta_M <- 
spliced_methylation_M_value_meta[index_meth_only_detect_spliced_int]-unspliced_methylation_M_value_meta[index_meth_only_detect_spliced_int]

transcript_info_ordered_spliced <-transcript_info_ordered

# only detect at pre
index_meth_only_detect_pre_int <- intersect(index_meth_only_detect_pre,intersect_pre_dominate)
SNP_1019[index_meth_only_detect_pre_int]


# ENST 
transcript_ids <- SNP_1019[index_meth_only_detect_pre_int]$transcript_overlapped

# clean
transcript_ids_clean <- gsub("\\..*", "", transcript_ids)

# find
transcript_info <- getBM(
  attributes = c("ensembl_transcript_id", "ensembl_gene_id", "external_gene_name", 
                 "chromosome_name", "start_position", "end_position", "strand"),
  filters = "ensembl_transcript_id",
  values = transcript_ids_clean,
  mart = ensembl
)

idx <- match(transcript_ids_clean, transcript_info$ensembl_transcript_id)
transcript_info_ordered <- transcript_info[idx, ]
transcript_info_ordered <-transcript_info_ordered [,c(4,3,7)]
transcript_info_ordered$position <- SNP_1019[index_meth_only_detect_pre_int]@ranges@start
transcript_info_ordered$position <- transcript_info_ordered$position-transcript_info_ordered$strand


transcript_info_ordered$meta_M <- 
unspliced_methylation_M_value_meta[index_meth_only_detect_pre_int]-spliced_methylation_M_value_meta[index_meth_only_detect_pre_int]

transcript_info_ordered_pre <-transcript_info_ordered

```
## run velocity
```{r,warning=FALSE}
library(ggplot2)
res_preprocess <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/res_preprocess.rds")
# the input data
test.list <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/test.list.rds")
control.list <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/control.list.rds")


spliced.meth.test <- test.list[[1]]
spliced.unmeth.test <- test.list[[2]]
unspliced.meth.test <- test.list[[3]]
unspliced.unmeth.test <- test.list[[4]]

kCells = 1; deltaT =1; kSites = 1
library(VeloRM)
# calculate the emat_size and nmat_size

emat_size <- colSums(spliced.meth.test + spliced.unmeth.test)
nmat_size <- colSums(unspliced.meth.test + unspliced.unmeth.test)

res <- methylation.site.relative.velocity.estimates(res_preprocess$test, res_preprocess$control,
                                                    kCells = kCells, deltaT = deltaT, kSites = kSites, 
                                                    fit.quantile = 0.05, n.cores = 1, emat.size = emat_size, 
                                                    nmat.size = nmat_size)

saveRDS(res, "D:/data/VeloRM_data/RNA-Seq_A-to-I/results/res_A_to_I_velocity.rds")

```

### load the embedding (tsne)
```{r,warning=FALSE}
sra_table <- read.csv("D:/data/VeloRM_data/RNA-Seq_A-to-I/SraRunTable_E12.5.csv", 
                      header = TRUE, 
                      stringsAsFactors = FALSE)

# Extract the first column which contains SRR 

classification_table <- read.csv("D:/data/VeloRM_data/RNA-Seq_A-to-I/Classification.csv")
GSM_info_2 <- classification_table[,1]
classification_label <- classification_table[,2]

GSM_info_2 <- GSM_info_2[grep("E12.5",classification_label)]
classification_label <- classification_label [grep("E12.5",classification_label)]
classification_label <- gsub(" \\[single cell RNA-seq\\]","", classification_label)
classification_label <- gsub("E12.5_","", classification_label)

SRR_info <- sra_table[match(GSM_info_2,sra_table[,17]),1]

which(colnames(res[[3]][["current"]][["M"]])!=SRR_info)

embedding <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/embedding.rds")
cell.colors <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/cell.colors.rds")
embedding <- embedding[match(names(cell.colors), rownames(embedding)), ]


index <- which(classification_label%in%rownames(embedding))

index_order <- match(names(cell.colors),classification_label[index])

```
#### filter
```{r,warning=FALSE}

SNP_1019 <- readRDS("D:/data/VeloRM_data/RNA-Seq_A-to-I/pre_process/SNP_1019.rds")

names(SNP_1019) <- rownames(test.list[[1]])

spliced_methylation_Beta_value_meta_index <- res_preprocess[["spliced_methylation_Beta_value_meta"]]
unspliced_methylation_Beta_value_meta_index <-res_preprocess[["unspliced_methylation_Beta_value_meta"]]


Beta_not_1or0 <- apply(res[[3]][["current"]]$Beta,1,function(x) length(which(x!=0&x!=1)))
index_site_name_int_intersect <-   names(which(Beta_not_1or0>=10))

```

#### trajectory M
```{r,warning=FALSE}



umap_results <- embedding
rownames(umap_results) <- SRR_info[index][index_order]
names(cell.colors) <- SRR_info[index][index_order]
umap_results <- as.data.frame(umap_results)
umap_results$color <- as.factor(cell.colors)

ggplot() +
  geom_point(
    data = umap_results,
    aes(x = V1, y = V2, color = color)
  ) +
  labs(x = "PC1", y = "PC2") +  # Change axis labels
  theme(
    legend.position = "none",
    legend.background = element_blank()
  )


arrow.scale=3; cell.alpha=0.4; cell.cex=1; fig.height=4;

vis_umap_Beta_emb_Beta_linear_index  <- show.velocity.on.embedding.cor(umap_results[,1:2],
             res[[3]][["current"]]$M[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$M[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$M[index_site_name_int_intersect,index],
             cell.colors=cell.colors,n=200,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,rm_uniform="mode1")
vis_umap_Beta_emb_Beta_linear_index[[1]]

arrow.scale=10; cell.alpha=0.4; cell.cex=1; fig.height=4;
vis_umap_Beta_emb_Beta_linear_index_grid <- show.velocity.on.embedding.cor(umap_results[,1:2],
             res[[3]][["current"]]$M[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$M[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$M[index_site_name_int_intersect,index],
             cell.colors=cell.colors,n=200,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,show.grid.flow = TRUE,grid.n = 20,rm_uniform="mode1")


vis_umap_Beta_emb_Beta_linear_index_grid[[1]]

```

