---
title: "scDART-seq_m6A_dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scDART-seq_m6A_dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(VeloRM)
```
## scDART_seq velocity pipeline
### data download
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE180954 
### alignment
```{bash eval=FALSE, include=FALSE}
(nohup trim_galore --paired -o /home/share/bowen/scDART/data/YTHcell_',i,' /home/share/bowen/scDART/data/YTHcell_',i,'/',sra,'_1.fastq',' /home/share/bowen/scDART/data/YTHcell_',i,'/',sra,'_2.fastq)&
  
(nohup STAR --runThreadN 10 --runMode genomeGenerate --genomeDir /home/share/bowen/genome/hg38/star_indx/ --genomeFastaFiles /home/share/bowen/genome/hg38/Homo_sapiens.GRCh38.dna.primary_assembly.fa --sjdbGTFfile /home/share/bowen/genome/hg38/Homo_sapiens.GRCh38.109.gtf --sjdbOverhang 150 > /home/share/bowen/genome/hg38/star_indx/build_index.out)&   

(nohup STAR --runThreadN 10 --outSAMtype BAM Unsorted --readFilesIn /home/share/bowen/scDART/data/YTHcell_1/SRR15268889_1_val_1.fq /home/share/bowen/scDART/data/YTHcell_1/SRR15268889_2_val_2.fq --genomeDir /home/share/bowen/genome/hg38/star_indx --sjdbGTFfile /home/share/bowen/genome/hg38/Homo_sapiens.GRCh38.109.gtf --outFileNamePrefix /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889 --outSAMattributes All --outSAMunmapped Within --sjdbOverhang 150 > /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889.log)& 
  
(nohup STAR --runThreadN 10 --outSAMtype BAM Unsorted --readFilesIn /home/share/bowen/scDART/data/YTHmutcell_1/SRR15268462_1_val_1.fq /home/share/bowen/scDART/data/YTHmutcell_1/SRR15268462_2_val_2.fq --genomeDir /home/share/bowen/genome/hg38/star_indx --sjdbGTFfile /home/share/bowen/genome/hg38/Homo_sapiens.GRCh38.109.gtf --outFileNamePrefix /home/share/bowen/scDART/data/YTHmutcell_1/bam/SRR15268462 --outSAMattributes All --outSAMunmapped Within --sjdbOverhang 150 > /home/share/bowen/scDART/data/YTHmutcell_1/bam/SRR15268462.log)&  

```
### variant calling
```{bash eval=FALSE, include=FALSE}
samtools sort -o /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889_sorted.bam /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889Aligned.out.bam

bcftools mpileup -f /home/share/bowen/genome/hg38/Homo_sapiens.GRCh38.dna.primary_assembly.fa /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889_sorted.bam | bcftools call -mv -Oz -o /home/share/bowen/scDART/data/YTHcell_1/bam/SRR15268889_calls.vcf 
```
### find relevant mutation
```{r eval=FALSE, include=FALSE}
data <- read.vcfR(SRR15268889_calls.vcf)

data2 <- as.data.frame(data@fix)

indx <- which(data2$REF == 'A' | data2$REF == 'G' | data2$REF == 'C' | data2$REF == 'T')

data2 <- data2[indx,]

indx <- which(data2$ALT == 'A' | data2$ALT == 'G' | data2$ALT == 'C' | data2$ALT == 'T')

data2 <- data2[indx,]

gr <- GRanges(seqnames = paste0('chr',data2$CHROM),
              IRanges(start = as.numeric(data2$POS),
                      width = 1),
              strand = '+',
              ref = data2$REF,
              alt = data2$ALT,
              quality = as.numeric(data2$QUAL),
              info = data2$INFO)

gr <- keepStandardChromosomes(gr,pruning.mode = "coarse")

ref <- as.character(DNAStringSet(Views(Hsapiens,gr)))

indx <- which(gr$ref != ref)

if(length(indx) != 0){
  gr <- gr[-indx]
}
  
saveRDS(gr,'/home/share/bowen/scDART/data/YTHcell_1/bam/SNP.rds')  
  
# filter mutation
transcript <- transcripts(TxDb.Hsapiens.UCSC.hg38.knownGene)

transcript_plus <- transcript[which(strand(transcript)== "+")]

transcript_minus <- transcript[which(strand(transcript)== "-")]

gr <- readRDS('/home/share/bowen/scDART/data/YTHcell_1/bam/SNP.rds')

gr_all <- GRanges()

## 1.	Coordinate (+) strand: ACX
gr_plus <- gr
olp <- findOverlaps(gr_plus,transcript_plus)
q <- queryHits(olp)
s <- subjectHits(olp)
gr_plus <- gr_plus[q]
gr_plus$transcript_overlapped <- transcript_plus[s]$tx_name
gr_plus$transcript_strand <- '+'
gr_plus <- unique(gr_plus)

gr_plus$seq <- as.character(DNAStringSet(Views(Hsapiens,gr_plus+1)))
indx <- which(gr_plus$seq == 'ACA'|gr_plus$seq == 'ACG'|gr_plus$seq == 'ACC'|gr_plus$seq == 'ACT')
gr_plus <- gr_plus[indx]
indx <- which(gr_plus$alt == 'T')
gr_plus <- gr_plus[indx]
gr_plus$sample <- list3[i]

## 2.	Coordinate (-) strand: XGT
gr_minus <- gr
strand(gr_minus) <- '-'

olp <- findOverlaps(gr_minus,transcript_minus)
q <- queryHits(olp)
s <- subjectHits(olp)
gr_minus <- gr_minus[q]
gr_minus$transcript_overlapped <- transcript_minus[s]$tx_name
gr_minus$transcript_strand <- '-'
gr_minus <- unique(gr_minus)

strand(gr_minus) <- '+'
gr_minus$seq <- as.character(DNAStringSet(Views(Hsapiens,gr_minus+1)))
indx <- which(gr_minus$seq == 'AGT'|gr_minus$seq == 'CGT'|gr_minus$seq == 'GGT'|gr_minus$seq == 'TGT')
gr_minus <- gr_minus[indx]
indx <- which(gr_minus$alt == 'A')
gr_minus <- gr_minus[indx]
gr_minus$sample <- list3[i]

gr <- c(gr_plus,gr_minus)
saveRDS(gr,paste0(list2[i],'SNP_filtered.rds'))
gr_all <- unique(c(gr_all,gr))
saveRDS(gr_all,'/home/share/bowen/scDART/SNP_all.rds')
```

###  Quality control

```{bash eval=FALSE, include=FALSE}


setwd('/home/share/bowen/velocity/')
list <- read.csv('/home/share/bowen/scDART/sample.csv')
gsm2sra <- read.csv('/home/share/bowen/scDART/SraRunTable.txt')
folders <- readRDS("/home/share/bowen/velocity/folders.rds")

path <- '/home/share/bowen/scDART/data/'
folders <- list.dirs(path, full.names = FALSE, recursive = FALSE)



folders <- strsplit(folders,'_')
folders <- unlist(lapply(folders,function(x) x[2]))
folders <- as.numeric(folders)
saveRDS(folders,'folders.rds')

j <- 0
for(i in folders){
  j <- j+1
  dir.create(paste0('/home/share/bowen/velocity/data/YTHcell_',i))
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ',i)
  gsm_name <- list$Accession[which(list$Title == name)]
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/home/share/bowen/scDART/data/','YTHcell_',i,'/bam/',sra,'_sorted.bam')
  newfile <- paste0('/home/share/bowen/velocity/data/YTHcell_',i,'/',sra,'_above.mapQ30.bam')
  run <- paste0('samtools view -h -b -q 30 ',path,' -o ',newfile)
  system(run)
  print(j)
}

```
### divide bam
```{bash eval=FALSE, include=FALSE}

###########################################################
# File prepare
###########################################################

## get exon-no-ambiguity bed by substract (exon.bed - intron.bed = exon-na.bed)
bedtools subtract -a UCSC_Exons.bed -b UCSC_Introns.bed > exon-na.bed

## get intron-no-ambituity bed by subtract (inton.bed - exon.bed = inron-na.bed)
bedtools subtract -a UCSC_Introns.bed -b UCSC_Exons.bed > inron-na.bed

###########################################################
### divide bam into spliced bam and unspliced bam
###########################################################
setwd('/home/share/bowen/velocity/')
list <- read.csv('/home/share/bowen/scDART/sample.csv')
gsm2sra <- read.csv('/home/share/bowen/scDART/SraRunTable.txt')
folders <- readRDS("/home/share/bowen/velocity/folders.rds")

j <- 0
for(i in folders){
  j <- j+1
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ',i)
  gsm_name <- list$Accession[which(list$Title == name)]
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/home/share/bowen/velocity/data/YTHcell_',i,'/')
  newfile <- paste0('/home/share/bowen/velocity/data/YTHcell_',i,'/',sra,'_above.mapQ30.bam')
  
  #### mRNA.bam
  run <- paste0('bedtools intersect -a ',newfile,' -b /home/share/bowen/velocity/UCSC_Introns.bed -v -wa -split > ','/home/share/bowen/velocity/data/YTHcell_',i,'/',sra,'_mRNA_temp.bam')
  system(run)
  rm(run)
  
  run2 <- paste0('bedtools intersect -a ','/home/share/bowen/velocity/data/YTHcell_',i,'/',sra,'_mRNA_temp.bam -b /home/share/bowen/velocity/UCSC_Exons.bed -wa -split > ','/home/share/bowen/velocity/data/YTHcell_',i,'/',sra,'_mRNA.bam')
  system(run2)
  rm(run2)
  
  #### mRNA_spliced.bam
  run3 <- paste0("samtools view -h ",path,sra,"_mRNA.bam | awk '$0 ~ /^@/ || $6 ~ /N/' | samtools view -b > ",path,sra,"_mRNA_spliced.bam")
  system(run3)
  rm(run3)
  
  #### pre-mRNA.bam
  run4 <- paste0('bedtools intersect -a ',newfile,' -b /home/share/bowen/velocity/intron-na.bed -wa -split > ',path,sra,'_pre-mRNA.bam')
  system(run4)
  rm(run4)
  
  #### isoform-a.bam
  run5 <- paste0('bedtools intersect -a ',newfile,' -b /home/share/bowen/velocity/UCSC_Introns.bed -wa -split > ',path,sra,'_maybe_pre_mRNA.bam')
  system(run5)
  rm(run5)
  
  run6 <- paste0('bedtools intersect -a ',path,sra,'_maybe_pre_mRNA.bam -b /home/share/bowen/velocity/intron-na.bed -v -wa -split > ',path,sra,'_isoform-a.bam')
  system(run6)
  rm(run6)
  
  print(j)
}

```
### base counting
```{r eval=FALSE, include=FALSE}
###########################################################
### base counting for spliced and unspliced isoform bam
###########################################################
setwd('/home/share/bowen/velocity/')
list <- read.csv('/home/share/bowen/scDART/sample.csv')
gsm2sra <- read.csv('/home/share/bowen/scDART/SraRunTable.txt')
folders <- readRDS("/home/share/bowen/velocity/folders.rds")


## annotate coverage 
library(GenomicRanges)
library(Biostrings)
library(GenomicAlignments)
library(Rsamtools)
gr_all <- readRDS('/home/share/bowen/scDART/SNP_all.rds')[,-c(1:5)]
seqlevels(gr_all) <- gsub('chr','',seqlevels(gr_all))

## base counting
j <- 0
for (i in folders) {
  j <- j+1
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ',i)
  gsm_name <- list$Accession[which(list$Title == name)]
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/home/share/bowen/velocity/data/YTHcell_',i,'/')
  
  ################################## mRNA_spliced.bam
  bamfilePath <- paste0(path,sra,'_mRNA_spliced.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  ################################## pre-mRNA.bam
  bamfilePath <- paste0(path,sra,'_pre-mRNA.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  
  ################################## isoform-mRNA.bam
  bamfilePath <- paste0(path,sra,'_isoform-a.bam.bam')
  flag <- scanBamFlag(isNotPassingQualityControls=FALSE,
                      isDuplicate=FALSE)
  param <- ScanBamParam(flag=flag, what=c("seq","qual"),which=gr_all)
  alns <- readGAlignments(bamfilePath, param=param)
  alns <- keepStandardChromosomes(alns,pruning.mode = "coarse")
  qseq <- mcols(alns)$seq
  qual <- mcols(alns)$qual
  seqlevels(alns) <- seqlevels(gr_all)
  
  nucl_piles <- pileLettersAt(qseq, seqnames(alns), start(alns),
                              cigar(alns), gr_all)
  frequency <- alphabetFrequency(nucl_piles, baseOnly=TRUE)
  
  saveRDS(nucl_piles,paste0(bamfilePath,'_nucl_piles_all.rds'))
  saveRDS(frequency,paste0(bamfilePath,'_frequency_all.rds'))
  rm(param,alns,qseq,qual,nucl_piles,frequency,flag)
  print(j)
}
###########################################################
#### combine the frequency as the matrix (pre mRNA spliced mRNA of the test cell)
###########################################################
# Rename 'list' to 'data_list' or another appropriate name
data_list <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/sample.csv')
gsm2sra <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/SraRunTable.txt')

# Ensure 'folders' and 'gsm2sra' are properly defined before running the loop

missing_index <- c(91,97:99,103,158,228,308,347,357,384,386,262)

# Read the frequency and remove missing indices
folders <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity/data_73/yth_folders.rds")
folders <- folders[-missing_index]

# Initialize the spliced_list
spliced_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHcell_', i, '/')
  spliced_list[[j]] <- readRDS(paste0(path, sra, "_mRNA_spliced.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of spliced_list
names(spliced_list) <- paste0("YTHcell_", folders)
saveRDS(spliced_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/frequency_spliced_YTH.rds")

# Initialize the pre_list
pre_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHcell_', i, '/')
  pre_list[[j]] <- readRDS(paste0(path, sra, "_pre-mRNA.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of pre_list
names(pre_list) <- paste0("YTHcell_", folders)
saveRDS(pre_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/frequency_pre_YTH.rds")

# Rename 'list' to 'data_list' or another appropriate name
data_list <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/sample.csv')
gsm2sra <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/SraRunTable.txt')

# Ensure 'folders' and 'gsm2sra' are properly defined before running the loop

missing_index <- c(91,97:99,103,158,228,308,347,357,384,386,262)

# Read the frequency and remove missing indices
folders <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity/data_73/yth_folders.rds")
folders <- folders[-missing_index]

# Initialize the isoform_list
isoform_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTH Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHcell_', i, '/')
  isoform_list[[j]] <- readRDS(paste0(path, sra, "_isoform-a.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of isoform_list
names(isoform_list) <- paste0("YTHcell_", folders)
saveRDS(isoform_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/isoform_73_YTH.rds")




###########################################################
#### combine the frequency as the matrix (pre mRNA spliced mRNA of the control cell)
###########################################################

# Rename 'list' to 'data_list' or another appropriate name
data_list <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/sample.csv')
gsm2sra <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/SraRunTable.txt')

# Ensure 'folders' and 'gsm2sra' are properly defined before running the loop




folders <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity/data_73/ythmut_folders.rds")


# Initialize the spliced_list
spliced_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTHmut Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHmutcell_', i, '/')
  spliced_list[[j]] <- readRDS(paste0(path, sra, "_mRNA_spliced.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of spliced_list
names(spliced_list) <- paste0("YTHmutcell_", folders)
saveRDS(spliced_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/frequency_spliced_YTHmut.rds")

# Initialize the pre_list
pre_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTHmut Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHmutcell_', i, '/')
  pre_list[[j]] <- readRDS(paste0(path, sra, "_pre-mRNA.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of pre_list
names(pre_list) <- paste0("YTHmutcell_", folders)
saveRDS(pre_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/frequency_pre_YTHmut.rds")



# Rename 'list' to 'data_list' or another appropriate name
data_list <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/sample.csv')
gsm2sra <- read.csv('/gpfs/work/bio/haozhewang17/data/velocity/data_73/SraRunTable.txt')

# Ensure 'folders' and 'gsm2sra' are properly defined before running the loop


folders <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity/data_73/ythmut_folders.rds")


# Initialize the spliced_list
isoform_list <- list()
j = 1
for (i in folders) {
  
  name <- paste0('iDART_SMART-seq APOBEC1-YTHmut Cell ', i)
  gsm_name <- data_list$Accession[which(data_list$Title == name)]  # Use data_list here
  sra <- gsm2sra$Run[which(gsm2sra$GEO_Accession..exp. == gsm_name)]
  path <- paste0('/gpfs/work/bio/haozhewang17/data/velocity/data_73/YTHmutcell_', i, '/')
  isoform_list[[j]] <- readRDS(paste0(path, sra, "_isoform-a.bam_frequency_all.rds"))
  j = j + 1
}

# Name the elements of isoform_list
names(isoform_list) <- paste0("YTHmutcell_", folders)
saveRDS(isoform_list, "/gpfs/work/bio/haozhewang17/data/velocity/data_73/isoform_73_YTHmut.rds")
```
We will get the 
Processed data:

SNP_all

Frequency data: frequency_pre_YTHmut, frequency_spliced_YTHmut, frequency_spliced_YTH, frequency_spliced_YTHmut,frequency_isoform_YTH,
frequency_isoform_YTHmut

Downloaded data (from https://www.cell.com/cell-genomics/fulltext/S2666-979X(24)00331-8):

expression_TPM

gene_informations

label_SigRM

## pre-process
### extract the modification read counts and un-modification read counts
```{r eval=FALSE, include=FALSE}

##  pre_YTH 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_pre_YTH.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_pre_YTH.rds")
saveRDS(frequency_all,"reads_counts_pre_YTH.rds")
saveRDS(error_reads_list,"error_reads_list_pre_YTH.rds")
saveRDS(methylation_rate,"methylation_rate_pre_YTH.rds")
##  pre_YTHmut 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_pre_YTHmut.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_pre_YTHmut.rds")
saveRDS(frequency_all,"reads_counts_pre_YTHmut.rds")
saveRDS(error_reads_list,"error_reads_list_pre_YTHmut.rds")
saveRDS(methylation_rate,"methylation_rate_pre_YTHmut.rds")



##  spliced_YTH 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_spliced_YTH.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_spliced_YTH.rds")
saveRDS(frequency_all,"reads_counts_spliced_YTH.rds")
saveRDS(error_reads_list,"error_reads_list_spliced_YTH.rds")
saveRDS(methylation_rate,"methylation_rate_spliced_YTH.rds")



##  spliced_YTHmut 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_spliced_YTHmut.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_spliced_YTHmut.rds")
saveRDS(frequency_all,"reads_counts_spliced_YTHmut.rds")
saveRDS(error_reads_list,"error_reads_list_spliced_YTHmut.rds")
saveRDS(methylation_rate,"methylation_rate_spliced_YTHmut.rds")


##  isoform_YTH 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_isoform_YTH.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_isoform_YTH.rds")
saveRDS(frequency_all,"reads_counts_isoform_YTH.rds")
saveRDS(error_reads_list,"error_reads_list_isoform_YTH.rds")
saveRDS(methylation_rate,"methylation_rate_isoform_YTH.rds")


##  isoform_YTHmut 



frequency_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/frequency_isoform_YTHmut.rds")


testSNP_all <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")


head(frequency_all[[1]])
testSNP_all


library(dplyr)
# get the number of cell number of sites

num_cell <- length(frequency_all)
num_sites <- length(testSNP_all)

# get the index of the + - STAND 

strand <- testSNP_all$transcript_strand
strand_plus <- which(strand=="+")
strand_minus <- which(strand=="-")

error_rate_list <- list()

error_reads_list <- list()
methylation_rate <- list()

site_length_nomatch <-  which(lapply(frequency_all, function(x) dim(x)[1])!=length(testSNP_all))


for (i in 1:num_cell){
  
  # change to data frame and add strand column
  
  frequency_all[[i]] <-  data.frame(frequency_all[[i]])
  
  # get methylated reads 
  
  methylated_reads <- rep(0, num_sites)
  methylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,4]
  methylated_reads[strand_minus] <-frequency_all[[i]][strand_minus,1]
  
  # get unmethylated reads 
  
  unmethylated_reads <- rep(0, num_sites)
  unmethylated_reads[strand_plus] <- frequency_all[[i]][strand_plus,2]
  unmethylated_reads[strand_minus] <- frequency_all[[i]][strand_minus,3]
  
  
  
  # get error reads 
  
  error_reads <- rep(0, num_sites)
  error_reads[strand_plus] <- frequency_all[[i]][strand_plus,1]+frequency_all[[i]][strand_plus,3]+
    frequency_all[[i]][strand_plus,5]
  error_reads[strand_minus] <- frequency_all[[i]][strand_minus,2]+frequency_all[[i]][strand_minus,4]+
    frequency_all[[i]][strand_minus,5]
  
  
  
  frequency_all[[i]] <- frequency_all[[i]] %>% 
    mutate(methylated_reads=methylated_reads)%>% 
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_reads=error_reads)%>% 
    select(methylated_reads,unmethylated_reads,error_reads)
  
  error_reads_list[[i]] <-  frequency_all[[i]]$error_reads
  
  frequency_all[[i]] <- frequency_all[[i]] %>%
    mutate(methylated_reads=methylated_reads)%>%
    mutate(unmethylated_reads=unmethylated_reads)%>%
    mutate(error_rate=error_reads/(methylated_reads+unmethylated_reads+error_reads))%>%
    select(methylated_reads,unmethylated_reads,error_rate)
  
  error_rate_list[[i]] <-  frequency_all[[i]]$error_rate
  frequency_all[[i]] <- frequency_all[[i]] %>% select(methylated_reads,unmethylated_reads)
  
  methylation_rate[[i]] <- frequency_all[[i]]$methylated_reads/(frequency_all[[i]]$methylated_reads+frequency_all[[i]]$unmethylated_reads)
}




setwd("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process")
saveRDS(error_rate_list,"error_rate_list_isoform_YTHmut.rds")
saveRDS(frequency_all,"reads_counts_isoform_YTHmut.rds")
saveRDS(error_reads_list,"error_reads_list_isoform_YTHmut.rds")
saveRDS(methylation_rate,"methylation_rate_isoform_YTHmut.rds")
```
### calculate the error rate for site
```{r eval=FALSE, include=FALSE}
# calculate teh meta_error_rate for each site

## isoform
error_reads_list_isoform_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_isoform_YTH.rds")

error_reads_list_isoform_YTH_sum <- Reduce("+", error_reads_list_isoform_YTH)

reads_counts_isoform_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_isoform_YTH.rds")

summed_vectors_list <- lapply(reads_counts_isoform_YTH, function(df) rowSums(df))

reads_counts_isoform_YTH_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_isoform_YTH <- error_reads_list_isoform_YTH_sum/
  (error_reads_list_isoform_YTH_sum+reads_counts_isoform_YTH_sum) 



error_reads_list_isoform_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_isoform_YTHmut.rds")

error_reads_list_isoform_YTHmut_sum <- Reduce("+", error_reads_list_isoform_YTHmut)

reads_counts_isoform_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_isoform_YTHmut.rds")

summed_vectors_list <- lapply(reads_counts_isoform_YTHmut, function(df) rowSums(df))

reads_counts_isoform_YTHmut_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_isoform_YTHmut <- error_reads_list_isoform_YTHmut_sum/
  (error_reads_list_isoform_YTHmut_sum+reads_counts_isoform_YTHmut_sum) 


error_rate_meta_isoform <- (error_reads_list_isoform_YTHmut_sum+error_reads_list_isoform_YTH_sum)/
  (error_reads_list_isoform_YTHmut_sum+reads_counts_isoform_YTHmut_sum+
     error_reads_list_isoform_YTH_sum+reads_counts_isoform_YTH_sum) 

saveRDS(error_rate_meta_isoform_YTH,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_isoform_YTH.rds")
saveRDS(error_rate_meta_isoform_YTHmut,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_isoform_YTHmut.rds")
saveRDS(error_rate_meta_isoform,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_isoform.rds")

## spliced

rm(list = ls())
gc()
error_reads_list_spliced_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_spliced_YTH.rds")

error_reads_list_spliced_YTH_sum <- Reduce("+", error_reads_list_spliced_YTH)

reads_counts_spliced_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_spliced_YTH.rds")

summed_vectors_list <- lapply(reads_counts_spliced_YTH, function(df) rowSums(df))

reads_counts_spliced_YTH_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_spliced_YTH <- error_reads_list_spliced_YTH_sum/
  (error_reads_list_spliced_YTH_sum+reads_counts_spliced_YTH_sum) 



error_reads_list_spliced_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_spliced_YTHmut.rds")

error_reads_list_spliced_YTHmut_sum <- Reduce("+", error_reads_list_spliced_YTHmut)

reads_counts_spliced_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_spliced_YTHmut.rds")

summed_vectors_list <- lapply(reads_counts_spliced_YTHmut, function(df) rowSums(df))

reads_counts_spliced_YTHmut_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_spliced_YTHmut <- error_reads_list_spliced_YTHmut_sum/
  (error_reads_list_spliced_YTHmut_sum+reads_counts_spliced_YTHmut_sum) 


error_rate_meta_spliced <- (error_reads_list_spliced_YTHmut_sum+error_reads_list_spliced_YTH_sum)/
  (error_reads_list_spliced_YTHmut_sum+reads_counts_spliced_YTHmut_sum+
     error_reads_list_spliced_YTH_sum+reads_counts_spliced_YTH_sum) 

saveRDS(error_rate_meta_spliced_YTH,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_spliced_YTH.rds")
saveRDS(error_rate_meta_spliced_YTHmut,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_spliced_YTHmut.rds")
saveRDS(error_rate_meta_spliced,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_spliced.rds")
## pre


rm(list = ls())
gc()
error_reads_list_pre_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_pre_YTH.rds")

error_reads_list_pre_YTH_sum <- Reduce("+", error_reads_list_pre_YTH)

reads_counts_pre_YTH <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_pre_YTH.rds")

summed_vectors_list <- lapply(reads_counts_pre_YTH, function(df) rowSums(df))

reads_counts_pre_YTH_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_pre_YTH <- error_reads_list_pre_YTH_sum/
  (error_reads_list_pre_YTH_sum+reads_counts_pre_YTH_sum) 



error_reads_list_pre_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_reads_list_pre_YTHmut.rds")

error_reads_list_pre_YTHmut_sum <- Reduce("+", error_reads_list_pre_YTHmut)

reads_counts_pre_YTHmut <- 
  readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_pre_YTHmut.rds")

summed_vectors_list <- lapply(reads_counts_pre_YTHmut, function(df) rowSums(df))

reads_counts_pre_YTHmut_sum <- Reduce("+", summed_vectors_list)

error_rate_meta_pre_YTHmut <- error_reads_list_pre_YTHmut_sum/
  (error_reads_list_pre_YTHmut_sum+reads_counts_pre_YTHmut_sum) 


error_rate_meta_pre <- (error_reads_list_pre_YTHmut_sum+error_reads_list_pre_YTH_sum)/
  (error_reads_list_pre_YTHmut_sum+reads_counts_pre_YTHmut_sum+
     error_reads_list_pre_YTH_sum+reads_counts_pre_YTH_sum) 

saveRDS(error_rate_meta_pre_YTH,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_pre_YTH.rds")
saveRDS(error_rate_meta_pre_YTHmut,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_pre_YTHmut.rds")
saveRDS(error_rate_meta_pre,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_pre.rds")

```
### filter the site 
From 1,599 cells (1,209 test cells and 390 YTH domain-mutated negative controls), we initially extracted 2,371,331 candidate sites. We applied stringent quality filters, excluding sites with either: (1) an error rate exceeding 1%, or (2) median read coverage below 10 in spliced mRNA or unspliced mRNA. This yielded 37,199 high-confidence m6A sites for downstream analysis.
```{r eval=FALSE, include=FALSE}
# Load necessary library
library(VeloRM)

# Read metadata
error_rate_meta_spliced <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_spliced.rds")
error_rate_meta_pre <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/error_rate_meta_pre.rds")
index_error <- intersect(which(error_rate_meta_spliced <= 0.01), which(error_rate_meta_pre <= 0.01))

# Free up memory
rm(error_rate_meta_spliced, error_rate_meta_pre)
gc()

# Read counts
reads_counts_spliced_YTHmut <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_spliced_YTHmut.rds")
reads_counts_spliced_YTH <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_spliced_YTH.rds")

# Process spliced data
spliced_meth_test <- as.data.frame(lapply(reads_counts_spliced_YTH, function(df) df[["methylated_reads"]]))
spliced_unmeth_test <- as.data.frame(lapply(reads_counts_spliced_YTH, function(df) df[["unmethylated_reads"]]))
spliced_meth_control <- as.data.frame(lapply(reads_counts_spliced_YTHmut, function(df) df[["methylated_reads"]]))
spliced_unmeth_control <- as.data.frame(lapply(reads_counts_spliced_YTHmut, function(df) df[["unmethylated_reads"]]))

# Free up memory
rm(reads_counts_spliced_YTH, reads_counts_spliced_YTHmut)
gc()

reads_counts_pre_YTHmut <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_pre_YTHmut.rds")
reads_counts_pre_YTH <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/reads_counts_pre_YTH.rds")

# Process pre-mRNA data
pre_meth_test <- as.data.frame(lapply(reads_counts_pre_YTH, function(df) df[["methylated_reads"]]))
pre_unmeth_test <- as.data.frame(lapply(reads_counts_pre_YTH, function(df) df[["unmethylated_reads"]]))
pre_meth_control <- as.data.frame(lapply(reads_counts_pre_YTHmut, function(df) df[["methylated_reads"]]))
pre_unmeth_control <- as.data.frame(lapply(reads_counts_pre_YTHmut, function(df) df[["unmethylated_reads"]]))

# Free up memory
rm(reads_counts_pre_YTH, reads_counts_pre_YTHmut)
gc()

# Set row names
rownames(spliced_meth_test) <- rownames(spliced_unmeth_test) <- rownames(spliced_meth_control) <- 
  rownames(spliced_unmeth_control) <- paste0("site_", 1:dim(spliced_unmeth_test)[1])

rownames(pre_meth_test) <- rownames(pre_unmeth_test) <- rownames(pre_meth_control) <- 
  rownames(pre_unmeth_control) <- paste0("site_", 1:dim(pre_unmeth_test)[1])

# Calculate spliced and pre-mRNA test matrices
spliced_test <- cbind(spliced_meth_test, spliced_meth_control) + cbind(spliced_unmeth_test, spliced_unmeth_control)


pre_test <- cbind(pre_meth_test, pre_meth_control) + cbind(pre_unmeth_test, pre_unmeth_control)


# Compute medians
spliced_median <- apply(spliced_test, 1, function(x) median(x, na.rm = TRUE))
rm(spliced_test)
gc()

pre_median <- apply(pre_test, 1, function(x) median(x, na.rm = TRUE))
rm(pre_test)
gc()

# Find common indices
index_median <- intersect(which(spliced_median >= 10), which(pre_median >= 10))

gc()



index <- intersect(index_error,index_median)

spliced_meth_test <- spliced_meth_test[index,]
spliced_unmeth_test <- spliced_unmeth_test[index,]
pre_meth_test <- pre_meth_test[index,]
pre_unmeth_test <- pre_unmeth_test[index,]

spliced_meth_control <- spliced_meth_control[index,]
spliced_unmeth_control <- spliced_unmeth_control[index,]
pre_meth_control <- pre_meth_control[index,]
pre_unmeth_control <- pre_unmeth_control[index,]


print("finish index")
saveRDS(spliced_median,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/spliced_median.rds")
saveRDS(pre_median,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/pre_median.rds")
saveRDS(index_median,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/index_median.rds")
saveRDS(index_error,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/index_error.rds")
saveRDS(index,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/index.rds")


reads_counts_isoform_YTHmut <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity_finish_2/pre_process/reads_counts_pre_YTHmut.rds")
reads_counts_isoform_YTH <- readRDS("/gpfs/work/bio/haozhewang17/data/velocity_finish_2/pre_process/reads_counts_pre_YTH.rds")

# Process pre-mRNA data
isoform_meth_test <- as.data.frame(lapply(reads_counts_isoform_YTH, function(df) df[["methylated_reads"]]))
isoform_unmeth_test <- as.data.frame(lapply(reads_counts_isoform_YTH, function(df) df[["unmethylated_reads"]]))
isoform_meth_control <- as.data.frame(lapply(reads_counts_isoform_YTHmut, function(df) df[["methylated_reads"]]))
isoform_unmeth_control <- as.data.frame(lapply(reads_counts_isoform_YTHmut, function(df) df[["unmethylated_reads"]]))

rm(reads_counts_isoform_YTH, reads_counts_isoform_YTHmut)
gc()

rownames(isoform_meth_test) <- rownames(isoform_unmeth_test) <- rownames(isoform_meth_control) <- 
  rownames(isoform_unmeth_control) <- paste0("site_", 1:dim(isoform_unmeth_test)[1])

isoform_meth_test <- isoform_meth_test[index,]
isoform_unmeth_test <- isoform_unmeth_test[index,]
isoform_meth_control <- isoform_meth_control[index,]
isoform_unmeth_control <- isoform_unmeth_control[index,]

test.list <- list(spliced_meth_test,spliced_unmeth_test,pre_meth_test,pre_unmeth_test,
                   isoform_meth_test,isoform_unmeth_test)
control.list <- list(spliced_meth_control,spliced_unmeth_control,pre_meth_control,
                      pre_unmeth_control,isoform_meth_control,isoform_unmeth_control)

res_preprocess <- methylation.sites.preprocess(test.list,control.list)

saveRDS(test.list,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/test.list.rds")
saveRDS(control.list,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/control.list.rds")
saveRDS(res_preprocess,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/res_preprocess.rds")

SNP <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/origin/SNP_all_7374.rds")
SNP_index <- SNP[index]
saveRDS(SNP_index,"/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/SNP_37199.rds")
```
### DESeq2 to test the differential methylation between the spliced mRNA and pre-mRNA
```{r eval=FALSE, include=FALSE}
# Load input data
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data

# Set random seed for reproducibility
set.seed(42)

# Record start time
start_time <- Sys.time()

# Load DESeq2 package
library(DESeq2)

# Define DESeq2 differential expression analysis function
DESeq2_test <- function(meth_control, meth_test, unmeth_control, unmeth_test) {
  # Convert input data to data frames
  meth_control <- data.frame(meth_control)
  meth_test <- data.frame(meth_test)
  unmeth_control <- data.frame(unmeth_control)
  unmeth_test <- data.frame(unmeth_test)
  
  # Function to calculate size factor
  sizeFactor <- function(data) {
    # Ensure data is not less than 0
    data <- as.matrix(data)
    
    # Log-transform the data
    log_data <- log(data)
    log_data[is.infinite(log_data)] <- NA
    log_mean <- rowMeans(log_data, na.rm = TRUE)
    log_s <- log_data - log_mean
    
    # Calculate size factor
    s_size <- exp(apply(log_s, 2, function(x) median(x, na.rm = TRUE)))
    return(s_size)
  }
  
  # Calculate size factors for spliced and unspliced data
  spliced_total <- meth_control + unmeth_control
  unspliced_total <- meth_test + unmeth_test
  
  spliced_size_factor <- sizeFactor(spliced_total)
  unspliced_size_factor <- sizeFactor(unspliced_total)
  
  # Get the number of control and test samples
  num_control <- dim(unmeth_control)[2]
  num_test <- dim(unmeth_test)[2]
  
  # Combine all data
  counts <- cbind(meth_control, meth_test, unmeth_control, unmeth_test)
  
  # Design the experiment matrix
  design = data.frame(
    Trt = rep(c(rep("control", num_control), rep("test", num_test)), 2), 
    Meth = c(rep("meth", (num_control + num_test)), rep("unmeth", (num_control + num_test)))
  )
  
  # Define the model
  model = ~ Meth + Trt + Meth:Trt
  
  # Create DESeqDataSet object
  dds <- DESeqDataSetFromMatrix(countData = counts,
                                colData = design,
                                design = model)
  
  # Set size factors
  sizeFactors(dds) = rep(c(spliced_size_factor, unspliced_size_factor), 2)
  
  # Perform differential expression analysis
  dds <- DESeq(dds, test = "Wald")
  
  # Extract results
  res = DESeq2::results(dds, name = "Methunmeth.Trttest", cooksCutoff = FALSE)
  
  return(res)
}

# Call the function to perform differential expression analysis
result <- DESeq2_test(spliced.meth.test, unspliced.meth.test, spliced.unmeth.test, unspliced.unmeth.test)

# Record end time
end_time <- Sys.time()

# Print the runtime
print(end_time - start_time)


# Save the results
saveRDS(result, "/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/res_DESeq2.rds")
```
### SigRM to test the differential methylation between the test cell and control group
```{r eval=FALSE, include=FALSE}
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/test.list.rds")
control.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/control.list.rds")

## SigRM test 
library(SigRM)

spliced.meth.test <- (test.list[[1]])
spliced.unmeth.test <- (test.list[[2]])
unspliced.meth.test <- (test.list[[3]])
unspliced.unmeth.test <- (test.list[[4]])
spliced.meth.control <- (control.list[[1]])
spliced.unmeth.control <- (control.list[[2]])
unspliced.meth.control <- (control.list[[3]])
unspliced.unmeth.control <- (control.list[[4]])


SigRMtest_res_spliced <- SigRMtest(spliced.meth.control,spliced.meth.test,
                                   spliced.unmeth.control,spliced.unmeth.test)

SigRMtest_res_unspliced <- SigRMtest(unspliced.meth.control,unspliced.meth.test,
                                     unspliced.unmeth.control,unspliced.unmeth.test)
saveRDS(SigRMtest_res_spliced,
        "/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/SigRMtest_res_spliced.rds")
saveRDS(SigRMtest_res_unspliced,
        "/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/SigRMtest_res_unspliced.rds")
```
We will get the 
Processed data:
SNP_37199: the SNP information for the m6A site for further analysis.
test_list: the test cell reads counts (spliced_meth_test,spliced_unmeth_test,pre_meth_test,pre_unmeth_test,
                   isoform_meth_test,isoform_unmeth_test)
control_list: the control cell reads counts (spliced_meth_control,spliced_unmeth_control,pre_meth_control,pre_unmeth_control,
                   isoform_meth_control,isoform_unmeth_control)
res_preprocess: the list processed by VeloRM
res_DESeq2: the DESeq2 results
SigRMtest_res_unspliced: the SigRM test results comparing to the control for unspliced read counts
SigRMtest_res_spliced: the SigRM test results comparing to the control for spliced read counts
## analysis
### splicing junction distance of the SNP sites
```{r eval=FALSE, include=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(dplyr)
library(tidyr)

snp <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SNP_37199.rds")
snp_pos <- snp@ranges@start
snp_chr_length <- snp@seqnames@lengths
snp_chr_seq <- snp@seqnames@values
snp_chr <- as.character(rep(snp_chr_seq,snp_chr_length)) 
snp_strand <- snp@elementMetadata@listData[["transcript_strand"]]
snp_data <- as.data.frame(cbind(snp_chr,snp_pos,snp_strand))
snp_data$snp_pos <- as.numeric(snp_data$snp_pos)




# Read the GTF file from Ensembl
gtf_file <- "D:/data/VeloRM_data/scDART-seq/origin/gencode.v47.chr_patch_hapl_scaff.annotation.gtf"
gtf_data <- import(gtf_file)

# Filter for exons
exons <- gtf_data[gtf_data$type == "exon"]

# Convert to data frame
exons_df <- data.frame(
chr = as.character(seqnames(exons)),
start = start(exons),
end = end(exons),
strand = as.character(strand(exons)),
gene_id = mcols(exons)$gene_id,
transcript_id = mcols(exons)$transcript_id
)

# Sort exons by transcript and position
exons_df <- exons_df %>%
arrange(transcript_id, chr, start)

# Extract splicing junctions using intron boundaries
junctions <- exons_df %>%
group_by(transcript_id) %>%
arrange(if_else(strand == "+", start, -start)) %>%  # Sort based on strand
mutate(
  donor_site = end,                # End of the upstream exon
  acceptor_site = lead(start)      # Start of the downstream exon
) %>%
filter(!is.na(acceptor_site)) %>%   # Remove the last exon (no downstream exon)
ungroup() %>%
select(chr, donor_site, acceptor_site, gene_id, transcript_id, strand)


junctions_unique <- junctions %>%
 distinct(chr, donor_site, acceptor_site, .keep_all = TRUE)



junctions_long <- junctions_unique %>%
 pivot_longer(cols = c(donor_site, acceptor_site), 
              names_to = "site_type", 
              values_to = "site_position") %>%
 select(chr, site_position, site_type, gene_id, transcript_id, strand)



# Calculate distance from each SNP to the nearest junction (matching chr and strand)
snp_with_distance <- snp_data %>%
 rowwise() %>%
 mutate(
   nearest_junction_distance = min(
     abs(junctions_long$site_position[junctions_long$chr == snp_chr & junctions_long$strand == snp_strand] - snp_pos),
     na.rm = TRUE
   )
 )


# get normal chr
standard_chromosomes <- paste0("chr", c(1:22, "X", "Y"))

# get the exon with the normal chr
exons_filtered <- exons[seqnames(exons) %in% standard_chromosomes]



# calculate the distance 
exon_lengths <- width(exons_filtered)

# calculate the prob 
prob_weights <- exon_lengths / sum(exon_lengths)

# random sample
set.seed(123) 


#random_distances_list <- list()

#for (i in 1:10){
 selected_exons <- sample(seq_along(exons_filtered), size = 37199, replace = TRUE, prob = prob_weights)
 # choose the site in the select the exons
 random_positions <- start(exons[selected_exons]) + 
   floor(runif(37199, min = 0, max = exon_lengths[selected_exons]))
 
 # create the data frame
 random_sites <- data.frame(
   chr = as.character(seqnames(exons[selected_exons])),
   position = random_positions,
   strand = as.character(strand(exons[selected_exons]))
 )
 
 
 # Calculate distance from each SNP to the nearest junction (matching chr and strand)
 random_sites_with_distance <- random_sites %>%
   rowwise() %>%
   mutate(
     nearest_junction_distance = min(
       abs(junctions_long$site_position[junctions_long$chr == chr & junctions_long$strand == strand] - position),
       na.rm = TRUE
     )
   )
 random_distances_list  <- random_sites_with_distance$nearest_junction_distance
#}



random_distances_df <- data.frame(
 distance = random_distances_list,
 group = "Random"
)

# merge the data frame
real_distances_df <- data.frame(
 distance = snp_with_distance$nearest_junction_distance,
 group = "Real"
)


all_distances_df <- bind_rows(random_distances_df, real_distances_df)
group_order <- c("Real","Random")
all_distances_df$group <- factor(all_distances_df$group, levels = group_order)

saveRDS(all_distances_df,"D:/data/VeloRM_data/scDART-seq/results/splicing_junction_analysis/all_distances_df.rds")

```
#### visulization
We can see that the SNP is close to the splicing junction comparing to random sample
```{r}
library(ggplot2)
all_distances_df <- readRDS("D:/data/VeloRM_data/scDART-seq/results/splicing_junction_analysis/all_distances_df.rds")
#  Boxplot
plot <- ggplot(all_distances_df, aes(x = group, y = log10(distance+1), fill = group)) +
 geom_violin(trim = FALSE, alpha = 0.5) +
 geom_boxplot(width = 0.1, outlier.shape = NA)+
 scale_fill_manual(values = c("red",rep("lightblue", 10))) +
 labs(title = NULL,
      x = NULL,
      y = "log10(Distance (bp)+1)") +
 theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5),
       legend.position = "none")+coord_cartesian(ylim = c(0,5))
plot

```
### overall analysis
#### overall reads counts
```{r}
rm(list = ls())
gc()
# Load input data
test.list <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/test.list.rds")

# Extract different parts of the data
spliced.meth.test <- test.list[[1]]  # Methylated spliced data
spliced.unmeth.test <- test.list[[2]]  # Unmethylated spliced data
unspliced.meth.test <- test.list[[3]]  # Methylated unspliced data
unspliced.unmeth.test <- test.list[[4]]  # Unmethylated unspliced data


# Load input data
control.list <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/control.list.rds")

# Extract different parts of the data
spliced.meth.control <- control.list[[1]]  # Methylated spliced data
spliced.unmeth.control <- control.list[[2]]  # Unmethylated spliced data
unspliced.meth.control <- control.list[[3]]  # Methylated unspliced data
unspliced.unmeth.control <- control.list[[4]]  # Unmethylated unspliced data

spliced.meth.sum <- sum(spliced.meth.test)+sum(spliced.meth.control)
spliced.unmeth.sum <- sum(spliced.unmeth.test)+sum(spliced.unmeth.control)
unspliced.meth.sum <- sum(unspliced.meth.test)+sum(unspliced.meth.control)
unspliced.unmeth.sum <- sum(unspliced.unmeth.test)+sum(unspliced.unmeth.control)


# spliced.meth.sum, spliced.unmeth.sum, unspliced.meth.sum, unspliced.unmeth.sum

# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Create a data frame
data <- data.frame(
  category = c("mature mRNA", "pre mRNA"),  # spliced = mature mRNA, unspliced = pre mRNA
  meth = c(spliced.meth.sum, unspliced.meth.sum),      # methylated counts
  unmeth = c(spliced.unmeth.sum, unspliced.unmeth.sum) # unmethylated counts
)

# Calculate proportions
data <- data %>%
  mutate(
    total = meth + unmeth,                # total counts for each category
    meth_ratio = meth / total             # methylation proportion
  )

# Calculate mature mRNA (spliced) proportion of total
spliced_ratio <- (spliced.meth.sum + spliced.unmeth.sum) / 
                 (spliced.meth.sum + spliced.unmeth.sum + 
                  unspliced.meth.sum + unspliced.unmeth.sum)

# Visualization 1: Pie chart of mature mRNA vs pre mRNA proportion
pie_data <- data.frame(
  group = c("mature mRNA", "pre mRNA"),
  value = c(spliced_ratio, 1 - spliced_ratio)
)

ggplot(pie_data, aes(x = "", y = value, fill = group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(value*100, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(title = "mature mRNA/(mature mRNA + pre mRNA) Ratio") +
  theme_void()

# Visualization 2: Bar plot of methylation ratio in each category
ggplot(data, aes(x = category, y = meth_ratio, fill = category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(meth_ratio*100, 1), "%")), vjust = -0.5) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.008)) +
  labs(title = "Methylation Ratio in Spliced and Unspliced",
       y = "Methylation Ratio",
       x = "") +
  theme_minimal()

```

#### read the SigRM results
```{r}
res_DESeq2 <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_DESeq2.rds")

log2FC <- res_DESeq2@listData[["log2FoldChange"]]
pvals <- res_DESeq2@listData[["pvalue"]]

plot_data <- data.frame(log2FC = log2FC, 
                       pvalue = pvals,
                       neg_log10_p = -log10(pvals))


ggplot(plot_data, aes(x = log2FC, y = neg_log10_p)) +
  # Add light gray background for -1 < log2FC < 1 (non-significant FC region)
  annotate("rect", xmin = -1, xmax = 1, ymin = 0, ymax = 30, 
           fill = "gray90", alpha = 0.3) +
  
  # Color points based on significance and log2FC:
  # - Red: log2FC > 1 & p < 0.05 (significantly up-regulated)
  # - Blue: log2FC < -1 & p < 0.05 (significantly down-regulated)
  # - Gray: Not significant or |log2FC|  1
  geom_point(aes(color = case_when(
    log2FC > 1 & pvalue < 0.05 ~ "Spliced",
    log2FC < -1 & pvalue < 0.05 ~ "Unspliced",
    TRUE ~ "Not significant"
  )), alpha = 0.6) +
  
  # Manual color scale (red, blue, gray)
  scale_color_manual(
    values = c(
      "Spliced" = "red",
      "Unspliced" = "blue",
      "Not significant" = "gray50"
    ),
    name = "Regulation"  # Legend title
  ) +
  
  # Add dashed lines for thresholds:
  # - Vertical: log2FC = 1
  # - Horizontal: p-value = 0.05 (-log10(0.05)  1.3)
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  
  # Axis labels and theme
  labs(
    x = "log2 Fold Change", 
    y = "-log10(p-value)",
    title = NULL
  ) +
  theme_minimal() +
  
  # Set axis limits (zoomed-in view)
  coord_cartesian(ylim = c(0, 30), xlim = c(-4, 4))


SigRMtest_res_spliced <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SigRMtest_res_spliced.rds")
SigRMtest_res_unspliced <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SigRMtest_res_unspliced.rds")

SigRMtest_res_spliced_p_value <- SigRMtest_res_spliced[["p value"]]
SigRMtest_res_unspliced_p_value <- SigRMtest_res_unspliced[["p value"]]


SigRMtest_res_spliced_p_value_median <- apply(SigRMtest_res_spliced_p_value,1,function(x) median(x,na.rm=TRUE))

SigRMtest_res_unspliced_p_value_median <- apply(SigRMtest_res_unspliced_p_value,1,function(x) median(x,na.rm=TRUE))


index_DESeq2_significant <- which(res_DESeq2@listData$pvalue<=0.05)
# Compute -log10(p-values)
log_p_spliced <- -log10(SigRMtest_res_spliced_p_value_median)
log_p_unspliced <- -log10(SigRMtest_res_unspliced_p_value_median)

# Create a combined dataframe for plotting
library(ggplot2)
df <- data.frame(
  log_p_value = c(log_p_spliced[index_DESeq2_significant], log_p_unspliced[index_DESeq2_significant]),
  Type = rep(c("Spliced", "Unspliced"), each = length(log_p_spliced[index_DESeq2_significant]))
)

# comparison the distribution
comparison_distribution <- ggplot(df, aes(x = log_p_value, fill = Type)) +
  geom_density(alpha = 0.5, bw = 0.3) +  # Adjust bw for smoothness
  geom_vline(xintercept = -log10(0.05), linetype = "dashed", color = "red", linewidth = 1) +  # Add vertical line
    annotate("text", x = -log10(0.05)+0.05, y = 0.5, label = "p-value = 0.05", color = "red", hjust = 0) +
  labs(title = "Median p-value Distribution",
       x = "-log10(p-value)",
       y = "Density") +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 3.5))+
  theme(legend.position = c(0.8, 0.8),
        plot.title = element_text(hjust = 0.5, vjust = 2))  # Move legend inside the plot
comparison_distribution

# the meta OR infomation
res_preprocess <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_preprocess.rds")

# want the site p_value<=0.05, meta log2(OR) >=1
spliced_num_sites <- length(which(SigRMtest_res_spliced_p_value_median<=0.05&res_preprocess$spliced_methylation_OR_value_meta-res_preprocess$unspliced_methylation_OR_value_meta>=1&res_DESeq2@listData$pvalue<=0.05))
unspliced_num_sites <- length(which(SigRMtest_res_unspliced_p_value_median<=0.05&res_preprocess$unspliced_methylation_OR_value_meta-res_preprocess$spliced_methylation_OR_value_meta>=1&res_DESeq2@listData$pvalue<=0.05))

Other_num_sites <- length(SigRMtest_res_spliced_p_value_median) - spliced_num_sites -unspliced_num_sites

library(ggplot2)

# Create a data frame for pie chart
df_pie <- data.frame(
  Type = c("Spliced", "Unspliced", "Other"),
  Count = c(spliced_num_sites, unspliced_num_sites, Other_num_sites)
)

# Calculate percentages
df_pie$Percentage <- round(df_pie$Count / sum(df_pie$Count) * 100, 1)

# Create separate data frames for each piece of text you want to display
df_pie$text_count <- as.character(df_pie$Count)
df_pie$text_percentage <- paste0(df_pie$Percentage, "%")

# Define custom colors
#custom_colors <- c( "#ef767b","#f9e7a7", "#43a3ef")

# Plot the pie chart with 4 separate texts
pie_chart <- ggplot(df_pie, aes(x = "", y = Count, fill = Type)) +
  geom_bar(stat = "identity", width = 1) +  
  coord_polar("y", start = 0) +  # Convert to pie chart
  geom_text(aes(label = text_count), 
            position = position_stack(vjust = 0.15), size = 5) +  # Position for the number
  geom_text(aes(label = text_percentage), 
            position = position_stack(vjust = 0.85), size = 5) +  # Apply custom colors
  labs(title = "Significant Sites Proportion") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank(),   # Remove x-axis labels
    axis.ticks = element_blank(),    # Remove x-axis ticks
    panel.grid = element_blank(),    # Remove grid
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, vjust = 0, size = 14),  # Move title outside center
    legend.title = element_text(size = 10),  # Larger legend title text
    legend.text = element_text(size = 8),  # Larger legend labels
    legend.key.size = unit(0.8, "cm")  # Increase the size of the legend keys (colors)
  )

print(pie_chart)
```
### regulatory mechanisms
#### Find the site that is differentially methylated in spliced and pre-mRNA.
```{r}
rm(list = ls())
gc()
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(MetaTX)
library(ggplot2)
library(dplyr)

res_DESeq2 <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_DESeq2.rds")


index_differentially_meth <- which(res_DESeq2@listData$pvalue<=0.05)


```
#### Finding that the sites in pre-mRNA are significant compared to the control but insignificant in spliced mRNA, as well as sites in spliced mRNA being significant compared to the control but insignificant in pre-mRNA.
```{r}
SigRMtest_res_spliced <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SigRMtest_res_spliced.rds")
SigRMtest_res_unspliced <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SigRMtest_res_unspliced.rds")

SigRMtest_res_spliced_pvalue_median <- apply(SigRMtest_res_spliced$`p value`,1,function(x) median(x,na.rm = TRUE))
SigRMtest_res_unspliced_pvalue_median <- apply(SigRMtest_res_unspliced$`p value`,1,function(x) median(x,na.rm = TRUE))

index_pre_siginificant_spliced_insiginificant <- which(SigRMtest_res_unspliced_pvalue_median<=0.05)

index_spliced_siginificant_pre_insiginificant <- which(SigRMtest_res_spliced_pvalue_median<=0.05)
```

#### Finding sites where methylation levels are high in pre-mRNA but low in spliced mRNA, as well as sites where methylation levels are high in spliced mRNA but low in pre-mRNA.
```{r}
res_preprocess <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_preprocess.rds")
spliced_methylation_OR_value_meta <- res_preprocess$spliced_methylation_OR_value_meta
unspliced_methylation_OR_value_meta <- res_preprocess$unspliced_methylation_OR_value_meta


index_pre_high_spliced_low <- which(unspliced_methylation_OR_value_meta-spliced_methylation_OR_value_meta>=1)
index_spliced_high_pre_low <- which(spliced_methylation_OR_value_meta-unspliced_methylation_OR_value_meta>=1)
```

#### metaTX plot 
```{r}
library(MetaTX)
intersect_pre_dominate <- intersect(intersect(index_pre_high_spliced_low,index_pre_siginificant_spliced_insiginificant),index_differentially_meth)

intersect_spliced_dominate <- intersect(intersect(index_spliced_high_pre_low,index_spliced_siginificant_pre_insiginificant),index_differentially_meth)

index_not_dominate <- setdiff(1:31799,union(intersect_pre_dominate,intersect_spliced_dominate)) 

# metaTX


SNP_37199 <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SNP_37199.rds")
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

SNP_intersect_pre_dominate <- SNP_37199[intersect_pre_dominate]
SNP_intersect_spliced_dominate <- SNP_37199[intersect_spliced_dominate]
SNP_intersect_not_dominate <- SNP_37199[index_not_dominate]

# minus transcript strand +1, plus transcript strand -1
SNP_intersect_pre_dominate_adjusted <- SNP_intersect_pre_dominate
pos_strand <- mcols(SNP_intersect_pre_dominate)$transcript_strand == "+"
neg_strand <- mcols(SNP_intersect_pre_dominate)$transcript_strand == "-"

SNP_intersect_pre_dominate_adjusted[pos_strand]@ranges@start <- 
   as.integer(SNP_intersect_pre_dominate_adjusted[pos_strand]@ranges@start - 1)
SNP_intersect_pre_dominate_adjusted[neg_strand]@ranges@start <- 
   as.integer(SNP_intersect_pre_dominate_adjusted[neg_strand]@ranges@start + 1)


SNP_intersect_spliced_dominate_adjusted <- SNP_intersect_spliced_dominate
pos_strand <- mcols(SNP_intersect_spliced_dominate)$transcript_strand == "+"
neg_strand <- mcols(SNP_intersect_spliced_dominate)$transcript_strand == "-"

SNP_intersect_spliced_dominate_adjusted[pos_strand]@ranges@start <- 
   as.integer(SNP_intersect_spliced_dominate_adjusted[pos_strand]@ranges@start - 1)
SNP_intersect_spliced_dominate_adjusted[neg_strand]@ranges@start <- 
   as.integer(SNP_intersect_spliced_dominate_adjusted[neg_strand]@ranges@start + 1)




SNP_intersect_not_dominate_adjusted <- SNP_intersect_not_dominate
pos_strand <- mcols(SNP_intersect_not_dominate)$transcript_strand == "+"
neg_strand <- mcols(SNP_intersect_not_dominate)$transcript_strand == "-"

SNP_intersect_not_dominate_adjusted[pos_strand]@ranges@start <- 
   as.integer(SNP_intersect_not_dominate_adjusted[pos_strand]@ranges@start - 1)
SNP_intersect_not_dominate_adjusted[neg_strand]@ranges@start <- 
   as.integer(SNP_intersect_not_dominate_adjusted[neg_strand]@ranges@start + 1)


remap_results_pre_dominate <- remapCoord(features = SNP_intersect_pre_dominate_adjusted, txdb = txdb)
remap_results_spliced_dominate <- remapCoord(features = SNP_intersect_spliced_dominate_adjusted, txdb = txdb)
remap_results_not_dominate <- remapCoord(features = SNP_intersect_not_dominate_adjusted, txdb = txdb)
  
# Step 2: Create plots for each subset
plot_pre_dominate <- directPlot(remap_results_pre_dominate)
plot_spliced_dominate <- directPlot(remap_results_spliced_dominate)
plot_not_dominate <- directPlot(remap_results_not_dominate)

data_a <- ggplot_build(plot_pre_dominate)$data[[1]] %>% mutate(Group = "Precursor")
data_b <- ggplot_build(plot_spliced_dominate)$data[[1]] %>% mutate(Group = "Mature")
data_c <- ggplot_build(plot_not_dominate)$data[[1]] %>% mutate(Group = "Other")

# Combine all datasets
combined_data <- bind_rows(data_a, data_b, data_c)#, data_intersect)

region_rects <- data.frame(
    xmin = c(0.2, 0.4, 0.6),  # Start positions of each region
    xmax = c(0.4, 0.6, 0.8),  # End positions of each region
    ymin = min(combined_data$y, na.rm = TRUE) - 0.1,  # Adjust below data range
    ymax = min(combined_data$y, na.rm = TRUE) - 0.05, # Slightly higher to make room
    Region = c("5' UTR", "CDS", "3' UTR")
)

# Calculate the height of 5' UTR and 3' UTR regions
utr_height <- region_rects$ymax[region_rects$Region == "5' UTR"] - region_rects$ymin[region_rects$Region == "5' UTR"]

# Increase the height of CDS region by 20% (can be adjusted as needed)
cds_height <- utr_height * 2 # Increase height by 30%

# Set ymin and ymax for CDS so that it is centered with respect to 5' UTR and 3' UTR
region_rects$ymin[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] - 0.5*utr_height  # Decrease ymin for CDS
region_rects$ymax[region_rects$Region == "CDS"] <- region_rects$ymin[region_rects$Region == "CDS"] + cds_height  # Increase ymax for CDS


compare_metaTX_plot <-ggplot(combined_data, aes(x = x, y = y, color = Group))+
  geom_line(size=1)+
    # Add region highlights using geom_rect, with consistent positioning across all facets
    geom_rect(data = region_rects, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = Region),
              inherit.aes = FALSE, color = "black", alpha = 0.4, 
              linewidth = 0.5) +  # Default linewidth for all regions
    # Set fill for CDS region to grey (lighter than black)
    geom_rect(data = subset(region_rects, Region == "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE, fill = "grey", color = "black", alpha = 0.4, linewidth = 0.5) +  # Grey fill for CDS
    # Adjust 5' UTR and 3' UTR fills to a very dark grey, closer to black
    geom_rect(data = subset(region_rects, Region != "CDS"), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                                                                fill = factor(Region, levels = c("5' UTR", "3' UTR"))), 
              inherit.aes = FALSE, color = "black", alpha = 0.3, linewidth = 0.5) + 
    scale_fill_manual(values = c("5' UTR" = grey(0.15),  # Very dark grey for 5' UTR, closer to black
                                "CDS" = "grey",           # Grey for CDS
                                "3' UTR" = grey(0.15))) +  # Very dark grey for 3' UTR
    # Adjust y position for the text annotations to place them above the rectangles
    geom_text(data = region_rects, aes(x = (xmin + xmax) / 2, y = ymax - 0.5, label = Region),
              inherit.aes = FALSE, color = "black", size = 3, hjust = 0.5) +
    # Remove Region fill legend
    guides(fill = "none")+scale_color_manual(values = c(
  "Other" = "#00ba38",      
  "Mature" = "#619cff",     
  "Precursor" = "#F8766d"         
),name = "Higher on")+theme_minimal()+
  theme(
    legend.position = c(0.8, 0.7)  )+xlab(NULL)+ylab(NULL)
compare_metaTX_plot
```

#### motif plot
```{r}
library(GenomicRanges)
library(Biostrings)
library(BSgenome)
library(BSgenome.Hsapiens.UCSC.hg38)
genome <- BSgenome.Hsapiens.UCSC.hg38

## pre 
SNP_intersect_pre_dominate <- SNP_37199[intersect_pre_dominate]
# transcript_strand  +
SNP_intersect_pre_dominate_plus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="+")] 


extended_range_intersect_pre_dominate_plus <- resize(SNP_intersect_pre_dominate_plus, width = 11, fix = "center")

# getseq
sequence_intersect_pre_dominate_plus <- getSeq(genome, extended_range_intersect_pre_dominate_plus)

sequence_intersect_pre_dominate_plus <- subseq(sequence_intersect_pre_dominate_plus, start = 3, end = 7)

# transcript_strand  -
SNP_intersect_pre_dominate_minus <-SNP_intersect_pre_dominate[which(SNP_intersect_pre_dominate$transcript_strand=="-")] 

extended_range_intersect_pre_dominate_minus <- resize(SNP_intersect_pre_dominate_minus, width = 11, fix = "center")

# getseq
sequence_intersect_pre_dominate_minus <- getSeq(genome, extended_range_intersect_pre_dominate_minus)

sequence_intersect_pre_dominate_minus <- (subseq(sequence_intersect_pre_dominate_minus, start = 5, end = 9))

## spliced
SNP_intersect_spliced_dominate <- SNP_37199[intersect_spliced_dominate]
# transcript_strand  +
SNP_intersect_spliced_dominate_plus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="+")] 

extended_range_intersect_spliced_dominate_plus <- resize(SNP_intersect_spliced_dominate_plus, width = 11, fix = "center")

# getseq
sequence_intersect_spliced_dominate_plus <- getSeq(genome, extended_range_intersect_spliced_dominate_plus)

sequence_intersect_spliced_dominate_plus <- subseq(sequence_intersect_spliced_dominate_plus, start = 3, end = 7)

# transcript_strand  -
SNP_intersect_spliced_dominate_minus <-SNP_intersect_spliced_dominate[which(SNP_intersect_spliced_dominate$transcript_strand=="-")] 

extended_range_intersect_spliced_dominate_minus <- resize(SNP_intersect_spliced_dominate_minus, width = 11, fix = "center")

# getseq
sequence_intersect_spliced_dominate_minus <- getSeq(genome, extended_range_intersect_spliced_dominate_minus)

sequence_intersect_spliced_dominate_minus <- (subseq(sequence_intersect_spliced_dominate_minus, start = 5, end = 9))






## not 
SNP_intersect_not_dominate <- SNP_37199[index_not_dominate]
# transcript_strand  +
SNP_intersect_not_dominate_plus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="+")] 


extended_range_intersect_not_dominate_plus <- resize(SNP_intersect_not_dominate_plus, width = 11, fix = "center")

# getseq
sequence_intersect_not_dominate_plus <- getSeq(genome, extended_range_intersect_not_dominate_plus)

sequence_intersect_not_dominate_plus <- subseq(sequence_intersect_not_dominate_plus, start = 3, end = 7)

# transcript_strand  -
SNP_intersect_not_dominate_minus <-SNP_intersect_not_dominate[which(SNP_intersect_not_dominate$transcript_strand=="-")] 

extended_range_intersect_not_dominate_minus <- resize(SNP_intersect_not_dominate_minus, width = 11, fix = "center")

# getseq
sequence_intersect_not_dominate_minus <- getSeq(genome, extended_range_intersect_not_dominate_minus)

sequence_intersect_not_dominate_minus <- (subseq(sequence_intersect_not_dominate_minus, start = 5, end = 9))


sequence_intersect_pre_dominate_minus_rc <- reverseComplement(sequence_intersect_pre_dominate_minus)
sequence_intersect_pre_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus_rc))
sequence_intersect_pre_dominate_plus_rc <- reverseComplement(sequence_intersect_pre_dominate_plus)
sequence_intersect_pre_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus_rc))


sequence_intersect_pre_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_plus))
sequence_intersect_pre_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_pre_dominate_minus))



sequence_intersect_spliced_dominate_minus_rc <- reverseComplement(sequence_intersect_spliced_dominate_minus)
sequence_intersect_spliced_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus_rc))
sequence_intersect_spliced_dominate_plus_rc <- reverseComplement(sequence_intersect_spliced_dominate_plus)
sequence_intersect_spliced_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus_rc))


sequence_intersect_spliced_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_plus))
sequence_intersect_spliced_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_spliced_dominate_minus))

sequence_intersect_not_dominate_minus_rc <- reverseComplement(sequence_intersect_not_dominate_minus)
sequence_intersect_not_dominate_minus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus_rc))
sequence_intersect_not_dominate_plus_rc <- reverseComplement(sequence_intersect_not_dominate_plus)
sequence_intersect_not_dominate_plus_rc <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus_rc))



sequence_intersect_not_dominate_plus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_plus))
sequence_intersect_not_dominate_minus <- chartr("T", "U", as.character(sequence_intersect_not_dominate_minus))




sequence_intersect_pre_dominate <- c(sequence_intersect_pre_dominate_plus,sequence_intersect_pre_dominate_minus_rc)

sequence_intersect_spliced_dominate <- c(sequence_intersect_spliced_dominate_plus,sequence_intersect_spliced_dominate_minus_rc)

sequence_intersect_not_dominate <- c(sequence_intersect_not_dominate_plus,sequence_intersect_not_dominate_minus_rc)



library(ggseqlogo)
seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate),
    "pre_dominate_plus" = as.character(sequence_intersect_pre_dominate_plus),
    "spliced_dominate_plus" = as.character(sequence_intersect_spliced_dominate_plus),
    "not_dominate_plus" = as.character(sequence_intersect_not_dominate_plus),
    "pre_dominate_minus" = as.character(sequence_intersect_pre_dominate_minus),
    "spliced_dominate_minus" = as.character(sequence_intersect_spliced_dominate_minus),
    "not_dominate_minus" = as.character(sequence_intersect_not_dominate_minus)
)
ggseqlogo(seq_list, method = 'bits') + 
    facet_wrap(~seq_group, ncol = 3, scales = "free_x")

seq_list <- list(
    "pre_dominate" = as.character(sequence_intersect_pre_dominate),
    "spliced_dominate" = as.character(sequence_intersect_spliced_dominate),
    "not_dominate" = as.character(sequence_intersect_not_dominate))


library(universalmotif)

motif_list <- lapply(seq_list,function(x) create_motif(x, alphabet = "RNA"))
# bits_matrix_list <- lapply(motif_list,function(x) convert_type(x, "ICM")@motif)
# 
# bits_matrix_list <- lapply(bits_matrix_list,function(x)log2(x+1))
# 
# ggseqlogo(bits_matrix_list, method = 'custom', seq_type = "DNA") + 
#     facet_wrap(~seq_group, ncol = 3, scales = "free_x") 
ggseqlogo(seq_list, method = 'bits', seq_type = "RNA") + 
     facet_wrap(~seq_group, ncol = 3, scales = "free_x")   +
  theme_grey() 

ggseqlogo(seq_list, method = 'prob', seq_type = "RNA") + 
     facet_wrap(~seq_group, ncol = 3, scales = "free_x")  +
  theme_grey() 
```




#### find the index of site only detect methylated reads in spliced or pre mRNA
```{r}
index_meth_only_detect_pre <-  which(res_preprocess[["vg.test"]] %in% res_preprocess[["index"]][["methylation only occurs at unspliced RNA"]])

index_meth_only_detect_spliced <-  which(res_preprocess[["vg.test"]] %in%res_preprocess[["index"]][["methylation only occurs at spliced RNA"]])

# only detect at spliced
index_meth_only_detect_spliced_int <-  intersect(index_meth_only_detect_spliced,intersect_spliced_dominate)
SNP_37199[index_meth_only_detect_spliced_int]

# + transcript strand need minus 1; - transcript strand need plus 1
#  chr1  19242390 EMC1 -
#  chr7  23140899 KLHL7 +
#  chr2  32678876 TTC27 +


# only detect at pre
index_meth_only_detect_pre_int <- intersect(index_meth_only_detect_pre,intersect_pre_dominate)
SNP_37199[index_meth_only_detect_pre_int]

#  chr3 119523752 TIMMDC1  +
#  chr2 208248516  IDH1  -
#  chr17  62448721  METTL2A  +
#  chr1  28407422  PHACTR4  +
#  chr15  41365455  NUSAP1  +
#  chr16  67616757  CTCF  +
```

### cell cycle trajectory using m6A velocity
#### find the 9 marker gens sites in 37199
```{r eval=FALSE, include=FALSE}


library(ggplot2)
library(svglite)
library(VeloRM)
## step1 : choose gene
g1_s_genes <- c("MCM6", "PCNA", "SLBP")
s_genes <- c( "RRM2", "DTL","MCM5")
g2_m_genes <- c("TOP2A",  "CCNB1", "AURKA" )






library(org.Hs.eg.db)
library(clusterProfiler)
library(stringi)



g1_s_genes_ENSG <- bitr(g1_s_genes, fromType="SYMBOL", toType="ENSEMBL", OrgDb="org.Hs.eg.db")
s_genes_ENSG <- bitr(s_genes, fromType="SYMBOL", toType="ENSEMBL", OrgDb="org.Hs.eg.db")
g2_m_genes_ENSG <-bitr(g2_m_genes, fromType="SYMBOL", toType="ENSEMBL", OrgDb="org.Hs.eg.db")


## step3 : select the tpm data
select_gene <- c(g1_s_genes_ENSG$ENSEMBL,s_genes_ENSG$ENSEMBL,g2_m_genes_ENSG$ENSEMBL)


library(biomaRt)

# Connect to Ensembl
ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")




# Query positional information with strand
pos_info <- getBM(attributes = c("ensembl_gene_id",
                                 "chromosome_name",
                                 "start_position",
                                 "end_position",
                                 "strand"),
                  filters = "ensembl_gene_id",
                  values = select_gene,
                  mart = ensembl)

# Convert strand information from numeric to "+" or "-"
pos_info$strand <- ifelse(pos_info$strand == 1, "+", "-")
pos_info <- pos_info[match(select_gene, pos_info$ensembl_gene_id), ]

# Print the ordered positional information
print(pos_info)


library(ggplot2)


SNP_37199 <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/SNP_37199.rds")

# find the 9 gens site

# the sites position
testSNP_all_position <- SNP_37199@ranges@start

# match the site to the gene
index_list <- list()
for ( i in 1:length(select_gene)){
  index_list[[i]] <-which(testSNP_all_position%in%(pos_info$start_position[i]:pos_info$end_position[i]))
  if(length(index_list[[i]])==0){

  }else{
    names(index_list[[i]]) <- paste0("gene_",i,"_",1:length(index_list[[i]]))
  }

}



# all index of sites
index_site <- unlist(index_list)
site_length <- unlist(lapply(index_list, length))


chr_info <- rep(SNP_37199@seqnames@values,SNP_37199@seqnames@lengths)
chr_info_get <- chr_info[index_site]
chr_info_target <- factor(rep(paste0("chr",pos_info$chromosome_name),site_length),levels = levels(chr_info_get))

index_correct <- which(chr_info_get ==chr_info_target)
site_length_correct <- list()
start <- 1
end <- site_length[1]
for (i in 1:8){
  site_length_correct[[i]] <- length(which(index_correct%in%c(start:end)))
  start <- start+site_length[i]
  end <- end + site_length[i+1]
}
site_length_correct[[9]] <- length(which(index_correct%in%c(start:end)))
site_index_correct <- index_site [index_correct]

index_site <- site_index_correct
index_site_SNP <- SNP_37199[index_site]
site_length <- unlist(site_length_correct)
names(site_length) <- c(g1_s_genes,s_genes,g2_m_genes )
saveRDS(index_site,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site.rds")
saveRDS(index_site_SNP,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site_SNP.rds")
saveRDS(site_length,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/site_length.rds")
```
#### run the m6A velocity and expression velocity
```{r eval=FALSE, include=FALSE}

res_preprocess <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/res_preprocess.rds")
# the input data
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/test.list.rds")
control.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/control.list.rds")


spliced.meth.test <- test.list[[1]]
spliced.unmeth.test <- test.list[[2]]
unspliced.meth.test <- test.list[[3]]
unspliced.unmeth.test <- test.list[[4]]

kCells = 1; deltaT = 0.01; kSites = 1
library(VeloRM)
# calculate the emat_size and nmat_size

emat_size <- colSums(spliced.meth.test + spliced.unmeth.test)
nmat_size <- colSums(unspliced.meth.test + unspliced.unmeth.test)

res <- methylation.site.relative.velocity.estimates(res_preprocess$test, res_preprocess$control,
                                                    kCells = kCells, deltaT = deltaT, kSites = kSites, 
                                                    fit.quantile = 0.05, n.cores = 1, emat.size = emat_size, 
                                                    nmat.size = nmat_size)

saveRDS(res, "/gpfs/work/bio/haozhewang17/data/VeloRM_data/results/res_m6A_velocity.rds")


## run the velocity


# the input data
test.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/test.list.rds")
control.list <- readRDS("/gpfs/work/bio/haozhewang17/data/VeloRM_data/pre_process/control.list.rds")


spliced.meth.test <- test.list[[1]]
spliced.unmeth.test <- test.list[[2]]
unspliced.meth.test <- test.list[[3]]
unspliced.unmeth.test <- test.list[[4]]

kCells = 1; deltaT = 0.01; kSites = 1
library(VeloRM)
# calculate the emat_size and nmat_size

emat_size <- colSums(spliced.meth.test + spliced.unmeth.test)
nmat_size <- colSums(unspliced.meth.test + unspliced.unmeth.test)

res <- site.relative.velocity.estimates(emat = spliced.meth.test+spliced.unmeth.test, 
                                        nmat = unspliced.meth.test+unspliced.unmeth.test,
                                                    kCells = kCells, deltaT = deltaT, kSites = kSites, 
                                                    fit.quantile = 0.05, n.cores = 1, emat.size = emat_size, 
                                                    nmat.size = nmat_size)

saveRDS(res, "/gpfs/work/bio/haozhewang17/data/VeloRM_data/results/res_expression_velocity.rds")

```



#### OR embedding
filter the sites
```{r eval=FALSE, include=FALSE}
library(ggplot2)

res <- readRDS("D:/data/VeloRM_data/scDART-seq/results/res_m6A_velocity.rds")
## PCA

label_SigRM <- readRDS("D:/data/VeloRM_data/scDART-seq/origin/label_SigRM.rds")

label_phase <- rep(NA,1209)
names(label_phase) <- colnames(res[[1]][["conv.nmat.norm"]])

label_phase[names(label_SigRM)] <- label_SigRM
index <- which(label_phase=="G1"|label_phase=="G2/M"|label_phase=="S")

# OR
index_site <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site.rds")
site_length <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/site_length.rds")
res_preprocess <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_preprocess.rds")
index_site_name <- res_preprocess[["vg.test"]][index_site]

index_site_name_int <- intersect(index_site_name,rownames(res[[3]][["current"]]$OR))

spliced_methylation_OR_value_meta_index <- res_preprocess[["spliced_methylation_OR_value_meta"]][index_site_name_int]
unspliced_methylation_OR_value_meta_index <-res_preprocess[["unspliced_methylation_OR_value_meta"]][index_site_name_int]


index_site_name_int_intersect <-index_site_name_int[which(spliced_methylation_OR_value_meta_index>=1&
                               unspliced_methylation_OR_value_meta_index>=1)] 


index_site_name_int_intersect_gene_info <- rep(names(site_length),site_length)[which(index_site_name%in%index_site_name_int_intersect)]

saveRDS(index_site_name_int_intersect,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site_name_int_intersect.rds")
```

##### OR embedding tsne
```{r warning=FALSE}
rm(list = ls())
gc()
res <- readRDS("D:/data/VeloRM_data/scDART-seq/results/res_m6A_velocity.rds")
index_site_name_int_intersect <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site_name_int_intersect.rds")
label_SigRM <- readRDS("D:/data/VeloRM_data/scDART-seq/origin/label_SigRM.rds")

label_phase <- rep(NA,1209)
names(label_phase) <- colnames(res[[1]][["conv.nmat.norm"]])

label_phase[names(label_SigRM)] <- label_SigRM
index <- which(label_phase=="G1"|label_phase=="G2/M"|label_phase=="S")
library(Rtsne)



data_scaled <- t(res[[3]][["current"]]$OR[index_site_name_int_intersect,])

data_scaled <- scale(data_scaled)



# Run t-SNE
set.seed(42)
tsne_result <- Rtsne(data_scaled, dims = 2, perplexity = 110,theta = 0.3,  max_iter = 1000,verbose = TRUE) # 100, 0.2, 1000/   # 100, 0.3, 1000, 

# View t-SNE embeddings
head(tsne_result$Y)
label_phase_ <- label_phase
label_phase_[label_phase=="unclear"] <- NA

tsne_data <- data.frame(cbind(tsne_result$Y,label_phase_))
colnames(tsne_data) <- c("x","y","color")
tsne_data$x <- as.numeric(tsne_data$x)
tsne_data$y <- as.numeric(tsne_data$y)

ggplot()+geom_point(data=tsne_data[index,],aes(x=x,y=y,color=color))
ggplot()+geom_point(data=tsne_data,aes(x=x,y=y,color=color))



#arrow.scale=5; cell.alpha=0.4; cell.cex=1; fig.height=4;
emb_tsne_OR_intersect <- tsne_result$Y
rownames(emb_tsne_OR_intersect) <- colnames(res[[3]][["current"]]$OR)

library(VeloRM)
arrow.scale=5; cell.alpha=0.4; cell.cex=1; fig.height=4;
vis_tsne_OR_emb_OR_log_index <- show.velocity.on.embedding.cor(emb_tsne_OR_intersect,
             res[[3]][["current"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$OR[index_site_name_int_intersect,index],
             cell.colors=label_phase[index],n=30,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5)
vis_tsne_OR_emb_OR_log_index[[1]]
arrow.scale=3;
vis_tsne_OR_emb_OR_log_index_grid <- show.velocity.on.embedding.cor(emb_tsne_OR_intersect,
             res[[3]][["current"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$OR[index_site_name_int_intersect,index],
             cell.colors=label_phase[index],n=30,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,show.grid.flow = TRUE,grid.n = 12)
vis_tsne_OR_emb_OR_log_index_grid[[1]]




# saveRDS(emb_tsne_OR_intersect,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/emb_tsne_OR_intersect.rds")
# saveRDS(vis_tsne_OR_emb_OR_log_index,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_tsne_OR_emb_OR_log_index.rds")
# saveRDS(vis_tsne_OR_emb_OR_log_index_grid,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_tsne_OR_emb_OR_log_index_grid.rds")
```

##### OR embedding UMAP
```{r,warning=FALSE}
library(uwot)


label_phase_ <- label_phase
label_phase_[label_phase=="unclear"] <- NA
# Run UMAP
umap_result <- umap(t(res[[3]][["current"]]$OR[index_site_name_int_intersect,]),n_neighbors=30,min_dist=1.5,seed=42,pca=5)

# View UMAP embeddings
head(umap_result)



umap_data <- data.frame(cbind(umap_result,label_phase_))
colnames(umap_data) <- c("x","y","color")
umap_data$x <- as.numeric(umap_data$x)
umap_data$y <- as.numeric(umap_data$y)

ggplot()+geom_point(data=umap_data[index,],aes(x=x,y=y,color=color))
ggplot()+geom_point(data=umap_data,aes(x=x,y=y,color=color))



emb_umap_OR_intersect <- umap_result
rownames(emb_umap_OR_intersect) <- colnames(res[[3]][["current"]]$OR)

library(VeloRM)
arrow.scale=5; cell.alpha=0.4; cell.cex=1; fig.height=4;
vis_umap_OR_emb_OR_log_index <- show.velocity.on.embedding.cor(emb_umap_OR_intersect,
             res[[3]][["current"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$OR[index_site_name_int_intersect,index],
             cell.colors=label_phase[index],n=30,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5)
vis_umap_OR_emb_OR_log_index[[1]]
arrow.scale=5
vis_umap_OR_emb_OR_log_index_grid <- show.velocity.on.embedding.cor(emb_umap_OR_intersect,
             res[[3]][["current"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["projected"]]$OR[index_site_name_int_intersect,index],
             res[[3]][["delta"]]$OR[index_site_name_int_intersect,index],
             cell.colors=label_phase[index],n=30,scale='log',
             cex=cell.cex,arrow.scale=arrow.scale,arrow.lwd=0.5,show.grid.flow = TRUE,grid.n = 11)
vis_umap_OR_emb_OR_log_index_grid[[1]]

# saveRDS(emb_umap_OR_intersect,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/emb_umap_OR_intersect.rds")
# saveRDS(vis_umap_OR_emb_OR_log_index,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_umap_OR_emb_OR_log_index.rds")
# saveRDS(vis_umap_OR_emb_OR_log_index_grid,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_umap_OR_emb_OR_log_index_grid.rds")
```



#### diffusion

##### tsne
```{r eval=FALSE, include=FALSE}
library(VeloRM)
# read the velocity results
res <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_1_site_1_delta_0.01/res.rds")

# read the visulaization results
vis_tsne_OR_emb_OR_log_index <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_tsne_OR_emb_OR_log_index.rds")
vis_tsne_OR_emb_OR_log_index <- vis_tsne_OR_emb_OR_log_index[[2]]

# read the embedding
emb_tsne_OR_intersect <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/emb_tsne_OR_intersect.rds")

# read the name of the sites
index_site_name_int_intersect <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site_name_int_intersect.rds")


# read the cell label 
label_SigRM <- readRDS("D:/data/VeloRM_data/scDART-seq/origin/label_SigRM.rds")

label_phase <- rep(NA,1209)
names(label_phase) <- colnames(res[[1]][["conv.nmat.norm"]])

label_phase[names(label_SigRM)] <- label_SigRM
index <- which(label_phase=="G1"|label_phase=="G2/M"|label_phase=="S")

diffusion_density_plot_tsne <- diffusion_density_plot(emb_tsne_OR_intersect[index,],vis_tsne_OR_emb_OR_log_index$tp,forward = TRUE,2500,scale="log")
saveRDS(diffusion_density_plot_tsne,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/diffusion_density_plot_tsne.rds")
```
###### visulization
```{r}
diffusion_density_plot_tsne <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/diffusion_density_plot_tsne.rds")
diffusion_density_plot_tsne
```

##### umap
```{r eval=FALSE, include=FALSE}
library(VeloRM)
# read the velocity results
res <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_1_site_1_delta_0.01/res.rds")

# read the visulaization results
vis_umap_OR_emb_OR_log_index <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/vis_umap_OR_emb_OR_log_index.rds")
vis_umap_OR_emb_OR_log_index <- vis_umap_OR_emb_OR_log_index[[2]]

# read the embedding
emb_umap_OR_intersect <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/emb_umap_OR_intersect.rds")

# read the name of the sites
index_site_name_int_intersect <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/index_site_name_int_intersect.rds")


# read the cell label 
label_SigRM <- readRDS("D:/data/VeloRM_data/scDART-seq/origin/label_SigRM.rds")

label_phase <- rep(NA,1209)
names(label_phase) <- colnames(res[[1]][["conv.nmat.norm"]])

label_phase[names(label_SigRM)] <- label_SigRM
index <- which(label_phase=="G1"|label_phase=="G2/M"|label_phase=="S")

diffusion_density_plot_umap <- diffusion_density_plot(emb_umap_OR_intersect[index,],vis_umap_OR_emb_OR_log_index$tp,forward = TRUE,2500,scale="log")
saveRDS(diffusion_density_plot_umap,"D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/diffusion_density_plot_umap.rds")
```
###### visulization
```{r}
diffusion_density_plot_umap <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_cycle_sites/diffusion_density_plot_umap.rds")
diffusion_density_plot_umap
```
#### site analysis
```{r,warning=FALSE}
median_OR <- apply(res[[3]][["current"]][["OR"]][index_site_name_int_intersect,index],1,median)

ind <- which.max(median_OR)

data=tsne_data[index,1:2]
data$OR = res[[3]][["current"]][["OR"]][index_site_name_int_intersect[ind],index]

plot_tsne_site120 <- ggplot() +
  geom_point(data = data, aes(x = x, y = y, color = OR), size = 3, alpha = 0.8) +
  scale_color_gradientn(
    colors = c("grey90", "grey", "#5AB4AC", "#01665E", "#003C30"),
    values = scales::rescale(c(0, 0.3, 0.6, 0.8, 1)),
    limits = c(min(data$OR), max(data$OR)),
    name = "Odds Ratio",
    guide = guide_colorbar(barwidth = 1, barheight = 15)
  ) +
  labs(title = NULL)+xlab("PC1")+ylab("PC2")

data=umap_data[index,1:2]
data$OR = res[[3]][["current"]][["OR"]][index_site_name_int_intersect[ind],index]
plot_tsne_site120

plot_umap_site120 <- ggplot() +
  geom_point(data = data, aes(x = x, y = y, color = OR), size = 3, alpha = 0.8) +
  scale_color_gradientn(
    colors = c("grey90", "grey", "#5AB4AC", "#01665E", "#003C30"),
    values = scales::rescale(c(0, 0.3, 0.6, 0.8, 1)),
    limits = c(min(data$OR), max(data$OR)),
    name = "Odds Ratio",
    guide = guide_colorbar(barwidth = 1, barheight = 15)
  ) +
  labs(title = NULL)+xlab("PC1")+ylab("PC2")
plot_umap_site120


data <- as.data.frame(cbind(res[[1]][["conv.emat.norm"]][index_site_name_int_intersect[ind],index],
              res[[1]][["conv.nmat.norm"]][index_site_name_int_intersect[ind],index],
              tsne_data [index,3]))
colnames(data) <- c("x","y","color")
data$x <- as.numeric(data$x)
data$y <- as.numeric(data$y)

slope <- res[[1]][["gamma"]][index_site_name_int_intersect[ind]]
intercept <- res[[1]][["ko"]][index_site_name_int_intersect[ind],1]

plot_site120_meth <- ggplot()+geom_point(data=data,aes(x=x,y=y),size=2,color="#2E5DAA") +
  geom_abline(slope = slope, intercept = intercept, color = "black", size = 1,linetype = "dashed")+
  coord_cartesian(xlim = c(0,0.02),ylim = c(0,0.06))+ylab("methylated unspliced")+xlab("methylated spliced")
plot_site120_meth


data <- as.data.frame(cbind(res[[2]][["conv.emat.norm"]][index_site_name_int_intersect[ind],index],
              res[[2]][["conv.nmat.norm"]][index_site_name_int_intersect[ind],index],
              tsne_data [index,3]))
colnames(data) <- c("x","y","color")
data$x <- as.numeric(data$x)
data$y <- as.numeric(data$y)
slope <- res[[2]][["gamma"]][index_site_name_int_intersect[ind]]
intercept <- res[[2]][["ko"]][index_site_name_int_intersect[ind],1]

plot_site120_unmeth <- ggplot()+geom_point(data=data,aes(x=x,y=y),size=2,color="#2E5DAA") +
  geom_abline(slope = slope, intercept = intercept, color = "black", size = 1,linetype = "dashed")+ylab("unmethylated unspliced")+xlab("unmethylated spliced")
plot_site120_unmeth


unspliced_M_current <- log2(((res[[1]][["conv.nmat.norm"]][index_site_name_int_intersect[ind],index]+1e-7)/
  (res[[2]][["conv.nmat.norm"]][index_site_name_int_intersect[ind],index]+1e-7))/res[[4]][["control_information_Beta"]][index_site_name_int_intersect[ind]])

spliced_M_current <- log2(((res[[1]][["conv.emat.norm"]][index_site_name_int_intersect[ind],index]+1e-7)/
                           (res[[2]][["conv.emat.norm"]][index_site_name_int_intersect[ind],index]+1e-7))/res[[4]][["control_information_Beta"]][index_site_name_int_intersect[ind]])

data <- as.data.frame(cbind(spliced_M_current,
                            unspliced_M_current,
                            tsne_data [index,3]))
colnames(data) <- c("x","y","color")
data$x <- as.numeric(data$x)
data$y <- as.numeric(data$y)

threshold <- 0

data_steady <- data[which(data$x*data$y>threshold^2),]
lm_res <- lm(y~x,data_steady)
lm_res$coefficients
slope <- lm_res$coefficients[2]
intercept <- lm_res$coefficients[1]

data_steady_left <- data[which(data$x<threshold&data$y<threshold),]
data_steady_right <- data[which(data$x>threshold&data$y>threshold),]

data_left <- data[which(data$x<threshold&data$y>threshold),]
data_right <- data[which(data$x>threshold&data$y<threshold),]


residuals_value_left <- predict(lm_res,data_left)-data_left$y
residuals_value_right <- predict(lm_res,data_right)-data_right$y

library(ggforce)
plot_site120_OR <- ggplot()+geom_point(data=data,aes(x=x,y=y),size=2,color="#2E5DAA") +
  geom_abline(slope = slope, intercept = intercept, color = "#555555", size = 1,linetype = "dashed")+
  geom_line(
    stat = "smooth",
    data = data_left,
    aes(x = x, y = y),
    method = "lm",
    formula = y ~ poly(x, 2),
    color = "#555555",
    size = 1,
    arrow = arrow(type = "closed", length = unit(0.15, "inches"))
  )+geom_line(
    stat = "smooth",
    data = data_right,
    aes(x = x, y = y),
    method = "lm",
    formula = y ~ poly(x, 2),
    color = "#555555",
    size = 1,
    arrow = arrow(type = "closed", length = unit(0.15, "inches"), ends = "first")
  )+stat_ellipse(data = data_steady_left, level = 0.95, linetype = 1, size = 2, 
    fill = "#2C7BB6",
                 aes(x = x, y = y), alpha = 0.3, segments = 50, geom = "polygon",type = "t") +stat_ellipse(data = data_steady_right, level = 0.99, linetype = 1, size = 2, 
               fill = "#2C7BB6",  aes(x = x, y = y), alpha = 0.3, segments = 50, geom = "polygon",type = "t") +ylab("unspliced OR")+xlab("spliced OR")
plot_site120_OR




```
#### transcriptional impact analysis 

##### run the transcriptional_impact_analysis
```{r eval=FALSE, include=FALSE}
library(VeloRM)
res_preprocess <- readRDS("D:/data/VeloRM_data/scDART-seq/pre_process/res_preprocess.rds")

test.list <- res_preprocess[[1]]
control.list <- res_preprocess[[2]]

res_methylation <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_1_site_1_delta_0.01/res.rds")
res_expression <- readRDS("D:/data/VeloRM_data/scDART-seq/results/cell_1_site_1_delta_0.01/res_expression.rds")


res_transcriptional_impact_analysis <- transcriptional.impact.analysis(test.list = test.list,
                                                                       control.list = control.list,
                                                                       methylation.site.relative.velocity.result = res_methylation,
                                                                       exp.site.velocity.result = res_expression,delta_exp = "delta_from_exp_velocity",delta_exp_TPM = FALSE,methylation_level = "meta_OR_sep",methylation_level_scale = "log2",cor_method="spearman")

saveRDS(res_transcriptional_impact_analysis,"D:/data/VeloRM_data/scDART-seq/results/transcriptional_impact_analysis/res_transcriptional_impact_analysis.rds")

```

##### analysis the res from transcriptional_impact_analysis
###### comparision spliced unspliced corelation
```{r}
rm(list = ls())
gc()
res_transcriptional_impact_analysis <- readRDS("D:/data/VeloRM_data/scDART-seq/results/transcriptional_impact_analysis/res_transcriptional_impact_analysis_spearman.rds")

delta_exp_and_methylation_cor_spliced <- mapply(function(meth, exp) {
    cor(meth, exp, method = "spearman")
  }, as.data.frame(t(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]])), 
  as.data.frame(t(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]])))


delta_exp_and_methylation_cor_unspliced <- mapply(function(meth, exp) {
    cor(meth, exp, method = "spearman")
  }, as.data.frame(t(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]])), 
  as.data.frame(t(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]])))


library(ggplot2)

# Create a data frame with both vectors and group labels
df <- data.frame(
  correlation = c(
    delta_exp_and_methylation_cor_spliced,
    delta_exp_and_methylation_cor_unspliced,
    res_transcriptional_impact_analysis[["delta_exp_and_methylation_cor"]]
  ),
  group = factor(
    c(
      rep("Spliced", length(delta_exp_and_methylation_cor_spliced)),
      rep("Unspliced", length(delta_exp_and_methylation_cor_unspliced)),
      rep("Meta", length(delta_exp_and_methylation_cor_unspliced))
    ),
    levels = c("Spliced", "Unspliced","Meta")
  )
)

p <- ggplot(df, aes(x = correlation, fill = group, color = group)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )+
  theme(
    legend.position = c(0.8, 0.7)  )
print(p)
```



##### analysis the methylation level and delta expression
```{r}
delta_exp_methylation_cor <- res_transcriptional_impact_analysis$delta_exp_and_methylation_cor


# find the site with high corelation 
cor_threshold_p <- 0.2
cor_threshold_n <- -0.2

index_delta_exp_methylation_cor_p <- which(delta_exp_methylation_cor>=cor_threshold_p)

index_delta_exp_methylation_cor_n <- which(delta_exp_methylation_cor<=cor_threshold_n)


# get the coefficients

lm_results_index_delta_exp_methylation_cor_coefficients_intercept <-
  unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients, function(x) x[1]))

lm_results_index_delta_exp_methylation_cor_coefficients_spliced <- unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients, function(x) x[2]))

lm_results_index_delta_exp_methylation_cor_coefficients_pre <- unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients, function(x) x[3]))


# get the coefficients pvalue

lm_results_index_delta_exp_methylation_cor_coefficients_intercept_pvalue <-
  unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients_pvalues, function(x) x[1]))

lm_results_index_delta_exp_methylation_cor_coefficients_spliced_pvalue <- unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients_pvalues, function(x) x[2]))

lm_results_index_delta_exp_methylation_cor_coefficients_pre_pvalue <- unlist(lapply(res_transcriptional_impact_analysis$delta_exp_lm_coefficients_pvalues, function(x) x[3]))

# all p-value need to significant
index_delta_exp_pvalue_significant <- which(lm_results_index_delta_exp_methylation_cor_coefficients_intercept_pvalue<=0.05&lm_results_index_delta_exp_methylation_cor_coefficients_spliced_pvalue<=0.05&lm_results_index_delta_exp_methylation_cor_coefficients_pre_pvalue<=0.05)

# for the evaluate metrics it should not be much worse than the training set
# use RMSE
delta_exp_train_RMSE <- res_transcriptional_impact_analysis$delta_exp_metrics_train_df$RMSE 
delta_exp_test_RMSE <- res_transcriptional_impact_analysis$delta_exp_metrics_test_df$RMSE

index_delta_exp_RMSE_significant <- which(delta_exp_test_RMSE>=delta_exp_train_RMSE*0.9&
                                  delta_exp_test_RMSE<=delta_exp_train_RMSE*1.1)

index_delta_exp_filter <- intersect(index_delta_exp_RMSE_significant,index_delta_exp_pvalue_significant)



```

###### positive correlation
```{r}
index_delta_exp_p_filter <- intersect(index_delta_exp_filter,index_delta_exp_methylation_cor_p)

# find the spliced dominant


index_delta_exp_spliced_dominant_p <-index_delta_exp_p_filter[which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_p_filter]>lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_p_filter]&lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_p_filter]>0)] 

num_delta_exp_spliced_dominant_p <- length(index_delta_exp_spliced_dominant_p)



# find the unspliced dominant 
index_delta_exp_pre_dominant_p <- index_delta_exp_p_filter[which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_p_filter]<lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_p_filter]&lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_p_filter]>0)]

num_delta_exp_pre_dominant_p <- length(index_delta_exp_pre_dominant_p)

num_delta_exp_spliced_dominant_p
num_delta_exp_pre_dominant_p
```
####### spliced dominant 
```{r}
length(which(lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_spliced_dominant_p]<0))

length(which(lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_spliced_dominant_p]>0))

ind <- which.max(res_transcriptional_impact_analysis[["delta_exp_and_methylation_cor"]][index_delta_exp_spliced_dominant_p])

# all the spliced dominant, pre mRNA is differential compare to spliced mRNA

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix"]][index_delta_exp_spliced_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_meta <- ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-48,-36),ylim=c(-0.001,0.001))+
  ylab("delta exp")+xlab("meta OR")+
                theme_minimal()
plot_meta

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_spliced_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_spliced <- ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-26,-17),ylim=c(-0.001,0.001))+
  ylab("delta exp")+xlab("spliced OR")+
                theme_minimal()
plot_spliced




data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_spliced_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_unspliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-23,-17),ylim=c(-0.001,0.001))+
  ylab("delta exp")+xlab("unspliced OR")+
                theme_minimal()
plot_unspliced









# Create data frames for both groups
data_spliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_spliced_dominant_p[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_p[ind],]),
    group = "Spliced"
)

data_unspliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_spliced_dominant_p[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_p[ind],]),
    group = "Unspliced"
)

# Combine the data
combined_data <- rbind(data_spliced, data_unspliced)

# Calculate correlations for each group
cor_spliced <- cor(data_spliced$x, data_spliced$y, method = "spearman")
cor_unspliced <- cor(data_unspliced$x, data_unspliced$y, method = "spearman")

# Create combined plot
combined_plot <- ggplot(combined_data, aes(x = x, y = y, color = group)) +
    geom_point(size = 1, alpha = 0.5) +
    coord_cartesian(xlim = c(-26, -17),ylim = c(-0.001,0.001)) +
    ylab("delta exp") +
    xlab("OR methylation level") +
    labs(color = "Group")  +
                theme_minimal()+
  theme(
    legend.position = c(0.8, 0.8)  )
            
combined_plot

```
####### pre dominant 
```{r}
length(which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_pre_dominant_p]<0))

length(which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_pre_dominant_p]>0))

# all the pre dominant, spliced mRNA is differential compare to spliced mRNA
ind <- which.max(res_transcriptional_impact_analysis[["delta_exp_and_methylation_cor"]][index_delta_exp_pre_dominant_p])

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix"]][index_delta_exp_pre_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_meta <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(23.5,35))+
  ylab("delta exp")+xlab("meta OR")+
                theme_minimal()
plot_meta


data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_pre_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_spliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(11,18))+
  ylab("delta exp")+xlab("spliced OR")+
                theme_minimal()

plot_spliced

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_pre_dominant_p[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_p[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_unspliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(11,18))+
  ylab("delta exp")+xlab("unspliced OR")+
                theme_minimal()
plot_unspliced





# Create data frames for both groups
data_spliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_pre_dominant_p[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_p[ind],]),
    group = "Spliced"
)

data_unspliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_pre_dominant_p[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_p[ind],]),
    group = "Unspliced"
)

# Combine the data
combined_data <- rbind(data_spliced, data_unspliced)

# Calculate correlations for each group
cor_spliced <- cor(data_spliced$x, data_spliced$y, method = "spearman")
cor_unspliced <- cor(data_unspliced$x, data_unspliced$y, method = "spearman")

# Create combined plot
combined_plot <- ggplot(combined_data, aes(x = x, y = y, color = group)) +
    geom_point(size = 1, alpha = 0.5) +
    coord_cartesian(xlim = c(11,18)) +
    ylab("delta exp") +
    xlab("OR methylation level") +
    labs(color = "Group")  +
                theme_minimal()
            
combined_plot

```
###### negative correlation
```{r}
index_delta_exp_n_filter <- intersect(index_delta_exp_filter,index_delta_exp_methylation_cor_n)

# find the spliced dominant
index_delta_exp_spliced_dominant_n <-index_delta_exp_n_filter[which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_n_filter]<lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_n_filter]&lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_n_filter]<0)] 

num_delta_exp_spliced_dominant_n <- length(index_delta_exp_spliced_dominant_n)
  
# find the unspliced dominant 
index_delta_exp_pre_dominant_n <- index_delta_exp_n_filter[which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_n_filter]>lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_n_filter]&lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_n_filter]<0)]

num_delta_exp_pre_dominant_n <- length(index_delta_exp_pre_dominant_n)

num_delta_exp_spliced_dominant_n
num_delta_exp_pre_dominant_n
```
####### spliced dominant 
```{r}
length(which(lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_spliced_dominant_n]<0))

length(which(lm_results_index_delta_exp_methylation_cor_coefficients_pre[index_delta_exp_spliced_dominant_n]>0))

# all the spliced dominant, pre mRNA is same compare to spliced mRNA


ind <- 2

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix"]][index_delta_exp_spliced_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_meta <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-48,-35))+
                theme_minimal()
plot_meta 



data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_spliced_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_spliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-24,-16))+
                theme_minimal()
plot_spliced



data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_spliced_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_unspliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-26,-15))+
                theme_minimal()
plot_unspliced




# Create data frames for both groups
data_spliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_spliced_dominant_n[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_n[ind],]),
    group = "Spliced"
)

data_unspliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_spliced_dominant_n[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_spliced_dominant_n[ind],]),
    group = "Unspliced"
)

# Combine the data
combined_data <- rbind(data_spliced, data_unspliced)

# Calculate correlations for each group
cor_spliced <- cor(data_spliced$x, data_spliced$y, method = "spearman")
cor_unspliced <- cor(data_unspliced$x, data_unspliced$y, method = "spearman")

# Create combined plot
combined_plot <- ggplot(combined_data, aes(x = x, y = y, color = group)) +
    geom_point(size = 1, alpha = 0.5) +
    coord_cartesian(xlim = c(-26,-15)) +
    ylab("delta exp") +
    xlab("OR methylation level") +
    labs(color = "Group")  +
                theme_minimal()
            
combined_plot
```
####### pre dominant 
```{r}
length(which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_pre_dominant_n]<0))

length(which(lm_results_index_delta_exp_methylation_cor_coefficients_spliced[index_delta_exp_pre_dominant_n]>0))

# all the pre dominant, all most all spliced mRNA is differential compare to pre mRNA

ind <- which.min(res_transcriptional_impact_analysis[["delta_exp_and_methylation_cor"]][index_delta_exp_pre_dominant_n])

data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix"]][index_delta_exp_pre_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_meta <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-47,-31))+
                theme_minimal()
plot_meta



data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_pre_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_spliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-25,-15))+
                theme_minimal()
plot_spliced



data_plot <- as.data.frame(cbind(as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_pre_dominant_n[ind],]),
                  as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_n[ind],])))
colnames(data_plot) <- c("x","y")


cor(data_plot$x,data_plot$y,method = "spearman")
plot_unspliced <-ggplot()+geom_point(data=data_plot,aes(x=x,y=y),
                                     size = 1,   
    alpha = 0.5,
    color = "#2c7fb8")+coord_cartesian(xlim=c(-25,-15))+
                theme_minimal()
plot_unspliced



# Create data frames for both groups
data_spliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_spliced"]][index_delta_exp_pre_dominant_n[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_n[ind],]),
    group = "Spliced"
)

data_unspliced <- data.frame(
    x = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["methylation_level_matrix_unspliced"]][index_delta_exp_pre_dominant_n[ind],]),
    y = as.numeric(res_transcriptional_impact_analysis[["other_information"]][["delta_exp_matrix"]][index_delta_exp_pre_dominant_n[ind],]),
    group = "Unspliced"
)

# Combine the data
combined_data <- rbind(data_spliced, data_unspliced)

# Calculate correlations for each group
cor_spliced <- cor(data_spliced$x, data_spliced$y, method = "spearman")
cor_unspliced <- cor(data_unspliced$x, data_unspliced$y, method = "spearman")

# Create combined plot
combined_plot <- ggplot(combined_data, aes(x = x, y = y, color = group)) +
    geom_point(size = 1, alpha = 0.5) +
    coord_cartesian(xlim = c(-25,-15)) +
    ylab("delta exp") +
    xlab("OR methylation level") +
    labs(color = "Group")  +
                theme_minimal()
            
combined_plot
```
##### visulization
```{r}
library(ggplot2)

# Create a data frame for plotting
data <- data.frame(
  Category = c("Positive", "Positive", "Negative", "Negative"),
  Type = c("Spliced", "Pre", "Spliced", "Pre"),
  Value = c(num_delta_exp_spliced_dominant_p, 
             num_delta_exp_pre_dominant_p, 
             num_delta_exp_spliced_dominant_n, 
             num_delta_exp_pre_dominant_n)
)

# Plot the stacked bar chart
delta_exp_vis <- ggplot(data, aes(x = Category, y = log10(Value + 1), fill = Type)) +
  geom_bar(stat = "identity", position = "stack") +  # Default is stack
  labs(title = NULL,
       x = NULL,
       y = "log10 (Count+1)") +
  theme_minimal() +
  theme(legend.position = c(0.73,0.8)) +
  scale_fill_brewer(palette = "Set2")  # Use color palette

delta_exp_vis


```

##### fisher test
```{r}
data <- matrix(c(9, 1, 2, 3074), nrow = 2, byrow = TRUE,
               dimnames = list(Category = c("Positive", "Negative"),
                               Type = c("Spliced", "Pre")))

# Print the table
print(data)

# Perform Fisher's Exact Test
result <- fisher.test(data)

# Print the result
print(result)
```
